---
title: "Length-Frequency Histograms"
author: "Marty Simonson"
date: "December 5, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
```



# Length frequency histograms
There are three main objectives to this markdown:
- overall length frequency histograms of each lake's carp and buffalo populations (10mm bins)
- comparisons betweeen seine hauls and electrofishing size disttributions
- when available, comparisons between 2017 and 2018 length frequency distributions

Much of this code will be absorbed from [[ LengthFrequencies.R ]] where I started this shindig. I'll try to finish it here. 

# 1.) Overall length frequency distributions of carp and buffalo in each of the lakes
```{r}
#reading in data
dframe<-read.csv("2018_COC_BIB_CMR_Data.csv")
dframe<-dframe[,-1]


# splitting data frame by Lakes
levels(dframe$Lake)
Blue <- subset(dframe, Lake == "Blue")
Center<-subset(dframe, Lake == "Center")
FiveIsland<-subset(dframe, Lake == "5 Island")
NTwin<-subset(dframe, Lake == "N. Twin")
Silver<-subset(dframe, Lake == "Silver")
STwin<-subset(dframe, Lake == "S. Twin")
Storm<-subset(dframe, Lake == "Storm")

```

## Blue Lake
```{r}
# No extreme need to split species (one BIB in dataset) but I'll do it anyway
blue.coc<-subset(Blue, Species == "COC" & !is.na(Length..mm.))

summary(blue.coc$Length..mm.)
length(!is.na(blue.coc$Length..mm.))

# Length Frequency Histograms
# 2017






#2018
BlueLF.2018<-ggplot(blue.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,700))+
                labs(title = "Blue Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency (Note larger y-axis)") +
                theme_classic()
BlueLF.2018


```

## Center Lake
```{r}
# Subsetting species
#                             Carp
center.coc<-subset(Center, Species == "COC" & !is.na(Length..mm.))
summary(center.coc$Length..mm.)
#                             Buffalo
center.bib<-subset(Center, Species == "BIB" & !is.na(Length..mm.))
summary(center.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
centerLF.c.2018<-ggplot(center.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "Center Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
centerLF.c.2018

centerLF.b.2018<-ggplot(center.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "Center Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
centerLF.b.2018

```

## Five Island Lake
```{r}
# Subsetting species
#                             Carp
FiveIsland.coc<-subset(FiveIsland, Species == "COC" & !is.na(Length..mm.))
summary(FiveIsland.coc$Length..mm.)
#                             Buffalo
FiveIsland.bib<-subset(FiveIsland, Species == "BIB" & !is.na(Length..mm.))
summary(FiveIsland.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
FiveIslandLF.c.2018<-ggplot(FiveIsland.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "Five Island Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
FiveIslandLF.c.2018

FiveIslandLF.b.2018<-ggplot(FiveIsland.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "Five Island Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
FiveIslandLF.b.2018

```

## North Twin Lake
```{r}
# Subsetting species
#                             Carp
NTwin.coc<-subset(NTwin, Species == "COC" & !is.na(Length..mm.))
summary(NTwin.coc$Length..mm.)
#                             Buffalo
NTwin.bib<-subset(NTwin, Species == "BIB" & !is.na(Length..mm.))
summary(NTwin.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
NTwinLF.c.2018<-ggplot(NTwin.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "North Twin Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
NTwinLF.c.2018

NTwinLF.b.2018<-ggplot(NTwin.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "North Twin Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
NTwinLF.b.2018

```

## Silver Lake
```{r}
# Subsetting species
#                             Carp
Silver.coc<-subset(Silver, Species == "COC" & !is.na(Length..mm.))
summary(Silver.coc$Length..mm.)
#                             Buffalo
Silver.bib<-subset(Silver, Species == "BIB" & !is.na(Length..mm.))
summary(Silver.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
SilverLF.c.2018<-ggplot(Silver.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "Silver Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
SilverLF.c.2018

SilverLF.b.2018<-ggplot(Silver.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "Silver Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
SilverLF.b.2018

```

## South Twin Lake
```{r}
# Subsetting species
#                             Carp
STwin.coc<-subset(STwin, Species == "COC" & !is.na(Length..mm.))
summary(STwin.coc$Length..mm.)
#                             Buffalo
STwin.bib<-subset(STwin, Species == "BIB" & !is.na(Length..mm.))
summary(STwin.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
STwinLF.c.2018<-ggplot(STwin.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "South Twin Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
STwinLF.c.2018

STwinLF.b.2018<-ggplot(STwin.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "South Twin Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
STwinLF.b.2018

```

## Storm Lake
```{r}
# Subsetting species
#                             Carp
Storm.coc<-subset(Storm, Species == "COC" & !is.na(Length..mm.))
summary(Storm.coc$Length..mm.)
#                             Buffalo
Storm.bib<-subset(Storm, Species == "BIB" & !is.na(Length..mm.))
summary(Storm.bib$Length..mm.)

# Length Frequency Histograms
# 2017






#2018
StormLF.c.2018<-ggplot(Storm.coc, aes(x=Length..mm., fill=Species))+
                geom_histogram(binwidth=10,colour="black",position="dodge", fill = "burlywood4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(title = "Storm Lake Carp Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                theme_classic()
StormLF.c.2018

StormLF.b.2018<-ggplot(Storm.bib, aes(x=Length..mm., fill = Species))+
                  geom_histogram(binwidth = 10, colour = "black", position = "dodge", fill = "cornflowerblue")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(title = "Storm Lake Buffalo Length-Frequency",
                      x = "Length (mm)",
                      y = "Frequency") +
                  theme_classic()
StormLF.b.2018

```

# now arrange the objects in a meaningful way

```{r}
# Carp
grid.arrange(centerLF.c.2018, centerLF.b.2018, 
             FiveIslandLF.c.2018, FiveIslandLF.b.2018,
             NTwinLF.c.2018, NTwinLF.b.2018,
             SilverLF.c.2018, SilverLF.b.2018,
             STwinLF.c.2018, STwinLF.b.2018,
             StormLF.c.2018, StormLF.b.2018,
             BlueLF.2018,
             nrow=7,ncol=2)
```




# these might be the same as from annual report. better to keep anyway since hard drive space is free!


- separate plots below:
## Blue Lake
```{r}
# No extreme need to split species (one BIB in dataset) but I'll do it anyway
blue.coc<-subset(Blue, Species == "COC" & !is.na(Length..mm.))
blue.coc<-blue.coc[,-c(1:2)]
blue.coc<-melt(blue.coc)

# Length Frequency Histograms
# 2017






#2018
BlueLF.2018<-ggplot(blue.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack",
                               show.legend=F,
                               fill = "#00BFC4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,700))+
                labs(x = "",
                      y = "(Note larger y-axis)") +
                theme_classic()+
                theme(plot.margin = unit(c(0.2,0.2,0.2,0.2),"pt"))+
                annotate("text",x=500,y=600,label="Blue Lake Carp")
BlueLF.2018


```

## Center Lake
```{r}
# Subsetting species
#                             Carp
center.coc<-subset(Center, Species == "COC" & !is.na(Length..mm.))
center.coc<-center.coc[,-c(1:2)]
center.c.m<-melt(center.coc)

#                             Buffalo
center.bib<-subset(Center, Species == "BIB" & !is.na(Length..mm.))
center.bib<-center.bib[,-c(1:2)]
center.b.m<-melt(center.bib)

# Length Frequency Histograms
# 2017





#2018 Carp
centerLF.c.2018<-ggplot(center.c.m, aes(x=value,  fill=Gear))+
                geom_histogram(binwidth=10, 
                               colour="black", 
                               position = "stack",
                               show.legend = F) +
                coord_cartesian(xlim = c(100,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(x="",y="")+
                theme_classic()+
                theme(axis.text.x=element_blank(),
                      plot.margin = unit(c(0.2,0.2,0.2,0.2),"pt"))+
                annotate("text",x=250, y = 150, label = "Center Lake Carp")
centerLF.c.2018
#2018 Buffalo
centerLF.b.2018<-ggplot(center.b.m, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend = F)+
                  coord_cartesian(xlim = c(100,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",y = "") +
                  theme_classic()+
                  annotate("text", x= 250, y = 150, label = "Center Lake Buffalo")
centerLF.b.2018

```

## Five Island Lake
```{r}
# Subsetting species
#                             Carp
FiveIsland.coc<-subset(FiveIsland, Species == "COC" & !is.na(Length..mm.))
FiveIsland.coc<-FiveIsland.coc[,-c(1:2)]
FiveIsland.coc<-melt(FiveIsland.coc)
#                             Buffalo
FiveIsland.bib<-subset(FiveIsland, Species == "BIB" & !is.na(Length..mm.))
FiveIsland.bib<-FiveIsland.bib[,-c(1:2)]
FiveIsland.bib<-melt(FiveIsland.bib)

# Length Frequency Histograms
# 2017






#2018 Carp
FiveIslandLF.c.2018<-ggplot(FiveIsland.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack", 
                               show.legend=F)+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(x = "",
                     y = "") +
                theme_classic()+
                annotate("text", x= 250, y = 150, label = "Five Island Lake Carp Length-Frequency")
FiveIslandLF.c.2018
#2018 Buffalo 
FiveIslandLF.b.2018<-ggplot(FiveIsland.bib, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend=F)+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",y = "") +
                  theme_classic()+
                  annotate("text",x=250,y=150,label="Five Island Lake Buffalo Length-Frequency")
FiveIslandLF.b.2018

```

## North Twin Lake
```{r}
# Subsetting species
#                             Carp
NTwin.coc<-subset(NTwin, Species == "COC" & !is.na(Length..mm.))
NTwin.coc<-NTwin.coc[,-c(1:2)]
NTwin.coc<-melt(NTwin.coc)
#                             Buffalo
NTwin.bib<-subset(NTwin, Species == "BIB" & !is.na(Length..mm.))
NTwin.bib<-NTwin.bib[,-c(1:2)]
NTwin.bib<-melt(NTwin.bib)

# Length Frequency Histograms
# 2017






#2018
NTwinLF.c.2018<-ggplot(NTwin.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack", 
                               show.legend=F)+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(x = "",y = "") +
                theme_classic()+
                annotate("text",x=250,y=150, label = "North Twin Lake Carp Length-Frequency")
NTwinLF.c.2018

NTwinLF.b.2018<-ggplot(NTwin.bib, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend = F)+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",
                      y = "") +
                  theme_classic()+
                  annotate("text",x=250,y=150,label="North Twin Lake Buffalo Length-Frequency")
NTwinLF.b.2018

```

## Silver Lake
```{r}
# Subsetting species
#                             Carp
Silver.coc<-subset(Silver, Species == "COC" & !is.na(Length..mm.))
Silver.coc<-Silver.coc[,-c(1:2)]
Silver.coc<-melt(Silver.coc)
#                             Buffalo
Silver.bib<-subset(Silver, Species == "BIB" & !is.na(Length..mm.))
Silver.bib<-Silver.bib[,-c(1:2)]
Silver.bib<-melt(Silver.bib)

# Length Frequency Histograms
# 2017






#2018
SilverLF.c.2018<-ggplot(Silver.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack", 
                               show.legend=F,
                               fill = "#00BFC4")+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(x = "",y = "") +
                theme_classic()+
                annotate("text",x=250,y=150,label = "Silver Lake Carp Length-Frequency")
SilverLF.c.2018

SilverLF.b.2018<-ggplot(Silver.bib, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend = F,
                                 fill = "#00BFC4")+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",y = "") +
                  theme_classic()+
                  annotate("text",x=250,y=150,label = "Silver Lake Buffalo Length-Frequency")
SilverLF.b.2018

```

## South Twin Lake
```{r}
# Subsetting species
#                             Carp
STwin.coc<-subset(STwin, Species == "COC" & !is.na(Length..mm.))
STwin.coc<-STwin.coc[,-c(1:2)]
STwin.coc<-melt(STwin.coc)
#                             Buffalo
STwin.bib<-subset(STwin, Species == "BIB" & !is.na(Length..mm.))
STwin.bib<-STwin.bib[,-c(1:2)]
STwin.bib<-melt(STwin.bib)

# Length Frequency Histograms
# 2017






#2018
STwinLF.c.2018<-ggplot(STwin.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack", 
                               show.legend = F)+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs(x = "",y = "") +
                theme_classic()+
                annotate("text",x=250, y=190,label = "South Twin Lake Carp Length-Frequency")
STwinLF.c.2018

STwinLF.b.2018<-ggplot(STwin.bib, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend=F)+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",y = "") +
                  theme_classic()+
                  annotate("text",x=250,y=190,label = "South Twin Lake Buffalo Length-Frequency")
STwinLF.b.2018

```

## Storm Lake
```{r}
# Subsetting species
#                             Carp
Storm.coc<-subset(Storm, Species == "COC" & !is.na(Length..mm.))
Storm.coc<-Storm.coc[,-c(1:2)]
Storm.coc<-melt(Storm.coc)
#                             Buffalo
Storm.bib<-subset(Storm, Species == "BIB" & !is.na(Length..mm.))
Storm.bib<-Storm.bib[,-c(1:2)]
Storm.bib<-melt(Storm.bib)

# Length Frequency Histograms
# 2017






#2018
StormLF.c.2018<-ggplot(Storm.coc, aes(x=value, fill=Gear))+
                geom_histogram(binwidth=10,
                               colour="black",
                               position="stack", 
                               show.legend=F)+
                coord_cartesian(xlim = c(0,1000))+
                scale_y_continuous(limits = c(0,200))+
                labs( x = "",y = "") +
                theme_classic()+
                annotate("text",x=250,y=150,label="Storm Lake Carp Length-Frequency")
StormLF.c.2018

StormLF.b.2018<-ggplot(Storm.bib, aes(x=value, fill = Gear))+
                  geom_histogram(binwidth = 10, 
                                 colour = "black", 
                                 position = "stack", 
                                 show.legend=F)+
                  coord_cartesian(xlim = c(0,1000))+
                  scale_y_continuous(limits = c(0,200))+
                  labs(x = "",y = "") +
                  theme_classic()+
                  annotate("text",x=250,y=150,label="Storm Lake Buffalo Length-Frequency")
StormLF.b.2018

```

## now arrange the objects in a meaningful way
```{r}
# Carp
grid.arrange(centerLF.c.2018,FiveIslandLF.c.2018,NTwinLF.c.2018,SilverLF.c.2018,STwinLF.c.2018,StormLF.c.2018,BlueLF.2018,
             centerLF.b.2018,FiveIslandLF.b.2018,NTwinLF.b.2018,SilverLF.b.2018,STwinLF.b.2018,StormLF.b.2018,
             nrow=2,ncol=7)
```

