---
title: "Chapter 1"
author: "Marty Simonson"
date: "2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(tidyr)
#packages(rcompanion)
#packages(GISTools)
#packages(maps)
#packages(mapproj)
```

# Objective:

Chapter 1 of my dissertation proposal deals with variation in CPUE, primarily as a function of true abundance/density. However, there are many covariates (environmental, temporal, human) that can affect catchability of carp and buffalo. 

1) Is there a linear or nonlinear relationship between carp/buffalo density and CPUE?

2) What covariates influence CPUE? In what conditions does sampling result in tight correlation with true abundance/density?

## Sub-Objective 1

For this objective, I need to take the data set up and split it into each year (closure assumption)
- remove commercial seine sampling
- remove run-n-gun sampling (standard transects only)
- keep month, lake, species as ID variables

Then, I will have to use the following equations for the initial comparisons:

\[
\frac{C}{f} = \alpha * (\frac{N}{a})^{\beta+1} * exp^\epsilon 
\]

where $\alpha$ is an estimate of catchability at low population density, $\beta$ is the degree of curvature between cpue and density, anhd $\epsilon$ is the error. These parameters can be estimated from the natural log-transformed model:

\[
ln(\frac{C}{f}) = b_0 + (b_1 * ln\frac{N}{a}) + \epsilon 
\]

where the intercept $b_0$ equals $ln(\alpha)$ and the slope $b_1$ equals $\beta + 1$. 

For reference, when $b_1$ = 1, catch rate is linearly related to density (proportional, $\beta$ = 0). When $b_1$ is less than than 1, density increases faster than catch rate (hyperstability), and when $b_1$ is greater than 1, catch rate increases faster than actual density (hyperdepletion)

Here goes nothing:

## Data Entry and cleaning:

note: we are going to start from the beginning with the master (compiled) data file. No searching other spreadsheets for biomass density data or schnabel estimates typed in from google, this is transforming raw data into the variables needed for analysis. Many threads may extend from compiled data file depending on desired outcome (average weight/biomass density, for example)

### CPUE Data First
_CPUE data needs to be checked, some of that code will come from manual entry in another script_

```{r}
# overall chapter data name will be "catchability" 
# subsets and manipulations of data with additional prefixes and suffixes as necessary (year, species, variable)

# read in compiled data:
catchability <- read.csv("2018 and 2017 Compiled Data.csv")
str(catchability)


# need to change date format to yyyy-mm-dd
catchability$Date<-as.Date(catchability$Date, format = "%m/%d/%Y")


# Subset for only electrofishing
levels(catchability$Gear)
catchability<-droplevels(subset(catchability, Gear != "Commercial Seine" & Gear != "Fyke"))
levels(catchability$Gear)

# Subset for ONLY standard transect sampling
levels(catchability$Start.Latitude)

catchability<-droplevels(subset(catchability, Start.Latitude == "5 island std 1" |
                                              Start.Latitude == "5 Island Std 1" |
                                              Start.Latitude == "5 island std 2" |
                                              Start.Latitude == "5 island std 3" |
                                              Start.Latitude == "5 island std 4" |
                                              Start.Latitude == "5 island std 5" |
                                  
                                              Start.Latitude == "Blue Std 1" |
                                              Start.Latitude == "Blue Std 2" |
                                              Start.Latitude == "Blue Std 3" |
                                              Start.Latitude == "Blue Std 4" |
                                    
                                              Start.Latitude == "Center std 1" |
                                              Start.Latitude == "Center std 2" |
                                              Start.Latitude == "Center std 3" |
                                              Start.Latitude == "Center std 4" |
                                              Start.Latitude == "Center std 5" |
                                              Start.Latitude == "Center trans. 1" |
                                              Start.Latitude == "Center Trans. 1" |
                                              Start.Latitude == "Center trans. 2" |
                                              Start.Latitude == "Center Trans. 2" |
                                              Start.Latitude == "Center trans. 3" |
                                              Start.Latitude == "Center Trans. 3" |
                                              Start.Latitude == "Center trans. 4" |
                                              Start.Latitude == "Center Trans. 4" |
                                              Start.Latitude == "Center trans. 5" |
                                              Start.Latitude == "Center Trans. 5" |
                                  
                                              Start.Latitude == "EF1" |
                                              Start.Latitude == "EF2" |
                                              Start.Latitude == "EF3" |
                                              Start.Latitude == "EF4" |
                                              Start.Latitude == "EF5" |
                                  
                                              Start.Latitude == "Five Island trans. 1" |
                                              Start.Latitude == "Five Island trans. 2" |
                                              Start.Latitude == "Five Island trans. 3" |
                                              Start.Latitude == "Five Island trans. 4" |
                                              Start.Latitude == "Five Island trans. 5" |
                                              
                                              Start.Latitude == "N twin std 1" |
                                              Start.Latitude == "N twin std 2" |
                                              Start.Latitude == "N twin std 3" |
                                              Start.Latitude == "N twin std 4 (labeled as 5 on trans coordinates)" |
                                            
                                              Start.Latitude == "S Twin Std 2" |
                                              Start.Latitude == "S Twin Std 4" |
                                              Start.Latitude == "S. Twin Std 1"|
                                  
                                              Start.Latitude == "Silver std 1" |
                                              Start.Latitude == "Silver Std 1" |
                                              Start.Latitude == "Silver std 2" |
                                              Start.Latitude == "Silver Std 2" |
                                              Start.Latitude == "Silver std 3" |
                                              Start.Latitude == "Silver Std 3" |
                                              Start.Latitude == "Silver std 4" |
                                              Start.Latitude == "Silver Std 4" |
                                              Start.Latitude == "Silver std 5" |
                                              Start.Latitude == "Silver std 6" |
                                  
                                              Start.Latitude == "Silver trans. 1" |
                                              Start.Latitude == "Silver Trans. 1" |
                                              Start.Latitude == "Silver trans. 2" |
                                              Start.Latitude == "Silver Trans. 2" |
                                              Start.Latitude == "Silver trans. 3" |
                                              Start.Latitude == "Silver Trans. 3" |
                                              Start.Latitude == "Silver trans. 4" |
                                              Start.Latitude == "Silver Trans. 4" |
                                              Start.Latitude == "Silver trans. 5" |
                                              Start.Latitude == "Silver Trans. 5" |
                                              Start.Latitude == "Silver trans. 6" |
                                              Start.Latitude == "Silver Trans. 6" |
                                  
                                              Start.Latitude == "Std Trans 3" |
                                              Start.Latitude == "Std. Trans 1" |
                                              Start.Latitude == "Std. Trans 2" |
                                              Start.Latitude == "Std. Trans 3" |
                                              Start.Latitude == "Std. Trans 4" |
                                              Start.Latitude == "Std. Trans 5" |
                                              Start.Latitude == "Std. Trans 6" |
                                              Start.Latitude == "Std. Trans 7" |
                                              Start.Latitude == "Std. Trans 9" |
                                  
                                              Start.Latitude == "Storm std 1" |
                                              Start.Latitude == "Storm std 2" |
                                              Start.Latitude == "Storm std 4 (by casino beach)" |
                                              Start.Latitude == "Storm std 5" |
                                              Start.Latitude == "Storm std 7" |
                                              Start.Latitude == "Storm std 9" |
                                  
                                              Start.Latitude == "Storm trans. 1" |
                                              Start.Latitude == "Storm trans. 2" |
                                              Start.Latitude == "Storm trans. 4" |
                                              Start.Latitude == "Storm trans. 5" |
                                              Start.Latitude == "Storm trans. 7" |
                                              Start.Latitude == "Storm trans. 9"))

# Species maybe
levels(catchability$Species)
catchability<-droplevels(subset(catchability, Species == "COC" | Species == "BIB"))
levels(catchability$Species)

# Drop unused columns (weather, etc)
str(catchability)
catchability<-catchability[,-c(7:12,14:16,18:20)]
str(catchability)

# For consistency in date-lake-run later, must consolidate transect names. 
#
#       Using # to denote standard electrofishing runs, 
#         for example the second standard run for a lake would be EF2
#
levels(catchability$Start.Latitude)

catchability$Start.Latitude<-as.character(catchability$Start.Latitude, stringsAsFactors = F)

catchability$Start.Latitude[catchability$Start.Latitude == "5 island std 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "5 Island Std 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "5 island std 2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "5 island std 3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "5 island std 4"] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "5 island std 5"] <- "5"
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Blue Std 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Blue Std 2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Blue Std 3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Blue Std 4"] <- "4"
                                    
catchability$Start.Latitude[catchability$Start.Latitude == "Center std 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Center std 2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Center std 3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Center std 4"] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Center std 5"] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Center trans. 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Center Trans. 1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Center trans. 2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Center Trans. 2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Center trans. 3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Center Trans. 3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Center trans. 4"] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Center Trans. 4"] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Center trans. 5"] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Center Trans. 5"] <- "5"
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Five Island trans. 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Five Island trans. 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Five Island trans. 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Five Island trans. 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Five Island trans. 5" ] <- "5"
                                              
catchability$Start.Latitude[catchability$Start.Latitude == "N twin std 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "N twin std 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "N twin std 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "N twin std 4 (labeled as 5 on trans coordinates)" ] <- "4"
                                            
catchability$Start.Latitude[catchability$Start.Latitude == "S Twin Std 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "S Twin Std 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "S. Twin Std 1"] <- "1"

catchability$Start.Latitude[catchability$Start.Latitude == "EF1"] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "EF2"] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "EF3"] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "EF4"] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "EF5"] <- "5"
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Std 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Std 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Std 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Std 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver std 6" ] <- "6" 
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Silver trans. 6" ] <- "6" 
catchability$Start.Latitude[catchability$Start.Latitude == "Silver Trans. 6" ] <- "6" 
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Std Trans 3" ] <- "3" 
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 1" ] <- "1" 
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 3" ] <- "3"
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 6" ] <- "6" 
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 7" ] <- "7" 
catchability$Start.Latitude[catchability$Start.Latitude == "Std. Trans 9" ] <- "9" 
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 4 (by casino beach)" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 7" ] <- "7" 
catchability$Start.Latitude[catchability$Start.Latitude == "Storm std 9" ] <- "9" 
                                  
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 1" ] <- "1"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 2" ] <- "2"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 4" ] <- "4"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 5" ] <- "5"
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 7" ] <- "7" 
catchability$Start.Latitude[catchability$Start.Latitude == "Storm trans. 9" ] <- "9" 

catchability$Start.Latitude<-as.factor(catchability$Start.Latitude)
levels(catchability$Start.Latitude) # No EF8, as it should be.

#
#
# Make sure that the each lake has the appropriate transects
levels(catchability$Lake)

summary(catchability$Start.Latitude[catchability$Lake == "5 Island"])
# all five transects represented. Low number of captures, though.

summary(catchability$Start.Latitude[catchability$Lake == "Blue"])
# all four transects represented. probably highest catches

summary(catchability$Start.Latitude[catchability$Lake == "Center"])
# all 5 transects represented, good catches

summary(catchability$Start.Latitude[catchability$Lake == "N. Twin"]) 
# all 4 transects represented. 

summary(catchability$Start.Latitude[catchability$Lake == "S. Twin"])
# This is more appropriate catches. all four transects represented.

summary(catchability$Start.Latitude[catchability$Lake == "Silver"])
# All six transects represented. more or less appropriate catches.

summary(catchability$Start.Latitude[catchability$Lake == "Storm"])
# correct transects represented. better catches than I was expecting.

#
#
#
#     Pasting Date/Lake/Run into single ID column, our minimum sample unit
catchability$DLR<-factor(paste(catchability$Date,
                               catchability$Lake,
                               catchability$Start.Latitude,
                               sep = ";"))
str(catchability$DLR) # 352 sample units across 7 lakes in 2 years

summary(catchability)
```

##### splitting by species and CPUE (additional 0 catch transects need to be incorporated to the averages)
to 1) comply with format used in CPUE Plots.Rmd code, and 2) do data checking against each other to make sure there's not samples for one species that aren't included for the other (looking at you, buffalo)

```{r}
# Using modified 'catchability' data frame from previous chunk

# Adding effort (in hours instead of seconds)
catchability$Effort<-catchability$On.Time..s./3600

# Splitting to carp and buffalo
coc.df<-subset(catchability, Species == "COC")
bib.df<-subset(catchability, Species == "BIB")

# Common Carp CPUE
coc.cpue<-aggregate(Effort~Lake+DLR, data = coc.df, mean) # df of three columns: Lake, DLR, and Effort(hours) for each transect
coc.df$dummy<-1 # each row is a single fish observation of carp
coc.n<-aggregate(dummy~Lake+DLR, data = coc.df, sum) #df like the CPUE but with sum of catch for each transect
coc.cpue$n<-coc.n$dummy # add counts to cpue df
coc.cpue$cpue<-coc.cpue$n/coc.cpue$Effort # this is c/f, CPUE for each run where at least one carp was observed

# Bigmouth Buffalo CPUE
bib.cpue<-aggregate(Effort~Lake+DLR, data = bib.df, mean) # df of three columns: Lake, DLR, and Effort for each transect
bib.df$dummy<-1 # each row is a single fish observation of carp
bib.n<-aggregate(dummy~Lake+DLR, data = bib.df, sum) #df like the CPUE but with sum of catch for each transect
bib.cpue$n<-bib.n$dummy # add counts to cpue df
bib.cpue$cpue<-bib.cpue$n/bib.cpue$Effort # this is c/f, CPUE for each run where at least one carp was observed

```

##### data checking CPUE
```{r}
###############################################
#
#         data checking
#
#         Need to confirm EACH date/lake/run
#         and add in 0 values
BlueCheck<-subset(coc.cpue, Lake == "Blue")
levels(droplevels(BlueCheck$DLR))

FICheck<-subset(coc.cpue, Lake == "5 Island")
levels(droplevels(FICheck$DLR))

CenterCheck<-subset(coc.cpue, Lake == "Center")
levels(droplevels(CenterCheck$DLR))

NTCheck<-subset(coc.cpue, Lake == "N. Twin")
levels(droplevels(NTCheck$DLR))

SilverCheck<-subset(coc.cpue, Lake == "Silver")
levels(droplevels(SilverCheck$DLR))

STCheck<-subset(coc.cpue, Lake == "S. Twin")
levels(droplevels(STCheck$DLR))

StormCheck<-subset(coc.cpue, Lake == "Storm")
levels(droplevels(StormCheck$DLR))

################################################
BlueCheck<-subset(bib.cpue, Lake == "Blue")
levels(droplevels(BlueCheck$DLR))

FICheck<-subset(bib.cpue, Lake == "5 Island")
levels(droplevels(FICheck$DLR))

CenterCheck<-subset(bib.cpue, Lake == "Center")
levels(droplevels(CenterCheck$DLR))

NTCheck<-subset(bib.cpue, Lake == "N. Twin")
levels(droplevels(NTCheck$DLR))

SilverCheck<-subset(bib.cpue, Lake == "Silver")
levels(droplevels(SilverCheck$DLR))

STCheck<-subset(bib.cpue, Lake == "S. Twin")
levels(droplevels(STCheck$DLR))

StormCheck<-subset(bib.cpue, Lake == "Storm")
levels(droplevels(StormCheck$DLR))

##########################################################################################
#
#           Blue Lake is good. Not worrying about the 1 buffalo
#
#           Five Island Lake additions:
#                         At least 1 carp per transect somehow
#                         44 runs without a buffalo though. 44/78 runs.
#                         Deleting runs from other file that weren't standard transects.


a<-c("5 Island","2018-05-23;5 Island;2", 928/3600, 0, 0)
b<-c("5 Island","2018-05-23;5 Island;3", 1298/3600, 0, 0)
c<-c("5 Island","2018-05-23;5 Island;4", 904/3600, 0, 0)
d<-c("5 Island","2018-05-23;5 Island;5", 905/3600, 0, 0)

i<-c("5 Island","2018-06-06;5 Island;3", 1021/3600, 0, 0)
j<-c("5 Island","2018-06-06;5 Island;5", 1051/3600, 0, 0)

o<-c("5 Island","2018-07-12;5 Island;1", 963/3600,0,0)
p<-c("5 Island","2018-07-12;5 Island;3", 909/3600,0,0)
q<-c("5 Island","2018-07-12;5 Island;4", 910/3600,0,0)
r<-c("5 Island","2018-07-12;5 Island;5", 969/3600,0,0)

x<-c("5 Island","2018-07-20;5 Island;2", 1161/3600,0,0)
y<-c("5 Island","2018-07-20;5 Island;3", 1142/3600,0,0)
z<-c("5 Island","2018-07-20;5 Island;4", 1389/3600,0,0)
aa<-c("5 Island","2018-07-20;5 Island;5", 1306/3600,0,0)

ee<-c("5 Island","2018-08-01;5 Island;2", 1025/3600,0,0)
ff<-c("5 Island","2018-08-01;5 Island;3", 1291/3600,0,0)
gg<-c("5 Island","2018-08-01;5 Island;4", 1162/3600,0,0)

mm<-c("5 Island","2018-09-09;5 Island;3", 901/3600,0,0)
nn<-c("5 Island","2018-09-09;5 Island;4", 1060/3600,0,0)
oo<-c("5 Island","2018-09-09;5 Island;5", 1021/3600,0,0)

pp<-c("5 Island","2018-10-06;5 Island;4", 1053/3600,0,0)
qq<-c("5 Island","2018-10-06;5 Island;5", 939/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,c,d,i,j,o,p,q,r,x,y,z,
                           aa,ee,ff,gg,mm,nn,oo,pp,qq)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)
#####################        AWESOME        ######################

##################################
#
# Center Lake
#       25 added to bib
#       13 added to coc
#                       both reduced to std runs only
#
a<-c("Center","2018-05-25;Center;2",931/3600,0,0)
b<-c("Center","2018-05-25;Center;3",900/3600,0,0)
c<-c("Center","2018-05-25;Center;4",909/3600,0,0)
d<-c("Center","2018-05-25;Center;5",939/3600,0,0)

g<-c("Center","2018-06-04;Center;2",1041/3600,0,0)
h<-c("Center","2018-06-04;Center;3",893/3600,0,0)
i<-c("Center","2018-06-04;Center;4",659/3600,0,0)
j<-c("Center","2018-06-04;Center;5",913/3600,0,0)

k<-c("Center","2018-06-11;Center;2",1010/3600,0,0)
l<-c("Center","2018-06-11;Center;3",672/3600,0,0)
m<-c("Center","2018-06-11;Center;4",599/3600,0,0)
n<-c("Center","2018-06-11;Center;5",762/3600,0,0)

p<-c("Center","2018-06-19;Center;4",645/3600,0,0)
q<-c("Center","2018-06-19;Center;5",896/3600,0,0)

r<-c("Center","2018-07-19;Center;4",770/3600,0,0)
s<-c("Center","2018-07-19;Center;5",971/3600,0,0)

v<-c("Center","2018-08-02;Center;1",1003/3600,0,0)
w<-c("Center","2018-08-02;Center;3",872/3600,0,0)
x<-c("Center","2018-08-02;Center;4",870/3600,0,0)
y<-c("Center","2018-08-02;Center;5",919/3600,0,0)


add.on<-data.frame(t(data.frame(a,b,c,d,g,h,i,j,k,l,m,n,p,q,r,s,v,w,x,y)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)
# Now Carp!
a<-c("Center","2018-05-17;Center;2",917/3600,0,0)
b<-c("Center","2018-05-17;Center;3",893/3600,0,0)
c<-c("Center","2018-05-17;Center;5",932/3600,0,0)

d<-c("Center","2018-05-25;Center;1",925/3600,0,0)
e<-c("Center","2018-05-25;Center;2",931/3600,0,0)

g<-c("Center","2018-07-19;Center;2",848/3600,0,0)

h<-c("Center","2018-08-02;Center;1",1003/3600,0,0)
i<-c("Center","2018-08-02;Center;4",870/3600,0,0)

j<-c("Center","2018-09-08;Center;2",885/3600,0,0)
k<-c("Center","2018-09-08;Center;4",778/3600,0,0)

l<-c("Center","2018-10-05;Center;2",849/3600,0,0)
m<-c("Center","2018-10-05;Center;4",549/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,c,d,e,g,h,i,j,k,l,m)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
coc.cpue<-rbind(coc.cpue,add.on)
####################################      AWESOME     ##########################



##############
# North Twin           
#
#       42 buffalo
#       16 carp
#               reduced to just standard runs
#
a<-c("N. Twin","2018-05-15;N. Twin;1",900/3600,0,0)
b<-c("N. Twin","2018-05-15;N. Twin;2",900/3600,0,0)

c<-c("N. Twin","2018-05-18;N. Twin;3",900/3600,0,0)

d<-c("N. Twin","2018-06-06;N. Twin;2",900/3600,0,0)

e<-c("N. Twin","2018-07-03;N. Twin;1",900/3600,0,0)
f<-c("N. Twin","2018-07-03;N. Twin;2",900/3600,0,0)
g<-c("N. Twin","2018-07-03;N. Twin;3",900/3600,0,0)

h<-c("N. Twin","2018-07-09;N. Twin;1",900/3600,0,0)
i<-c("N. Twin","2018-07-09;N. Twin;3",900/3600,0,0)

k<-c("N. Twin","2018-08-01;N. Twin;2",900/3600,0,0)
l<-c("N. Twin","2018-08-01;N. Twin;3",900/3600,0,0)

m<-c("N. Twin","2018-08-06;N. Twin;1",900/3600,0,0)
n<-c("N. Twin","2018-08-06;N. Twin;2",900/3600,0,0)
o<-c("N. Twin","2018-08-06;N. Twin;3",900/3600,0,0)

p<-c("N. Twin","2018-11-01;N. Twin;1",905/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,c,d,e,f,g,h,i,k,l,m,n,o,p)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)

# COC now
a<-c("N. Twin","2018-05-18;N. Twin;1",900/3600,0,0)

b<-c("N. Twin","2018-07-09;N. Twin;1",900/3600,0,0)

e<-c("N. Twin","2018-08-06;N. Twin;5",900/3600,0,0)

f<-c("N. Twin","2018-10-10;N. Twin;3",900/3600,0,0)

g<-c("N. Twin","2018-10-12;N. Twin;1",900/3600,0,0)
h<-c("N. Twin","2018-10-12;N. Twin;2",900/3600,0,0)
i<-c("N. Twin","2018-10-12;N. Twin;5",900/3600,0,0)

j<-c("N. Twin","2018-11-01;N. Twin;2",849/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,e,f,g,h,i,k)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
coc.cpue<-rbind(coc.cpue,add.on)



###################################
# Silver Lake
#       50 bib additions
#       8 carp additions
#                         reduced to just standard runs
#

b<-c("Silver","2018-05-18;Silver;5",927/3600,0,0)
c<-c("Silver","2018-07-11;Silver;6",996/3600,0,0)
d<-c("Silver","2018-07-18;Silver;3",1052/3600,0,0)
e<-c("Silver","2018-07-18;Silver;5",899/3600,0,0)
f<-c("Silver","2018-08-02;Silver;1",1116/3600,0,0)
g<-c("Silver","2018-10-05;Silver;5",549/3600,0,0)
h<-c("Silver","2018-10-05;Silver;6",804/3600,0,0)

add.on<-data.frame(t(data.frame(b,c,d,e,f,g,h)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
coc.cpue<-rbind(coc.cpue,add.on)


# now 50 buffalo
a<-c("Silver","2018-05-24;Silver;6",894/3600,0,0)
c<-c("Silver","2018-06-04;Silver;1",901/3600,0,0)
d<-c("Silver","2018-06-04;Silver;2",914/3600,0,0)
m<-c("Silver","2018-07-11;Silver;3",710/3600,0,0)
n<-c("Silver","2018-07-11;Silver;5",1124/3600,0,0)
p<-c("Silver","2018-07-18;Silver;2",2568/3600,0,0)
q<-c("Silver","2018-07-18;Silver;5",899/3600,0,0)
x<-c("Silver","2018-08-02;Silver;1",1116/3600,0,0)
y<-c("Silver","2018-08-02;Silver;2",1307/3600,0,0)
z<-c("Silver","2018-08-02;Silver;3",1121/3600,0,0)

aa<-c("Silver","2018-08-02;Silver;4",854/3600,0,0)
bb<-c("Silver","2018-08-02;Silver;5",1202/3600,0,0)
cc<-c("Silver","2018-08-02;Silver;6",923/3600,0,0)
dd<-c("Silver","2018-08-06;Silver;1",997/3600,0,0)
ee<-c("Silver","2018-08-06;Silver;2",1173/3600,0,0)
ff<-c("Silver","2018-08-06;Silver;3",1004/3600,0,0)
gg<-c("Silver","2018-08-06;Silver;4",688/3600,0,0)
hh<-c("Silver","2018-08-06;Silver;5",952/3600,0,0)
ii<-c("Silver","2018-08-06;Silver;6",810/3600,0,0)
mm<-c("Silver","2018-09-08;Silver;1",1156/3600,0,0)
nn<-c("Silver","2018-09-08;Silver;2",1261/3600,0,0)
oo<-c("Silver","2018-09-08;Silver;3",1110/3600,0,0)
pp<-c("Silver","2018-09-08;Silver;5",1181/3600,0,0)
qq<-c("Silver","2018-09-08;Silver;6",961/3600,0,0)
ss<-c("Silver","2018-10-05;Silver;2",964/3600,0,0)
tt<-c("Silver","2018-10-05;Silver;3",1208/3600,0,0)
uu<-c("Silver","2018-10-05;Silver;4",890/3600,0,0)
vv<-c("Silver","2018-10-05;Silver;5",594/3600,0,0)
ww<-c("Silver","2018-10-05;Silver;6",804/3600,0,0)
xx<-c("Silver","2018-10-05;Silver;1",1132/3600,0,0)

add.on<-data.frame(t(data.frame(a,c,d,m,n,p,q,x,y,z,
                           aa,bb,cc,dd,ee,ff,gg,hh,ii,mm,nn,oo,pp,qq,ss,tt,uu,vv,ww,xx)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)

#################################
#
#         South Twin
# buff
a<-c("S. Twin","2018-05-21;S. Twin;2",900/3600,0,0)

b<-c("S. Twin","2018-06-05;S. Twin;1",900/3600,0,0)
c<-c("S. Twin","2018-06-05;S. Twin;3",900/3600,0,0)
d<-c("S. Twin","2018-06-05;S. Twin;4",900/3600,0,0)

e<-c("S. Twin","2018-06-18;S. Twin;1",900/3600,0,0)
f<-c("S. Twin","2018-06-18;S. Twin;2",900/3600,0,0)
g<-c("S. Twin","2018-06-18;S. Twin;3",900/3600,0,0)
h<-c("S. Twin","2018-06-18;S. Twin;4",900/3600,0,0)

l<-c("S. Twin","2018-07-05;S. Twin;1",900/3600,0,0)
m<-c("S. Twin","2018-07-05;S. Twin;3",900/3600,0,0)

q<-c("S. Twin","2018-07-13;S. Twin;1",900/3600,0,0)
r<-c("S. Twin","2018-07-13;S. Twin;2",900/3600,0,0)
s<-c("S. Twin","2018-07-13;S. Twin;3",900/3600,0,0)
t<-c("S. Twin","2018-07-13;S. Twin;4",900/3600,0,0)

u<-c("S. Twin","2018-08-31;S. Twin;2",900/3600,0,0)

v<-c("S. Twin","2018-09-04;S. Twin;1",900/3600,0,0)
w<-c("S. Twin","2018-09-04;S. Twin;4",900/3600,0,0)

x<-c("S. Twin","2018-08-03;S. Twin;1",900/3600,0,0)
y<-c("S. Twin","2018-08-03;S. Twin;2",900/3600,0,0)
z<-c("S. Twin","2018-08-03;S. Twin;3",900/3600,0,0)
aa<-c("S. Twin","2018-08-03;S. Twin;4",900/3600,0,0)

bb<-c("S. Twin","2018-08-07;S. Twin;1",900/3600,0,0)
cc<-c("S. Twin","2018-08-07;S. Twin;2",900/3600,0,0)
dd<-c("S. Twin","2018-08-07;S. Twin;3",900/3600,0,0)
ee<-c("S. Twin","2018-08-07;S. Twin;4",900/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,c,d,e,f,g,h,l,m,q,r,s,t,u,v,w,x,y,z,
                           aa,bb,cc,dd,ee)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)


# carp now
e<-c("S. Twin","2018-10-11;S. Twin;4",900/3600,0,0)

add.on<-data.frame(t(data.frame(e)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
coc.cpue<-rbind(coc.cpue,add.on)


##########################################################
#
#
#         STORM LAKE. THE FINAL FRONTIER
#
# 5 CARP

c<-c("Storm","2018-08-08;Storm;1", 902/3600,0,0)
d<-c("Storm","2018-08-08;Storm;5", 804/3600,0,0)
f<-c("Storm","2018-08-08;Storm;9", 804/3600,0,0)


add.on<-data.frame(t(data.frame(c,d,f)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
coc.cpue<-rbind(coc.cpue,add.on)

# 60+ buffalo
a<-c("Storm","2018-05-21;Storm;1", 910/3600,0,0)
b<-c("Storm","2018-05-21;Storm;2", 903/3600,0,0)
c<-c("Storm","2018-05-21;Storm;4", 928/3600,0,0)
d<-c("Storm","2018-05-21;Storm;5", 925/3600,0,0)
e<-c("Storm","2018-05-21;Storm;7", 901/3600,0,0)
f<-c("Storm","2018-05-21;Storm;9", 903/3600,0,0)

i<-c("Storm","2018-06-07;Storm;1", 1205/3600,0,0)
j<-c("Storm","2018-06-07;Storm;2", 955/3600,0,0)
k<-c("Storm","2018-06-07;Storm;4", 863/3600,0,0)


rrr<-c("Storm","2018-06-08;Storm;1", 901/3600,0,0)
l<-c("Storm","2018-06-08;Storm;2", 944/3600,0,0)
m<-c("Storm","2018-06-08;Storm;4", 914/3600,0,0)
n<-c("Storm","2018-06-08;Storm;5", 1579/3600,0,0)
o<-c("Storm","2018-06-08;Storm;9", 1107/3600,0,0)

s<-c("Storm","2018-06-27;Storm;1", 918/3600,0,0)
t<-c("Storm","2018-06-27;Storm;2", 1009/3600,0,0)
u<-c("Storm","2018-06-27;Storm;4", 1151/3600,0,0)
v<-c("Storm","2018-06-27;Storm;8", 998/3600,0,0)
w<-c("Storm","2018-06-27;Storm;9", 948/3600,0,0)

y<-c("Storm","2018-07-16;Storm;1", 948/3600,0,0)
z<-c("Storm","2018-07-16;Storm;2", 944/3600,0,0)
aa<-c("Storm","2018-07-16;Storm;4", 959/3600,0,0)
bb<-c("Storm","2018-07-16;Storm;5", 1089/3600,0,0)
cc<-c("Storm","2018-07-16;Storm;9", 891/3600,0,0)

xx<-c("Storm","2018-08-08;Storm;9", 804/3600,0,0)
yy<-c("Storm","2018-08-08;Storm;7", 999/3600,0,0)
zz<-c("Storm","2018-08-08;Storm;1", 804/3600,0,0)
aaa<-c("Storm","2018-08-08;Storm;2", 948/3600,0,0)
bbb<-c("Storm","2018-08-08;Storm;4", 889/3600,0,0)
ccc<-c("Storm","2018-08-08;Storm;5", 956/3600,0,0)

hhh<-c("Storm","2018-09-09;Storm;1", 903/3600,0,0)
iii<-c("Storm","2018-09-09;Storm;4", 866/3600,0,0)
jjj<-c("Storm","2018-09-09;Storm;5", 1040/3600,0,0)
kkk<-c("Storm","2018-09-09;Storm;9", 902/3600,0,0)

lll<-c("Storm","2018-10-05;Storm;1", 945/3600,0,0)
mmm<-c("Storm","2018-10-05;Storm;2", 828/3600,0,0)
nnn<-c("Storm","2018-10-05;Storm;4", 867/3600,0,0)
ooo<-c("Storm","2018-10-05;Storm;5", 1623/3600,0,0)
ppp<-c("Storm","2018-10-05;Storm;7", 1230/3600,0,0)
qqq<-c("Storm","2018-10-05;Storm;9", 944/3600,0,0)

add.on<-data.frame(t(data.frame(a,b,c,d,e,f,i,j,k,l,m,n,o,s,t,u,v,w,y,z,
                           aa,bb,cc,dd,xx,yy,zz,
                           aaa,bbb,ccc,hhh,iii,jjj,kkk,lll,mmm,nnn,ooo,ppp,qqq,rrr)),
                   stringsAsFactors = F)
names(add.on)<-c("Lake","DLR","Effort","n","cpue")
add.on$n<-as.numeric(add.on$n)
add.on$cpue<-as.numeric(add.on$cpue)
add.on$Effort<-as.numeric(add.on$Effort)
add.on$Lake<-as.factor(add.on$Lake)
add.on$DLR<-as.factor(add.on$DLR)
str(add.on)
bib.cpue<-rbind(bib.cpue,add.on)

```

```{r}
setdiff(coc.cpue$DLR,bib.cpue$DLR) # entries in coc not in bib (BLUE lake)
setdiff(bib.cpue$DLR,coc.cpue$DLR) # entries in bib not in coc 
```

