---
title: "CJS Plots"
author: "Martin A. Simonson"
date: "February 6, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Front Matter
```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
#packages(gridExtra)
packages(reshape2)
#packages(FSA)
#packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
#packages(lattice)
#packages(tikzDevice)
#packages(GGally)
#packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(RMark)
packages(tidyr)
```

plotting the top two models for CJS survival and recapture estimates from carp and buffalo in three iowa lakes in 2017 and 2018.

- Three lakes are grouped alphabetically, Center, N Twin, S Twin
- carp and buffalo are completely separate MARK .dbf files so they need to be merged here
- survival (Phi, S) is the first 21 parameters (7 for each lake)
- recap rate (p) is second 21 parameters. These are completely group and time dependent, meaning each season's recap rate in each lake is it's own parameter.

## The Plot:

side-by-side boxplot, or using lattice to split Open vs. Winter
- potential bargraph but those are whack sauce

survival (phi) on y axis
Open water (spring summer fall) vs. Winter
grouped within that is Center vs Twins 

duplicated for each species. Shown separately due to different top models


```{r,}
# Create DF with CPUE and 95% CI for each lake...

Lakes<-c("Center","N+S Twin")
Carp.S.open <- c(0.9364316,0.9364316)
Carp.S.open.L<-c(0.886910,0.8886910)
Carp.S.open.U<-c(0.9795353,0.9795353)

Carp.S.winter <- c(0.5137947,0.9364316)
Carp.S.winter.L<-c(0.0662681,0.8886910)
Carp.S.winter.U<-c(0.9951693,0.9795353)

Buff.S.open <- c(NA,81894,104513,26802)
Buff.S.open.L<-c(NA,59.40,17.99,36.56)
Buff.S.open.U<-c(NA,145.17,31.14,95.32)

Buff.S.winter <- c(NA,327.6,225.7,48.3)
Buff.S.winter.L<-c(63432,45202,5387,29180)
Buff.S.winter.u<-c(224140,77317,17058,92395)

Carp.DF<-data.frame(Lakes,Carp.S.open,Carp.S.open.L,Carp.S.open.U,
                          Carp.S.winter,Carp.S.winter.L,Carp.S.winter.U)

summary(Carp.DF)

```

Now to create 2 carp plots, 
- Carp actual top model (Phi(Center(.),N=S(.)) p (g*t))
```{r,}
Lakes<-c("Center","Twins")
Carp.S<-c(0.8821662, 0.9522349)
Carp.S.L<-c(0.7190165,0.9047440)
Carp.S.U<-c(0.9970485,0.9907159)
Carp.DF<-data.frame(Lakes,
                    Carp.S,
                    Carp.S.L,
                    Carp.S.U)

# Carp FIrst
carp1<-ggplot(Carp.DF, aes(x=Lakes, y=Carp.S)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = Carp.S.U, ymin = Carp.S.L)) +
  labs(title = "Common Carp Survival",
       x = "Lake",
       y = "Survival") +
  theme_bw() 
carp1


```

- carp model that matches buff top model (and within 2 AICc points)
- Phi(Center(winter), center(open)=N=S(.)) p(g*t)

```{r,}
Lakes<-c("Center","Twins","Center","Twins")
Season<-c("Open","Open","Winter","Winter")
Carp.S<-c(0.9444109,0.9444109,0.8517614,0.9444109)
Carp.S.L<-c(0.8979603,0.8979603,0.4493000,0.8979603)
Carp.S.U<-c(0.9866552,0.9866552,0.9979746,0.9866552)
Carp.DF<-data.frame(Lakes,
                    Season,
                    Carp.S,
                    Carp.S.L,
                    Carp.S.U)

# Carp FIrst
carp2<-ggplot(Carp.DF, aes(x=Lakes, y=Carp.S)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = Carp.S.U, ymin = Carp.S.L)) +
  labs(title = "Common Carp Survival",
       x = "Lake",
       y = "Survival") +
  facet_wrap(~Season)+
  theme_bw() 
carp2


```




- two Buff Plots:
top model:

Phi(Center, n=s(.); p (g*t))

```{r}
Lakes<-c("Center","Twins")
Carp.S<-c(0.6295946, 0.9510359)
Carp.S.L<-c(0.4252622,0.903211)
Carp.S.U<-c(0.8689247,0.9881958)
Carp.DF<-data.frame(Lakes,
                    Carp.S,
                    Carp.S.L,
                    Carp.S.U)

# Buffalo
Buff1<-ggplot(Carp.DF, aes(x=Lakes, y=Carp.S)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = Carp.S.U, ymin = Carp.S.L)) +
  labs(title = "Buffalo Survival",
       x = "Lake",
       y = "Survival") +
  theme_bw() 
Buff1

```






- second ranked model
phi(center(winter), center(open)=n=s=(.) and p(g*t))


```{r,}
Lakes<-c("Center","Twins","Center","Twins")
Season<-c("Open","Open","Winter","Winter")
Buff.S<-c(0.9358976,0.9358976, 0.5108930,0.9358976)
Buff.S.L<-c(0.8892595,0.8892595,0.0586412,0.8892595)
Buff.S.U<-c(0.9759545,0.9759545,0.9938855,0.9759545)
Buff.DF<-data.frame(Lakes,
                    Season,
                    Buff.S,
                    Buff.S.L,
                    Buff.S.U)

# Buffalo
Buff<-ggplot(Buff.DF, aes(x=Lakes, y=Buff.S)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = Buff.S.U, ymin = Buff.S.L)) +
  labs(title = "Buffalo Survival",
       x = "Lake",
       y = "Survival") +
  facet_wrap(~Season)+
  theme_bw() 
Buff


```