---
title: "Meerbeek Summary"
author: "Martin Simonson"
date: "March 4, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
#packages(GISTools)
#packages(maps)
#packages(mapproj)
```


# CPUE

comprehensive spring electrofishing and fall fyke netting surveys for all species.
Jon wants CPUE in n/hour and n/net. 

He wants species as rows. gear as broad column and lake as subcolumn.
values in each cell are to the tenth of a unit.

Also I guess it's only for my main 4 lakes? 
- Don't do any extra work, stupid. He will tell you if he needs more.

```{r}
# load and clean up data to what I need to work with.
df<-read.csv("2018_COC_BIB_CMR_Data.csv", header = T)
head(df)

# Year
df<-droplevels(subset(df, Year == "2018"))
df<-droplevels(subset(df, Lake == "Center" |
                          Lake == "Five Island" |
                          Lake == "Silver" |
                          Lake == "Storm"))
df.fyke<-droplevels(subset(df, Gear == "Fyke"))
df.shock<-droplevels(subset(df, Gear == "Electroshocking"))
df.shock<-droplevels(subset(df.shock, Site.Transect == "EF.Comp.1" |
                                      Site.Transect == "EF.Comp.2" |
                                      Site.Transect == "EF.Comp.3" |
                                      Site.Transect == "EF.Comp.4" |
                                      Site.Transect == "EF.Comp.5" |
                                      Site.Transect == "EF.Comp.6" |
                                      Site.Transect == "EF.Comp.7" |
                                      Site.Transect == "EF.Comp.9"))
##############################################################################
summary(df.fyke)
summary(df.shock)

```

on to CPUE

- FYKE

```{r}
df.fyke$Effort[df.fyke$Lake == "Center"] <-6
df.fyke$Effort[df.fyke$Lake == "Five Island"] <- 9
df.fyke$Effort[df.fyke$Lake == "Silver"] <- 11
df.fyke$Effort[df.fyke$Lake == "Storm"] <- 10


cpue1<-aggregate(Effort~Lake + Species, data = df.fyke, mean)
df.fyke$dummy<-1
cpue2<-aggregate(dummy~Lake+Species, data = df.fyke, sum)
cpue1$n<-cpue2$dummy
cpue1$fyke.cpue<-cpue1$n/cpue1$Effort
length1<-aggregate(Length..mm.~Lake+Species, data = df.fyke, mean)
length1$Length..in.<-length1$Length..mm./25.4
```

- EFishing

```{r}
df.shock$Effort<-df.shock$On.Time..s./3600


df.shock$dummy<-1
cpue4<-aggregate(dummy~Lake+Species + Site.Transect + Effort, data = df.shock, sum)
cpue4$shock.cpue<-cpue4$dummy/cpue4$Effort
cpue3<-aggregate(shock.cpue ~ Lake + Species, data = cpue4, mean)
length2<-aggregate(Length..mm. ~ Lake + Species, data = df.shock, mean)
length2$Length..in.<-length2$Length..mm./25.4
```

- put that ish together

```{r}

cpue<-merge(cpue1,cpue3, by = c("Species","Lake"), all = T)
cpue<-cpue[,-c(3,4)]
# Melt and cast

cpue.m<-melt(cpue)

cpue.cast<-dcast(cpue.m, Species ~ variable + Lake)

write.csv(cpue.cast, "2018_All_Species_CPUE.csv")
################################################################################
lengths<-merge(length1,length2, by = c("Species","Lake"), all = T)
lengths.m<-melt(lengths)
lengths.cast<-dcast(lengths.m, Species ~ variable + Lake)
write.csv(lengths.cast, "2018_All_Species_Lengths.csv", row.names = F)
```


# And other garbage

this is to attempt to count what age structures we have collected from ALL species in 2018

```{r}
df<-read.csv("2018_COC_BIB_CMR_Data.csv", header = T)
head(df)

# Year
df<-droplevels(subset(df, Year == "2018"))
df<-droplevels(subset(df, Lake == "Center" |
                          Lake == "Five Island" |
                          Lake == "Silver" |
                          Lake == "Storm"))

# now how the trickle am I supposed to figure out which structures we have on hand????
levels(df$Notes)

df<-droplevels(subset(df, Notes == "euthanized" |
                          Notes == "Aging Fish" |
                          Notes == "Aging Fish " |
                          Notes == "Aging Fish - 2017 Tag Loss" |
                          Notes == "Aging Fish - 2017 Tag Loss Female"|
                          Notes == "Aging Fish - Female" |
                          Notes == "Aging Fish - Male" |
                          Notes == "Aging Fish - Mirror" |
                          Notes == "Aging Fish - Scoliosis"|
                          Notes == "Aging Fish 2017 Tag Loss Female" |
                          Notes == "dorsal fin flipped too" |
                          Notes == "dorsal fin clips" ))


df.dummy<-1
df.m<-melt(df)

df.cast<-dcast(df.m, Species ~ variable + Lake, subset = .(variable == "dummy"), fun.aggregate = "sum")


```

