---
title: "Conversion To RMark IMP"
author: "Marty Simonson"
date: "December 6, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Front Matter
```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
packages(RMark)
packages(tidyr)
```

# Loading in the CSV
```{r}
# compiled data
dframe<-read.csv("2018 and 2017 Compiled Data.csv")
head(dframe)
#Setting new "detect" column used in the web resource I'm using.
dframe$detect<-1
str(dframe)
###############
#
#   Well shoot. Looks like I should split up by lake first.
#   I'll do blue (COC ONLY) first (play with MARK models below)
#   then come back and finish splitting the lakes. 
```


# Blue Lake capture history
```{r,}
input.data<-subset(dframe, Lake == "Blue" & Species == "COC" & !is.na(dframe$TAG.NUMBER))


#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","Date"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Date)
head(y)

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
blue.ch<-unite(y,"ch", 2:41, sep = "",remove=T)
blue.ch<-blue.ch[,-c(3:4)]
#blue.ch$freq<-1 --------- does not need to be included if all are 1
# That looks like what we want!


# looking for the 12 incorrect ch (error from data.proc below)




```


# Blue Lake MARK Models
OK. So here's some potential complications:
- unequal time intervals and unequal samples between years (could collapst to month?)
- DUPLICATE TAGS! 8901-9000 is still a problem that's unresolved.
- further, we have no difference between 2017 and 2018 at the moment
- so many carp with only one capture. Such is life.
- this is gonna put lots of dbf and fpt files in my directory Need a new branch I think!
```{r}
str(blue.ch)

# Process.data
blue.proc<-process.data(blue.ch,model="CJS")
# errors around 12 incorrect capture histories

# OK here's the badness: this lake sucks because of the duplicate tags 
# (or falsified data but I really really really doubt that's a thing)

# Gonna have to take a step back and work on another lake

```


# Center Lake carp
```{r}
input.data<-subset(dframe, Lake == "Center" & Species == "COC" & !is.na(dframe$TAG.NUMBER))


#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","Date"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Date)
head(y)
y[is.na(y)]<-0
# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
center.ch<-unite(y,"ch", 2:28, sep = "",remove=T)
center.ch<-center.ch[,-c(3:4)]
```


# Center Lake MARK Models
OK. So here's some potential complications:
- unequal time intervals and unequal samples between years (could collapst to month?)
- need to figure out appropriate model (CJS?)
- all the assumptions!
- also some double-tagged fish. How to handle?
```{r}
str(center.ch)

# Process.data
center.proc<-process.data(center.ch,model="CJS")
center.ddl<-make.design.data(center.proc)
names(center.ddl)
summary(center.ddl$Phi)

center.carp.1<-mark(center.ch) # dot model
PIMS(center.carp.1,"Phi",simplified=F)
# obviously need to consolidate the data to at least month if not more. 
# give it some thought and come back



```

