---
title: "Conversion To RMark IMP"
author: "Marty Simonson"
date: "December 6, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Front Matter
```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
#packages(ggplot2)
#packages(gridExtra)
packages(reshape2)
#packages(FSA)
#packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
#packages(lattice)
#packages(tikzDevice)
#packages(GGally)
#packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(RMark)
packages(tidyr)





```


__Gonna branch into new file for running models, and change this file to terminate at the broadest capture history I can think of, it can always be adjusted. Old chunks of code prior to 1/31/2019 will be moved to bottom of this file. and to new CJS file.__ Robust design coming down the pipe.


# Loading in the CSV
```{r}
# compiled data
dframe<-read.csv("2018 and 2017 Compiled Data.csv")
head(dframe)
#Setting new "detect" column used in the web resource I'm using to create capture history
#       basically just turns into "yes this is observed" and later is summed in casted data frame
dframe$detect<-1
str(dframe)
###############
#
#   Well shoot. Looks like I should split up by lake first.
#   I'll do blue (COC ONLY) and Center (COC and BIB) first (play with MARK models below)
#   then come back and edit models

dframe$Date<-as.Date(dframe$Date, format = "%m/%d/%Y")
str(dframe$Date)

```

-- fun side note today after practice talk (2/1/2019)
- create new capture history: seasonal survival so we have shorter capture history
                            - define seasons and [season length]:
                                  - Spring: March - May [3 months]
                                  - Summer: June - Aug [3 months]
                                  - Fall: Sept - October [2 months]
                                  - Winter: November - February [4 months]
                            - create .xlsx of samples seperately 
                                    (putting in 0 when for each period w/o sample)
                                    
MJW dind't describe how to do this in terms of R, just MARK. But for example, we want
\* tag number *\        00001000                           100;                                where
tag # as comment        ch[length 8]   Lake Groups (binary for center, n. twin, s. twin)

  for each observation



ch[length 8] is 
      - spring 17
      - summer 17
      - fall 17
      - winter 17-18
      - spring 18
      - summer 18
      - fall 18
      - winter 18-19

This will be repeated for each species.
## Carp
```{r}
c.frame<-droplevels(subset(dframe, Species == "COC" &
                        !is.na(dframe$TAG.NUMBER) &
                        Notes != "Aging Fish" &             # remove euthanized 1
                        Fin.Clipped != "Dorsal")) #           remove euthanized 2

summary(c.frame) # verified only one species. 
#                  min  length 122 mm, max 904mm


# remove extraneous columns
c.frame<-c.frame[,-c(1,5,8,9,10,12:22)]

# condense to three lakes of interest for 2017-2018 comparisons
c.frame<-droplevels(subset(c.frame, Lake == "Center" |
                                    Lake == "N. Twin" |
                                    Lake == "S. Twin"))
# down to 3359 observations in three years across two lakes of one species.

# need to create the column to combine month and year
c.frame$YearMonth<-factor(strftime(c.frame$Date, "%Y %m"))
levels(c.frame$YearMonth)
length(levels(c.frame$YearMonth))

# need to delineate year and seasons. the eight cols of capture history
#           there are 15 month/year combinations across three lakes
c.frame$Season[c.frame$YearMonth == "2017 05"] <- "Sp17" # spring 2017 one month
c.frame$Season[c.frame$YearMonth == "2017 06" |
               c.frame$YearMonth == "2017 07" |
               c.frame$YearMonth == "2017 08"] <- "Su17" # Summer 2017 two months
c.frame$Season[c.frame$YearMonth == "2017 09" |
               c.frame$YearMonth == "2017 10"] <- "Fa17" # Fall 2017 two months
c.frame$Season[c.frame$YearMonth == "2017 11" |
               c.frame$YearMonth == "2017 12"] <- "Wi17" # Winter 2017 two months

c.frame$Season[c.frame$YearMonth == "2018 05"] <- "Sp18" # Spring 2018 one month
c.frame$Season[c.frame$YearMonth == "2018 06" |
               c.frame$YearMonth == "2018 07" |
               c.frame$YearMonth == "2018 08"] <- "Su18" # summer 2018 three months
c.frame$Season[c.frame$YearMonth == "2018 09" |
               c.frame$YearMonth == "2018 10"] <- "Fa18" # fall 2018 two months
c.frame$Season[c.frame$YearMonth == "2018 11"] <- "Wi18" # winter 2017 one month
# 15 year month combinations have been converted into seasons (8)

# remove two fish labeled as "no tag" from n/s twin 2017
c.frame<-droplevels(subset(c.frame, TAG.NUMBER != "NO TAG"))
c.frame<-droplevels(subset(c.frame, TAG.NUMBER != "no tag"))


# correcting funky tag names
c.frame$TAG.NUMBER<-as.character(c.frame$TAG.NUMBER)
# "1330 and 1332" renamed to "1330" --- searched and no subsequent 1332 or 1330
c.frame$TAG.NUMBER[c.frame$TAG.NUMBER == "1330 and 1332"]<- "1330"
# "1333 and 1334" renamed to "1333" --- searched and no subsequent 1333 or 1334
c.frame$TAG.NUMBER[c.frame$TAG.NUMBER == "1333 and 1334"]<- "1333"
# "22133/4835" and "4835"
c.frame$TAG.NUMBER[c.frame$TAG.NUMBER == "22133/4835"]<- "22133"
c.frame$TAG.NUMBER[c.frame$TAG.NUMBER == "4835"]<- "22133"
# back to factor
c.frame$TAG.NUMBER<-as.factor(c.frame$TAG.NUMBER)



##############################################33
# time to melt and cast I think:
junk<-melt(c.frame,id.var=c("TAG.NUMBER","Lake","Season"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Season)
head(y)
y<-y[,c(1,4,6,2,8,5,7,3,9)] # reorder columns to desired sequential order
z<-dcast(junk,TAG.NUMBER~Lake)
head(z)


# data frame of whether a lake has a sample or not in each season:
c.samples<-dcast(junk,Lake~Season)



# We have two data frames of 3,220 observations. Need to cbind
yz<-cbind(y,z[,-1])
head(yz)
max(yz[,2:12])
# Set any observations greater than 1 to 1
yz$Sp17[yz$Sp17 >= 2]<-1
yz$Su17[yz$Su17 >= 2]<-1
yz$Fa17[yz$Fa17 >= 2]<-1
yz$Wi17[yz$Wi17 >= 2]<-1

yz$Sp18[yz$Sp18 >= 2]<-1
yz$Su18[yz$Su18 >= 2]<-1
yz$Fa18[yz$Fa18 >= 2]<-1
yz$Wi18[yz$Wi18 >= 2]<-1

yz$Center[yz$Center >= 2]<-1
yz$`N. Twin`[yz$`N. Twin` >= 2]<-1
yz$`S. Twin`[yz$`S. Twin` >= 2]<-1
#
max(yz[,2:12])
#









# In business. pasting capture history, pasting lakes. Then pasting both with a space
# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first following MARK convention:
yz$c<-"/*"
yz$d<-"*/"
yz$TAG.NUMBER<-factor(paste(yz$c,yz$TAG.NUMBER,yz$d,sep = " "))
#nicely done. Now to unite the capture histories
yz<-unite(yz,"ch", 2:9, sep = "",remove=T) # capture history column
yz<-unite(yz,"L:CNS", 3:5, sep = " ",remove=T) # lake column, in Center North Twin South Twin Order
yz[,3]<-factor(paste(yz[,3],";", sep="")) # adding semicolon to end of lake column
yz<-yz[,-c(4:5)] # comment columns removal.
head(yz)

```

__ Writing the .txt file to then manually re-save as .inp file for Program MARK __

copy paste the following to the re-saved file for MARK, adjust as needed:

/* Input file 2019-02-01 from Center, N, & S Twin Lakes between May 2017 and November 2018. */
/* COMMON CARP data. First column is a commented tag number */
/* ch: 8 events, binary split into 4 seasons in 2 years [Spr17, Sum17, ... , Fall18, Win18] */
/* Spring Mar-May 3months, Summer Jun-Aug 3 months, Fall Oct-Sep 2 Months, Winter 4 remain. */
/* Lake recorded as binary variable in alphabetical order: [Center, North Twin, South Twin] */

```{r}
write.table(yz, "CarpCJS.txt", sep = " ",
            quote=F, col.names = F,row.names=F)
```

## Repeat for Buffalo
```{r}
b.frame<-droplevels(subset(dframe, Species == "BIB" &
                        !is.na(dframe$TAG.NUMBER) &
                        Notes != "Aging Fish" &             # remove euthanized 1
                        Fin.Clipped != "Dorsal")) #           remove euthanized 2

summary(b.frame) # verified only one species. 
#                  min  length 122 mm, max 904mm


# remove extraneous columns
b.frame<-b.frame[,-c(1,5,8,9,10,12:22)]

# condense to three lakes of interest for 2017-2018 comparisons
b.frame<-droplevels(subset(b.frame, Lake == "Center" |
                                    Lake == "N. Twin" |
                                    Lake == "S. Twin"))
# down to 6147 observations in three years across two lakes of one species.

# need to create the column to combine month and year
b.frame$YearMonth<-factor(strftime(b.frame$Date, "%Y %m"))
levels(b.frame$YearMonth)
length(levels(b.frame$YearMonth))

# need to delineate year and seasons. the eight cols of capture history
#           there are 15 month/year combinations across three lakes
b.frame$Season[b.frame$YearMonth == "2017 05"] <- "Sp17" # spring 2017 one month
b.frame$Season[b.frame$YearMonth == "2017 06" |
               b.frame$YearMonth == "2017 07" |
               b.frame$YearMonth == "2017 08"] <- "Su17" # Summer 2017 two months
b.frame$Season[b.frame$YearMonth == "2017 09" |
               b.frame$YearMonth == "2017 10"] <- "Fa17" # Fall 2017 two months
b.frame$Season[b.frame$YearMonth == "2017 11" |
               b.frame$YearMonth == "2017 12"] <- "Wi17" # Winter 2017 two months

b.frame$Season[b.frame$YearMonth == "2018 05"] <- "Sp18" # Spring 2018 one month
b.frame$Season[b.frame$YearMonth == "2018 06" |
               b.frame$YearMonth == "2018 07" |
               b.frame$YearMonth == "2018 08"] <- "Su18" # summer 2018 three months
b.frame$Season[b.frame$YearMonth == "2018 09" |
               b.frame$YearMonth == "2018 10"] <- "Fa18" # fall 2018 two months
b.frame$Season[b.frame$YearMonth == "2018 11"] <- "Wi18" # winter 2017 one month
# 15 year month combinations have been converted into seasons (8)

# remove ONE fish labeled as "no tag" from n/s twin WINTER (DECEMBER) 2017
b.frame<-droplevels(subset(b.frame, TAG.NUMBER != "NO TAG PUT IN"))


# correcting funky tag names
b.frame$TAG.NUMBER<-as.character(b.frame$TAG.NUMBER)
# "1304 and 1305" renamed to "1304" --- searched and no subsequent 1304 or 1305
b.frame$TAG.NUMBER[b.frame$TAG.NUMBER == "1304 and 1305"]<- "1304"
# "1308 and 1309" renamed to "1308" --- searched and no subsequent 1308 or 1309
b.frame$TAG.NUMBER[b.frame$TAG.NUMBER == "1308 and 1309"]<- "1308"
# "21118/4314" and "4835"
b.frame$TAG.NUMBER[b.frame$TAG.NUMBER == "21118/4314"]<- "21118"
b.frame$TAG.NUMBER[b.frame$TAG.NUMBER == "4314"]<- "22133"
# "21439/23741" was retagged with the 23741
b.frame$TAG.NUMBER[b.frame$TAG.NUMBER == "21439/23741"]<- "21439"
# back to factor
b.frame$TAG.NUMBER<-as.factor(b.frame$TAG.NUMBER)



##############################################33
# time to melt and cast I think:
junk<-melt(b.frame,id.var=c("TAG.NUMBER","Lake","Season"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Season)
head(y)
y<-y[,c(1,4,6,2,8,5,7,3,9)] # reorder columns to desired sequential order
z<-dcast(junk,TAG.NUMBER~Lake)
head(z)


# data frame of whether a lake has a sample or not in each season:
b.samples<-dcast(junk,Lake~Season)


# We have two data frames of 3,220 observations. Need to cbind
yz<-cbind(y,z[,-1])
head(yz)
max(yz[,2:12])
# Set any observations greater than 1 to 1
yz$Sp17[yz$Sp17 >= 2]<-1
yz$Su17[yz$Su17 >= 2]<-1
yz$Fa17[yz$Fa17 >= 2]<-1
yz$Wi17[yz$Wi17 >= 2]<-1

yz$Sp18[yz$Sp18 >= 2]<-1
yz$Su18[yz$Su18 >= 2]<-1
yz$Fa18[yz$Fa18 >= 2]<-1
yz$Wi18[yz$Wi18 >= 2]<-1

yz$Center[yz$Center >= 2]<-1
yz$`N. Twin`[yz$`N. Twin` >= 2]<-1
yz$`S. Twin`[yz$`S. Twin` >= 2]<-1
#
max(yz[,2:12])
#

# In business. pasting capture history, pasting lakes. Then pasting both with a space
# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first following MARK convention:
yz$c<-"/*"
yz$d<-"*/"
yz$TAG.NUMBER<-factor(paste(yz$c,yz$TAG.NUMBER,yz$d,sep = " "))
#nicely done. Now to unite the capture histories
yz<-unite(yz,"ch", 2:9, sep = "",remove=T) # capture history column
yz<-unite(yz,"L:CNS", 3:5, sep = " ",remove=T) # lake column, in Center North Twin South Twin Order
yz[,3]<-factor(paste(yz[,3],";", sep="")) # adding semicolon to end of lake column
yz<-yz[,-c(4:5)] # comment columns removal.
head(yz)

```

__ Writing the .txt file to then manually re-save as .inp file for Program MARK __

copy paste the following to the re-saved file for MARK, adjust as needed:

/* Input file 2019-02-02 from Center, N, & S Twin Lakes between May 2017 and November 2018. */
/* BIGMOUTH BUFFALO data. First column is a commented tag number */
/* ch: 8 events, binary split into 4 seasons in 2 years [Spr17, Sum17, ... , Fall18, Win18] */
/* Spring Mar-May 3months, Summer Jun-Aug 3months, Fall Oct-Sep 2 Months, Winter 4 remain. */
/* Lake recorded as binary variable in alphabetical order: [Center, North Twin, South Twin] */

```{r}
write.table(yz, "BuffCJS.txt", sep = " ",
            quote=F, col.names = F,row.names=F)
```


## combining frames of samples in month for both spp
```{r}

checkingbothspecies<-rbind(c.samples,b.samples)
# visually see that all the zeros line up for both species. back to just one

yz<-c.samples
yz<-yz[,c(1,4,6,2,8,5,7,3,9)] # setting desired order

yz$Sp17[yz$Sp17 >= 2]<-1
yz$Su17[yz$Su17 >= 2]<-1
yz$Fa17[yz$Fa17 >= 2]<-1
yz$Wi17[yz$Wi17 >= 2]<-1

yz$Sp18[yz$Sp18 >= 2]<-1
yz$Su18[yz$Su18 >= 2]<-1
yz$Fa18[yz$Fa18 >= 2]<-1
yz$Wi18[yz$Wi18 >= 2]<-1

write.csv(yz,"CB_CJS_Samples.csv", row.names = F)
```







# One Capture History To Rule Them All
- removing non-tagged species
- removing euthanized carp and buff
                  - *known fates: Storm, etc.*
                  - *funky notation and double tags: consistency*
                  - *duplicate 100 tags at Blue Lake 2018*
- clean/groom the factors and their unused levels

1.) consolidate relevant individual covariates & variables:
          - Year
          - Month
          - MonthYear
          - Day
          - Lake
          - TL
          - Weight
          - Notes (append to end)
          - Gear
          - Water Temp

2.) Melt and cast to pivot table of capture history
```{r}
#Input data 
input.data<-subset(dframe, !is.na(dframe$TAG.NUMBER) &         # non-coc/bib species won't have tags or envelope #
                           Notes != "Aging Fish" &             # remove euthanized 1
                           Fin.Clipped != "Dorsal")            # remove euthanized 2      
input.data$TAG.NUMBER<-droplevels(input.data$TAG.NUMBER)       # remove all factor levels that dont have observation
                                                               # check for any factors with envelope numbers (dead)
summary(input.data$TAG.NUMBER)#                                  check fish with most obs/recaps
input.data$Species<-droplevels(input.data$Species)#              remove all unused factor levels 
summary(input.data$Species)#                                     raw # individuals, hopefully many recaps
  
input.data$Year<-factor(strftime(input.data$Date, "%Y"))#        new column of year
input.data$Month<-factor(strftime(input.data$Date, "%m"))#        new column of month
input.data$Day<-factor(strftime(input.data$Date, "%d"))#        new column of day
input.data$YearMonth<-factor(strftime(input.data$Date, "%Y %m"))#        new column of year month
head(input.data$YearMonth)
levels(input.data$YearMonth)#                                     levels line up with samples



#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","Date","Year", "Month", "Day", "YearMonth", "Lake",
                               "Total.Length..mm.","Weight..g.","Gear","Water.Temp..C.","Notes"),
           measure.vars="detect")

# casting the whole kit n caboodle
y<-dcast(junk,TAG.NUMBER+Species+Year+Month+Day+YearMonth+Lake+Total.Length..mm.+Weight..g.+Gear+Water.Temp..C.+Notes
         ~Date)
head(y)
#y[is.na(y)]<-0# not needed after defaulting to length, although some are 2

#############################
#
#     Really need to write function/for() loop to change multiple counts to just 1
#
#############################

# That way I can cast based on the question at hand and figure out the business.





# after consolidating to month there are some fish with 2 captures per month 
# so need to correct those
y$`2017 05`[y$`2017 05` >= 2] <- 1
y$`2017 06`[y$`2017 06` >= 2] <- 1
y$`2018 05`[y$`2018 05` >= 2] <- 1
y$`2018 06`[y$`2018 06` >= 2] <- 1
y$`2018 07`[y$`2018 07` >= 2] <- 1
y$`2018 08`[y$`2018 08` >= 2] <- 1
y$`2018 09`[y$`2018 09` >= 2] <- 1
y$`2018 10`[y$`2018 10` >= 2] <- 1

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first following MARK convention:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-factor(paste(y$c,y$TAG.NUMBER,y$d,sep = " "))
#nicely done. Now to unite the capture histories
center.ch<-unite(y,"ch", 3:10, sep = "",remove=T)
center.ch<-center.ch[,-c(4:5)] # comment columns removal.
head(center.ch)
```






















# Blue Lake capture history
```{r,}
input.data<-subset(dframe, Lake == "Blue" & Species == "COC" & !is.na(dframe$TAG.NUMBER))


#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","Date"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Date)
head(y)

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
blue.ch<-unite(y,"ch", 2:41, sep = "",remove=T)
blue.ch<-blue.ch[,-c(3:4)]
#blue.ch$freq<-1 --------- does not need to be included if all are 1
# That looks like what we want!


# looking for the 12 incorrect ch (error from data.proc below)




```


# Blue Lake MARK Models
OK. So here's some potential complications:
- unequal time intervals and unequal samples between years (could collapst to month?)
- DUPLICATE TAGS! 8901-9000 is still a problem that's unresolved.
- further, we have no difference between 2017 and 2018 at the moment
- so many carp with only one capture. Such is life.
- this is gonna put lots of dbf and fpt files in my directory Need a new branch I think!
```{r}
str(blue.ch)

# Process.data
blue.proc<-process.data(blue.ch,model="CJS")
# errors around 12 incorrect capture histories

# OK here's the badness: this lake sucks because of the duplicate tags 
# (or falsified data but I really really really doubt that's a thing)

# Gonna have to take a step back and work on another lake

```


# Center Lake carp: 2018 only
```{r}
input.data<-subset(dframe, Lake == "Center" &                  # Lake
                           Species == "COC" &                  # Species
                           !is.na(dframe$TAG.NUMBER) &         # must have a tag number
                           Year == "2018" &                    # within a year
                           Notes != "Aging Fish" &             # remove euthanized 1
                           Fin.Clipped != "Dorsal")            # remove euthanized 2
# check tag numbers
levels(droplevels(input.data$TAG.NUMBER))
summary(input.data$TAG.NUMBER)

input.data$Month<-strftime(input.data$Date, "%m")
levels(factor(input.data$Month))


#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","Month"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER~Month)
head(y)

#y[is.na(y)]<-0
# after consolidating to month there are some fish with 2 captures per month so need to correct those

y$`05`[y$`05` >= 2] <- 1
#y$`06`[y$`06` >= 2] <- 1
#y$`07`[y$`07` >= 2] <- 1
#y$`08`[y$`08` >= 2] <- 1
#y$`09`[y$`09` >= 2] <- 1
#y$`10`[y$`10` >= 2] <- 1

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
center.ch<-unite(y,"ch", 2:7, sep = "",remove=T)
center.ch<-center.ch[,-c(3:4)] # comment columns removal.

```


## Center Lake MARK Models
OK. So here's some potential complications:
- unequal time intervals and unequal samples between years (could collapse to month?)
- need to figure out appropriate model (CJS?)
- all the assumptions!
- also some double-tagged fish. How to handle?
```{r}
str(center.ch)

# Process.data
center.proc<-process.data(center.ch,model="CJS")
str(center.proc)



center.ddl<-make.design.data(center.proc)
names(center.ddl)
summary(center.ddl$Phi)

center.carp.1<-mark(center.ch) # dot model
PIMS(center.carp.1,"Phi",simplified=F) # pims for full time dependent model

# Other model structures:
phi.0<-list(formula=~1) # dot model
phi.1<-list(formula=~time) # time dependent
phi.2<-list(formula=~Time) # Time trend
p.1<-list(formula=~1) # dot model recapture rate
p.2<-list(formula=~time) # time dependent recapture

# Running other models
#
#   See if Phi-dot and phi-dot match above
#             Phi = 0.20586
#             p = 0.0698

center.m.1<-mark(center.proc, center.ddl, model.parameters=list(Phi=phi.0,p=p.1))
# dot model matches

# Phi time and p dot
center.m.2<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.1,p=p.1))

# Phi time trend and p dot
center.m.3<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.2,p=p.1))

# phi time and p time
center.m.4<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.1,p=p.2))

# phi dot and p time
center.m.5<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.0,p=p.2))

# phi Time trend and p time
center.m.6<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.2,p=p.2))

center.model.results<-collect.models()
center.model.results


```


## Center Lake: Carp and Buffalo
- need to separate months by year. Probs combined column
- and trying to run for both buffalo and carp together (as groups). 
```{r}
input.data<-subset(dframe, Lake == "Center" & 
                           !is.na(dframe$TAG.NUMBER) &         # non-coc/bib species won't have tags or envelope #
                           Notes != "Aging Fish" &             # remove euthanized 1
                           Fin.Clipped != "Dorsal")            # remove euthanized 2      
input.data$TAG.NUMBER<-droplevels(input.data$TAG.NUMBER)       # remove all factor levels that dont have observation
                                                               # check for any factors with envelope numbers (dead)
summary(input.data$TAG.NUMBER)#                                  check fish with most obs/recaps
input.data$Species<-droplevels(input.data$Species)#              remove all unused factor levels 
summary(input.data$Species)#                                     raw # individuals, hopefully many recaps
  
input.data$MonthYear<-factor(strftime(input.data$Date, "%Y %m"))# new column of month and year
head(input.data$MonthYear)
levels(input.data$MonthYear)#                                     levels line up with samples



#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","MonthYear", "Species"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER+Species~MonthYear)
head(y)
#y[is.na(y)]<-0                                   # not needed after defaulting to length, although some are 2



# after consolidating to month there are some fish with 2 captures per month 
# so need to correct those
y$`2017 05`[y$`2017 05` >= 2] <- 1
y$`2017 06`[y$`2017 06` >= 2] <- 1
y$`2018 05`[y$`2018 05` >= 2] <- 1
y$`2018 06`[y$`2018 06` >= 2] <- 1
y$`2018 07`[y$`2018 07` >= 2] <- 1
y$`2018 08`[y$`2018 08` >= 2] <- 1
y$`2018 09`[y$`2018 09` >= 2] <- 1
y$`2018 10`[y$`2018 10` >= 2] <- 1

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first following MARK convention:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-factor(paste(y$c,y$TAG.NUMBER,y$d,sep = " "))
#nicely done. Now to unite the capture histories
center.ch<-unite(y,"ch", 3:10, sep = "",remove=T)
center.ch<-center.ch[,-c(4:5)] # comment columns removal.
head(center.ch)
```


# Center Lake COC&BIB MARK models 2017-2018 by month
SETTING UP PROCESS DATA AND DESIGN DATA
```{r}
# SETTING UP PROCESS DATA AND DESIGN DATA
str(center.ch)

# Process.data
center.proc<-process.data(center.ch,#         the data frame
                          model="CJS",#       cormack jolly seber paramaterization
                          time.intervals = c(1,11,1,1,1,1,1),#    center only may-june 17 then begin may18
                          groups = "Species")#  grouping by species
str(center.proc)

release.gof(center.proc) # why won't this work? need to know goodness of fit.

center.ddl<-make.design.data(center.proc) # makes design data that can be formed into PIMS
names(center.ddl)
summary(center.ddl$Phi) # Let's estimate these
```

Model structures:
building up to the hypothesis that survival is species- and lake- and time-dependent. 

__Shoot, from here it's getting too complicated. Gonna branch into new file for running models, and change this file to terminate at the broadest capture history I can think of, it can always be adjusted. 1/31/2019__


```{r}
# model structures:
#       Phi
phi.0<-list(formula=~1) # dot model
phi.1<-list(formula=~time) # time dependent
phi.2<-list(formula=~Time) # Time trend
phi.species<-list(formula=~Species) # group by species
phi.species.t<-list(formula=~Species+time) # species plus time
#       p
p.0<-list(formula=~1) # dot model recapture rate
p.t<-list(formula=~time) # time dependent recapture
p.species<-list(formula=~Species) # group by species
p.species.t<-list(formula=~Species+time) # species and time


################################################################################
# dot model with no variation between species
center.m.1<-mark(center.proc, center.ddl, model.parameters=list(Phi=phi.0,p=p.1))

################################################################################
# Phi time and p dot with no variation between species
center.m.2<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.1,p=p.1))

# Phi species and p dot
center.m.spp<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.species,p=p.1))

# Phi species and time and p dot
center.m.spp.t<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.species.t,p=p.1))

# Phi time trend and p dot
center.m.3<-mark(center.proc,center.ddl,model.parameters=list(Phi=phi.2,p=p.1))


################################################################################
# phi dot and p time with no species variation
center.m.4<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.0,p=p.2))

# phi time and p time with no species variation
center.m.5<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.1,p=p.2))

# phi Time trend and p time with no variation in species
center.m.6<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.2,p=p.2))

# phi dot and p species
center.m.phiDot.pSpp<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.0,p=p.species))

# phi time and p species
center.m.phiDot.pSpp<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.1,p=p.species))

# phi dot and p species + time
center.m.phiDot.pSpp<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.0,p=p.species.t))

# phi time and p species + time
center.m.phiDot.pSpp<-mark(center.proc,center.ddl,model.parameters = list (Phi=phi.1,p=p.species.t))


center.model.results<-collect.models()
center.model.results



```


# North Twin Lake: Carp and Buffalo
- need to separate months by year. Probs combined column
- and trying to run for both buffalo and carp together (as groups). 
```{r}
levels(dframe$Lake)
input.data<-subset(dframe, Lake == "N. Twin" & 
                           !is.na(dframe$TAG.NUMBER) &
                           Notes != "Aging Fish" &             # remove euthanized 1
                           Fin.Clipped != "Dorsal")            # remove euthanized 2   
input.data<-subset(input.data, TAG.NUMBER != "no tag")
levels(droplevels(input.data$TAG.NUMBER))
input.data$TAG.NUMBER<-droplevels(input.data$TAG.NUMBER)
summary(input.data$TAG.NUMBER)
input.data$Species<-droplevels(input.data$Species)
summary(input.data$Species)

input.data$MonthYear<-factor(strftime(input.data$Date, "%Y %m"))
head(input.data$MonthYear)
levels(input.data$MonthYear)



#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","MonthYear", "Species"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER+Species~MonthYear)
head(y)
#y[is.na(y)]<-0
# after consolidating to month there are some fish with 2 captures per month so need to correct those

y$`2017 06`[y$`2017 06` >= 2] <- 1
y$`2017 07`[y$`2017 07` >= 2] <- 1
y$`2017 08`[y$`2017 08` >= 2] <- 1
y$`2017 09`[y$`2017 09` >= 2] <- 1
y$`2017 10`[y$`2017 10` >= 2] <- 1
y$`2017 11`[y$`2017 11` >= 2] <- 1
y$`2018 05`[y$`2018 05` >= 2] <- 1
y$`2018 06`[y$`2018 06` >= 2] <- 1
y$`2018 07`[y$`2018 07` >= 2] <- 1
y$`2018 08`[y$`2018 08` >= 2] <- 1
y$`2018 09`[y$`2018 09` >= 2] <- 1
y$`2018 10`[y$`2018 10` >= 2] <- 1
y$`2018 11`[y$`2018 11` >= 2] <- 1

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
Ntwin.ch<-unite(y,"ch", 3:15, sep = "",remove=T)
Ntwin.ch<-Ntwin.ch[,-c(4:5)] # comment columns removal.

```


# North Twin Lake COC&BIB MARK models 2017-2018 by month

```{r}
# SETTING UP PROCESS DATA AND DESIGN DATA
str(Ntwin.ch)

# Process.data
Ntwin.proc<-process.data(Ntwin.ch,
                          model="CJS",
                          time.intervals = c(1,1,1,1,1,6,1,1,1,1,1,1),
                          groups = "Species")
str(Ntwin.proc)

release.gof(Ntwin.proc)

Ntwin.ddl<-make.design.data(Ntwin.proc)
names(Ntwin.ddl)
summary(Ntwin.ddl$Phi)
```

```{r}


# model structures:
phi.0<-list(formula=~1) # dot model
phi.1<-list(formula=~time) # time dependent
phi.2<-list(formula=~Time) # Time trend
phi.species<-list(formula=~Species)
phi.species.t<-list(formula=~Species+time)
p.1<-list(formula=~1) # dot model recapture rate
p.2<-list(formula=~time) # time dependent recapture
p.species<-list(formula=~Species)
p.species.t<-list(formula=~Species+time)


################################################################################
# dot model with no variation between species
Ntwin.m.1<-mark(Ntwin.proc, Ntwin.ddl, model.parameters=list(Phi=phi.0,p=p.1))

################################################################################
# Phi time and p dot with no variation between species
Ntwin.m.2<-mark(Ntwin.proc,Ntwin.ddl,model.parameters=list(Phi=phi.1,p=p.1))

# Phi species and p dot
Ntwin.m.spp<-mark(Ntwin.proc,Ntwin.ddl,model.parameters=list(Phi=phi.species,p=p.1))

# Phi species and time and p dot
Ntwin.m.spp.t<-mark(Ntwin.proc,Ntwin.ddl,model.parameters=list(Phi=phi.species.t,p=p.1))

# Phi time trend and p dot
Ntwin.m.3<-mark(Ntwin.proc,Ntwin.ddl,model.parameters=list(Phi=phi.2,p=p.1))


################################################################################
# phi dot and p time with no species variation
Ntwin.m.4<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.0,p=p.2))

# phi time and p time with no species variation
Ntwin.m.5<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.1,p=p.2))

# phi Time trend and p time with no variation in species
Ntwin.m.6<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.2,p=p.2))

# phi dot and p species
Ntwin.m.phiDot.pSpp<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.0,p=p.species))

# phi time and p species
Ntwin.m.phiDot.pSpp<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.1,p=p.species))

# phi dot and p species + time
Ntwin.m.phiDot.pSpp<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.0,p=p.species.t))

# phi time and p species + time
Ntwin.m.phiDot.pSpp<-mark(Ntwin.proc,Ntwin.ddl,model.parameters = list (Phi=phi.1,p=p.species.t))


Ntwin.model.results<-collect.models()
Ntwin.model.results


Ntwin.m.6$results$lnl # code format to examine -2 log likelihood for models. Perhaps able to append it to model.results table with more code.
```


# South Twin Lake: Carp and Buffalo
- need to separate months by year. Probs combined column
- and trying to run for both buffalo and carp together (as groups). 
```{r}
levels(dframe$Lake)
input.data<-subset(dframe, Lake == "S. Twin" & 
                           !is.na(dframe$TAG.NUMBER) &
                           Notes != "Aging Fish" &             # remove euthanized 1
                           Fin.Clipped != "Dorsal")            # remove euthanized 2   
input.data<-subset(input.data, TAG.NUMBER != "NO TAG" & TAG.NUMBER != "NO TAG PUT IN")
levels(droplevels(input.data$TAG.NUMBER))
input.data$TAG.NUMBER<-droplevels(input.data$TAG.NUMBER)
summary(input.data$TAG.NUMBER)
input.data$Species<-droplevels(input.data$Species)
summary(input.data$Species)

input.data$MonthYear<-factor(strftime(input.data$Date, "%Y %m"))
head(input.data$MonthYear)
levels(input.data$MonthYear)



#reshaping functions to pivot the data
junk<-melt(input.data,id.var=c("TAG.NUMBER","MonthYear", "Species"),
           measure.vars="detect")
y<-dcast(junk,TAG.NUMBER+Species~MonthYear)
head(y)
#y[is.na(y)]<-0
# after consolidating to month there are some fish with 2 captures per month so need to correct those

y$`2017 06`[y$`2017 06` >= 2] <- 1
y$`2017 07`[y$`2017 07` >= 2] <- 1
y$`2017 08`[y$`2017 08` >= 2] <- 1
y$`2017 09`[y$`2017 09` >= 2] <- 1
y$`2017 10`[y$`2017 10` >= 2] <- 1
y$`2017 11`[y$`2017 11` >= 2] <- 1
y$`2017 12`[y$`2017 12` >= 2] <- 1
y$`2018 05`[y$`2018 05` >= 2] <- 1
y$`2018 06`[y$`2018 06` >= 2] <- 1
y$`2018 07`[y$`2018 07` >= 2] <- 1
y$`2018 08`[y$`2018 08` >= 2] <- 1
y$`2018 09`[y$`2018 09` >= 2] <- 1
y$`2018 10`[y$`2018 10` >= 2] <- 1
y$`2018 11`[y$`2018 11` >= 2] <- 1

# collapse the capture history to single column
# capture history: using "unite" in dplyr
# Want to make each tag a "comment" first:
y$c<-"/*"
y$d<-"*/"
y$TAG.NUMBER<-paste(y$c,y$TAG.NUMBER,y$d,sep = " ")
#nicely done. Now to unite the capture histories
Stwin.ch<-unite(y,"ch", 3:16, sep = "",remove=T)
Stwin.ch<-Stwin.ch[,-c(4:5)] # comment columns removal.

```


# North Twin Lake COC&BIB MARK models 2017-2018 by month

```{r}
# SETTING UP PROCESS DATA AND DESIGN DATA
str(Stwin.ch)

# Process.data
Stwin.proc<-process.data(Stwin.ch,
                          model="CJS",
                          time.intervals = c(1,1,1,1,1,1,5,1,1,1,1,1,1),
                          groups = "Species")
str(Stwin.proc)

release.gof(Stwin.proc)

Stwin.ddl<-make.design.data(Stwin.proc)
names(Stwin.ddl)
summary(Stwin.ddl$Phi)
```

```{r}


# model structures:
phi.0<-list(formula=~1) # dot model
phi.1<-list(formula=~time) # time dependent
phi.2<-list(formula=~Time) # Time trend
phi.species<-list(formula=~Species)
phi.species.t<-list(formula=~Species+time)
p.1<-list(formula=~1) # dot model recapture rate
p.2<-list(formula=~time) # time dependent recapture
p.species<-list(formula=~Species)
p.species.t<-list(formula=~Species+time)


################################################################################
# dot model with no variation between species
Stwin.m.1<-mark(Stwin.proc, Stwin.ddl, model.parameters=list(Phi=phi.0,p=p.1))

################################################################################
# Phi time and p dot with no variation between species
Stwin.m.2<-mark(Stwin.proc,Stwin.ddl,model.parameters=list(Phi=phi.1,p=p.1))

# Phi species and p dot
Stwin.m.spp<-mark(Stwin.proc,Stwin.ddl,model.parameters=list(Phi=phi.species,p=p.1))

# Phi species and time and p dot
Stwin.m.spp.t<-mark(Stwin.proc,Stwin.ddl,model.parameters=list(Phi=phi.species.t,p=p.1))

# Phi time trend and p dot
Stwin.m.3<-mark(Stwin.proc,Stwin.ddl,model.parameters=list(Phi=phi.2,p=p.1))


################################################################################
# phi dot and p time with no species variation
Stwin.m.4<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.0,p=p.2))

# phi time and p time with no species variation
Stwin.m.5<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.1,p=p.2))

# phi Time trend and p time with no variation in species
Stwin.m.6<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.2,p=p.2))

# phi dot and p species
Stwin.m.phiDot.pSpp<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.0,p=p.species))

# phi time and p species
Stwin.m.phiDot.pSpp<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.1,p=p.species))

# phi dot and p species + time
Stwin.m.phiDot.pSpp<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.0,p=p.species.t))

# phi time and p species + time
Stwin.m.phiDot.pSpp<-mark(Stwin.proc,Stwin.ddl,model.parameters = list (Phi=phi.1,p=p.species.t))


Stwin.model.results<-collect.models()
Stwin.model.results


Stwin.m.6$results$lnl # code format to examine -2 log likelihood for models. Perhaps able to append it to model.results table with more code.



