---
title: "2018 EDA_Summary"
author: "Marty Simonson"
date: "November 12, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
```



# Load in data and summarize, look for glaring errors:
```{r}
data<-read.csv("2018_Data_compiled_2.0.csv",header=T)
# data includes seine, electrofishing, and fyke nets
#           includes all species data from spring std runs (may) and fall fyke netting (sept)
#           some bigly data here, boys.


str(data)

############## 
#
#     Data Management: Columns formatting
#
#

# Date
levels(data$Date)
data$Date<-as.Date(as.character(data$Date), format = "%m/%d/%Y")
str(data$Date)


# Lake
summary(data$Lake)
data$Lake[data$Lake == "SIlver"] <- "Silver"
data$Lake[data$Lake == "silver"] <- "Silver"
data$Lake[data$Lake == "Silver (Dickinson)"] <- "Silver"
data$Lake[data$Lake == "5 Island"] <- "Five Island"
data$Lake[data$Lake == "N Twin"] <- "N. Twin"
data$Lake[data$Lake == "S Twin"] <- "S. Twin"
data$Lake<-droplevels(data$Lake)
summary(data$Lake)

# TAG.NUMBER - Already a factor
summary(data$TAG.NUMBER)
data$TAG.NUMBER[data$TAG.NUMBER == ""] <- NA # set non-tagged fish to NA
summary(data$TAG.NUMBER)
data$TAG.NUMBER<-droplevels(data$TAG.NUMBER)
    # NAs checked
    # One fish in Storm (20375) captured 4 times in 2018
    # When doing pivot tables / .INP files for MARK I'll probably find more errors.

# Species
levels(data$Species) #KOI could be an issue, as with BMB and CAP for abbreviations
summary(data$Species)
summary(data$Lake[data$Species == "KOI"])
data$Species<-as.character(data$Species)
data$Species[data$Species == "KOI"] <- "COC"
data$Species[data$Species == "CAP"] <- "COC"
data$Species[data$Species == "BMB"] <- "BIB"
data$Species[data$Species == "YLP"] <- "YEP"
data$Species[data$Species == "YL Bass"] <- "YLB"
data$Species[data$Species == "YBH"] <- "YEB"
data$Species[data$Species == "GZS"] <- "GIS"
#data$Lake[data$Species == "YLB"]
data$Species[data$Species == "painted turtle"] <- "Painted turtle"
data$Species[data$Species == "Spottail Shiner"] <- "Spottail shiner"
data$Species[data$Species == "White Sucker"] <- "WHS"
data$Species[data$Species == "Northern Pike"] <- "NOP"
data$Species[data$Species == "Shortnose Gar"] <- "SNG"
data$Species[data$Species == "GSFxBLG"] <- "BLGxGSF"
data$Species<-as.factor(data$Species)
summary(data$Species)
# Confirm the abbreviations: Yellow Bullhead = YEB; Yellow Bass = YLB
#       Confirmed 1/3/2019

#       HOWEVER! Found out that the YLB in storm are really WHB. Super awesome
#       Thanks jon meerbeek though.
data$Species[data$Lake == "Storm" & 
             data$Species == "YLB"] <- "WHB"



# Length..mm.
str(data$Length..mm.)
summary(data$Length..mm.)
data$Length..mm.[data$Length..mm. ==""] <- NA

data$Length..mm.<-droplevels(data$Length..mm.)
data$Length..mm.<-as.numeric(as.character(data$Length..mm.))

summary(data$Length..mm.)

# seems legit! 1,210 missing values.
# Update 1/3/2019 we've added many many bached fish from the fyke nets. mostly yoy white bass

# Min 26mm is YLB in 5I; Max is 958 buffalo in Storm
# Don't forget to put the character command within the numeric!


# Weight..g.
# Needs to be rounded to whole numbers
summary(data$Weight..g.[data$Species == "COC"]) # Max is good (24.5 lb carp in Silver), minimum is also silver (11 inch fish .5 lbs)
summary(data$Weight..g.[data$Species == "BIB"]) # Maximum is good (30.6 lb buffalo in Storm) minimum is non-tagged voucher fish from N twin.
summary(data$Weight..g.)
# Minimum is 1... many small YOY fish
data$Weight..g.<-round(data$Weight..g.) # set to just grams

#######
#
# Recap Column
#
summary(data$Recap)
# Simply set recaps to 1. I'll double check each "1" after. Year is currently irrelevant (i'll add 2017 tag data later)
data$Recap[data$Recap == "2017 R"] <- "1"
data$Recap[data$Recap == "2017 R "] <- "1"
data$Recap[data$Recap == "2017R "] <- "1"
data$Recap[data$Recap == "2017R"] <- "1"
data$Recap[data$Recap == "2018 R"] <- "1"
data$Recap[data$Recap == "2018R"] <- "1"
data$Recap[data$Recap == "R"] <- "1"
data$Recap[data$Recap == "r 2017"] <- "1"
data$Recap[data$Recap == "R 2017"] <- "1"
data$Recap[data$Recap == "r 2018"] <- "1"
data$Recap[data$Recap == "R 2018"] <- "1"
data$Recap[data$Recap == "R2018"] <- "1"
data$Recap[data$Recap == "recap 2017"] <- "1"
data$Recap[data$Recap == "recap 2018"] <- "1"
data$Recap[data$Recap == "Recap 2018"] <- "1"
data$Recap[data$Recap == "2017 tag loss"] <- ""
data$Recap[data$Recap == "2017 tag loss R"] <- ""
data$Recap[data$Recap == "2017 tagloss"] <- ""
data$Recap[data$Recap == "2018 tag loss "] <- ""
data$Recap[data$Recap == "2018 tag loss R"] <- ""
data$Recap[data$Recap == "R (tag loss)"] <- ""

# Those entered as tag loss will be entered as "1" in tag loss column
data$Tag.Loss[data$Recap == "2017 tag loss"] <- "1"
data$Tag.Loss[data$Recap == "2017 tag loss R"] <- "1"
data$Tag.Loss[data$Recap == "2017 tagloss"] <- "1"
data$Tag.Loss[data$Recap == "2018 tag loss "] <- "1"
data$Tag.Loss[data$Recap == "2018 tag loss R"] <- "1"
data$Tag.Loss[data$Recap == "R (tag loss)"] <- "1"
data$Recap<-droplevels(data$Recap)
summary(data$Recap)
##### Need to check number of recaps (443 in Google, 503 here. Google does not include 2017-only recaps)
##### Need to make sure each recap in this column has comment


#####################
#
# Tag Loss
#
str(factor(data$Tag.Loss))
data$Tag.Loss[data$Tag.Loss == "2017"] <- "1"
data$Tag.Loss[data$Tag.Loss == "2018"] <- "1"
data$Tag.Loss<-factor(data$Tag.Loss)
summary(data$Tag.Loss)
# 56 tags lost including both years. 

########################################################
#### Need to check for comments in each of 56 tags lost!
########################################################



############
#
#   Fin Clipped
#
#Levels
summary(data$Fin.Clipped)
data$Fin.Clipped[data$Fin.Clipped == "dorsal"] <-"Dorsal"
data$Fin.Clipped[data$Fin.Clipped == "Both Pectoral"] <- "R/L Pectoral"
data$Fin.Clipped[data$Fin.Clipped == "N/A"] <-NA
data$Fin.Clipped[data$Fin.Clipped == ""] <- NA
data$Fin.Clipped<-droplevels(data$Fin.Clipped)
summary(data$Fin.Clipped)

#############
#
#     Notes
#
summary(data$Notes)
# just making some more consistent
data$Notes[data$Notes == "2018 recap"] <- "2018 Recap"
data$Notes[data$Notes == "male"] <- "Male"
data$Notes[data$Notes == "female"] <- "Female"
data$Notes[data$Notes == "baby"] <- "Baby"
# consider adding gender column?????
data$Notes[data$Notes == "Baby fish"] <- "Baby"
data$Notes[data$Notes == "biggie"] <- "Rotund"
data$Notes[data$Notes == "Biggie"] <- "Rotund"
data$Notes[data$Notes == "Chubby"] <- "Rotund"
data$Notes[data$Notes == "Chubs"] <- "Rotund"
data$Notes[data$Notes == "DELTS"] <- "DELTs"
data$Notes[data$Notes == "Fat"] <- "Rotund"
data$Notes[data$Notes == "Fatty"] <- "Rotund"
data$Notes[data$Notes == "Fatty DTF"] <- "Rotund Female"
data$Notes[data$Notes == "Football"] <- "Rotund"
data$Notes[data$Notes == "Hog"] <- "Rotund"
data$Notes[data$Notes == "Plump girl"] <- "Rotund"
data$Notes[data$Notes == "probs female very round"] <- "Rotund Female"
data$Notes[data$Notes == "rotund"] <- "Rotund"
data$Notes[data$Notes == "ROTUND"] <- "Rotund"
#data$Notes[data$Notes == "rotund male"] <- "Rotund Male"
data$Notes[data$Notes == "Thick"] <- "Rotund"
data$Notes<-droplevels(data$Notes)
summary(data$Notes)
#######


#######
#
# Gear
levels(data$Gear)
data$Gear[data$Gear == "Seine"] <- "Commercial Seine"
data$Gear<-droplevels(data$Gear)
summary(data$Gear)

#########
#
#  Site.Transect
summary(data$Site.Transect)
# need to correct 1+ ones (2 through 8)
# need to correct ALL + MM samples (combine effort)
data$Site.Transect[data$Site.Transect == "Extra 1"] <- "Bonus 1"
data$Site.Transect[data$Site.Transect == "Extra 2"] <- "Bonus 2"
data$Site.Transect[data$Site.Transect == "Extra 3"] <- "Bonus 3"
# what's up with the NAs? they are fyke nets. changed to fyke in excel
data$Site.Transect[data$Site.Transect == "Standard 1"] <- "1"
data$Site.Transect[data$Site.Transect == "Standard 2"] <- "2"
data$Site.Transect[data$Site.Transect == "Standard 3"] <- "3"
data$Site.Transect[data$Site.Transect == "Standard 4"] <- "4"
data$Site.Transect[data$Site.Transect == "Bonus 1 (bridge)"] <- "Bridge"
data$Site.Transect<-droplevels(data$Site.Transect)
summary(data$Site.Transect)

################
#
# On.Time..s.
#

summary(data$On.Time..s.)
str(data$On.Time..s.)
# check tag 18499 and 19026: gear listed as seine but sites given std transect names, but also seine on that day 5/22/18 five island --- checked and has already been corrected 1/4/2018

# check 7/26/2018  Silver lake transect 6 has NA for effort. (possibly from Jon Meerbeek scanned sheets?) Those are the only NAs left so can find and replace if needed



# then all blanks (9961 of them) can become NAs because they come from either seines or fykes
data$On.Time..s.<-as.numeric(data$On.Time..s.)
summary(data$On.Time..s.)




##########
#
# water temp
#
summary(data$Water.Temp..C.)
# actually looks good! 88 degrees one day apparently but other than that 2.4 was nov. 8 so that's happy

##############
#
# Secchi
#
summary(data$Secchi..cm.)
data$Secchi..cm.<-as.character(data$Secchi..cm.)
data$Secchi..cm.[data$Secchi..cm. == "210cm"]<-"210"
data$Secchi..cm.<-as.numeric(data$Secchi..cm.)
summary(data$Secchi..cm.)
# this is fairly junky data. Mostly absent and i plan to use more of Ellen's data anyway.

#########
#
# Weather
#
levels(data$Weather)
# more arbitrary than anything. more or less useless

##########
#
# Add year column for later comparisons
data$Year<-factor("2018")
# at this time 2018 data is entered and looking OK.


###################################################################################################
#
#
#           Need to specify Standard Transects vs. Run N Gun
#           Need to specify ALL SPECIES spring run as well.
levels(data$Site.Transect)
levels(data$Start.Latitude)

summary(data$Length..mm.[data$Lake == "Center" &
                         data$Species == "YLB"]) 
# forked up man. No WHB recorded but listed as popular fish in the lake. feck.
# anyway let's set the Site.Transect

data$Site.Transect<-as.character(data$Site.Transect)
# Center spring ef
data$Site.Transect[data$Date == as.Date("2018-05-17") &
                   data$Lake == "Center" &
                   data$Site.Transect == "1"] <- "EF.Comp.1"
data$Site.Transect[data$Date == as.Date("2018-05-17") &
                   data$Lake == "Center" &
                   data$Site.Transect == "2"] <- "EF.Comp.2"
data$Site.Transect[data$Date == as.Date("2018-05-17") &
                   data$Lake == "Center" &
                   data$Site.Transect == "3"] <- "EF.Comp.3"
data$Site.Transect[data$Date == as.Date("2018-05-17") &
                   data$Lake == "Center" &
                   data$Site.Transect == "4"] <- "EF.Comp.4"
data$Site.Transect[data$Date == as.Date("2018-05-17") &
                   data$Lake == "Center" &
                   data$Site.Transect == "5"] <- "EF.Comp.5"
# Five Island spring EF
data$Site.Transect[data$Date == as.Date("2018-05-23") &
                   data$Lake == "Five Island" &
                   data$Site.Transect == "1"] <- "EF.Comp.1"
data$Site.Transect[data$Date == as.Date("2018-05-23") &
                   data$Lake == "Five Island" &
                   data$Site.Transect == "2"] <- "EF.Comp.2"
data$Site.Transect[data$Date == as.Date("2018-05-23") &
                   data$Lake == "Five Island" &
                   data$Site.Transect == "3"] <- "EF.Comp.3"
data$Site.Transect[data$Date == as.Date("2018-05-23") &
                   data$Lake == "Five Island" &
                   data$Site.Transect == "4"] <- "EF.Comp.4"
data$Site.Transect[data$Date == as.Date("2018-05-23") &
                   data$Lake == "Five Island" &
                   data$Site.Transect == "5"] <- "EF.Comp.5"
# Silver spring EF
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "1"] <- "EF.Comp.1"
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "2"] <- "EF.Comp.2"
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "3"] <- "EF.Comp.3"
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "4"] <- "EF.Comp.4"
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "5"] <- "EF.Comp.5"
data$Site.Transect[data$Date == as.Date("2018-05-18") &
                   data$Lake == "Silver" &
                   data$Site.Transect == "6"] <- "EF.Comp.6"
# Storm Spring EF
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "1"] <- "EF.Comp.1"
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "2"] <- "EF.Comp.2"
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "4"] <- "EF.Comp.4"
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "5"] <- "EF.Comp.5"
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "7"] <- "EF.Comp.7"
data$Site.Transect[data$Date == as.Date("2018-05-21") &
                   data$Lake == "Storm" &
                   data$Site.Transect == "9"] <- "EF.Comp.9"
# See what's up now, dog.

# Fyke should take care of itself.
# I am really, really questioning if life is worth it.

# which means I'm not happy where I am at.
# which means I quit? I should get [[[[more]]]] help.

########################################################################################################




```

```{r}

# This code is to overwrite this file name already in directory
# Should re-write this line later for 17 and 18 data in new file name
#
#
write.csv(data, "2018_COC_BIB_CMR_Data.csv", row.names = F)
```



## Adding in 2017 data:
Each of four lakes where fish were tagged in 2017 (Center, N&S Twin, Blue) have separate data files, if I'm not mistaken. Now that I have my personal PC back (12/10/2018) I am more familiar with the filing and where to get the best data files based on most updated and whatnot. 

Feels good to be back.

Got to load in 4 lakes of data,
Change column formats to match df "data" in chunk above,
append each to data frame,
write new CSV with consolidated data in long format


__Blue Lake. Because alphabet__
```{r}
# Also, this one is gonna suck. I didn't have much foresight when it came to using this data later :(
blue.df <- read.csv("2017 BL Carp Tag Data.csv")
str(blue.df)

# Date format
levels(blue.df$Date)
blue.df$Date<-as.Date(blue.df$Date, format = "%m/%d/%Y")
str(blue.df$Date)

# Lake is good
summary(blue.df$Lake)

# Tag number: 
summary(blue.df$TAG.NUMBER)
blue.df$TAG.NUMBER[blue.df$TAG.NUMBER == ""] <- NA # set non-tagged fish to NA
summary(blue.df$TAG.NUMBER)
blue.df$TAG.NUMBER<-droplevels(blue.df$TAG.NUMBER)
# already a factor and 7 recaps all year look OK

# Species
#
summary(blue.df$Species)
blue.df$Species<-as.character(blue.df$Species)
blue.df$Species[blue.df$Species == "BLCR"] <- "BLC" # Black Crappie
blue.df$Species[blue.df$Species == "BBH"] <- "BLB" # Black Bullhead
blue.df$Species<-as.factor(blue.df$Species)
blue.df$Species<-droplevels(blue.df$Species)
summary(blue.df$Species) 


# length
summary(blue.df$Length..mm.)
summary(blue.df$Length..mm.[blue.df$Species == "COC"])
# looks good

# Weight
summary(blue.df$Weight..g.)
summary(blue.df$Weight..g.[blue.df$Species== "COC"])
# Looks OK (again most of these were checked last year)

# recaps
summary(blue.df$Recap)
sum(!is.na(blue.df$Recap)) # well shoot missing 2 recaps
# NopE! turns out miskeys for tags 560 and 561 
# (supposed to be 590 and 591 respectively)


# Tag Loss
summary(blue.df$Tag.Loss)
sum(!is.na(blue.df$Tag.Loss))

# Skipping Notes and Gear. Not relevant to this dataset
summary(blue.df$Gear)

# Site/Transect
summary(blue.df$Site.Transect)
# so they had std runs 1-4 then just willy nilly the rest of the year up to trans39 + aging
# 40 runs total.

# skipping to On.Time..s.
summary(blue.df$On.Time..s.)
# seems OK after checking min and max

# secchi and weather never measured.


# remove 3 extra columns at end
lapply(data,class)
lapply(blue.df,class)

# shoot homie. NA level generated for an invalid factor level. 
# seems related to mis-matched classes of columns (e.g., numeric vs. integer)

blue.df$Length..mm.<-as.numeric(blue.df$Length..mm.)
blue.df$Weight..g.<-as.numeric(blue.df$Weight..g.)
blue.df$Recap<-as.factor(blue.df$Recap)
blue.df$Tag.Loss<-as.factor(blue.df$Tag.Loss)
blue.df$On.Time..s.<-as.numeric(blue.df$On.Time..s.)
blue.df$Water.Temp..C.<-as.numeric(blue.df$Water.Temp..C.)
blue.df$Secchi..cm.<-as.numeric(blue.df$Secchi..cm.)
blue.df$Weather<-as.factor(blue.df$Weather)
blue.df$Year<-as.factor("2017")

data<-rbind(data,blue.df)
```

__Twin Lakes__
```{r}
twin.df <- read.csv("2017 TwinLakes Tag Data.csv", header = TRUE)
str(twin.df)
twin.df$Date<-twin.df$ï..Date



# Date format
levels(twin.df$Date )
twin.df$Date<-as.Date(twin.df$Date, format = "%m/%d/%Y")
str(twin.df$Date)

# Lake is good
summary(twin.df$Lake)
twin.df$Lake<-as.character(twin.df$Lake)
twin.df$Lake[twin.df$Lake == "N Twin"] <- "N. Twin"
twin.df$Lake[twin.df$Lake == "S Twin"] <- "S. Twin"
twin.df$Lake<-as.factor(twin.df$Lake)
twin.df$Lake<-droplevels(twin.df$Lake)
summary(twin.df$Lake)

# Tag number: 
summary(twin.df$TAG.NUMBER)
twin.df$TAG.NUMBER[twin.df$TAG.NUMBER == ""] <- NA # set non-tagged fish to NA
summary(twin.df$TAG.NUMBER)
#twin.df$TAG.NUMBER<-droplevels(twin.df$TAG.NUMBER)
# already a factor and 7 recaps all year look OK

# Species
#
summary(twin.df$Species)
# no all-species samples in these lakes in 2017


# length
summary(twin.df$Length..mm.)
summary(twin.df$Length..mm.[twin.df$Species == "COC"])
summary(twin.df$Length..mm.[twin.df$Species == "BIB"])

# Weight
summary(twin.df$Weight..g.)
summary(twin.df$Weight..g.[twin.df$Species== "COC"])
summary(twin.df$Weight..g.[twin.df$Species== "BIB"])
# Looks OK (again most of these were checked last year)

# recaps
summary(twin.df$Recap)
twin.df$Recap<-as.character(twin.df$Recap)
twin.df$Recap[twin.df$Recap == "R"] <- "1"
twin.df$Recap<-as.factor(twin.df$Recap)
twin.df$Recap<-droplevels(twin.df$Recap)
sum(!is.na(twin.df$Recap)) 


# Tag Loss
summary(twin.df$Tag.Loss)
sum(!is.na(twin.df$Tag.Loss)) # 3 2017 tag losses

# Skipping Notes and Gear. Not relevant to this dataset


# Site/Transect
summary(twin.df$Site.Transect)
# over 90 runs for both lakes. This is unique by each run regardless of lake. 
# Since Black Hawk office compiled lake data I found it easier to make each run a new number
# no standard transect runs nor all-species runs in the twin lakes 2017

# skipping to On.Time..s.
summary(twin.df$On.Time..s.)
# seems OK after checking min and max

# secchi and weather never measured.

# year
twin.df$Year <- "2017"

# shoot homie. NA level generated for an invalid factor level. 
# seems related to mis-matched classes of columns (e.g., numeric vs. integer)




twin.df$Weight..g.<-as.numeric(twin.df$Weight..g.)
twin.df$Tag.Loss<-as.factor(twin.df$Tag.Loss)
twin.df$On.Time..s.<-as.numeric(twin.df$On.Time..s.)
twin.df[,13]<-as.factor(twin.df[,13])
twin.df[,14]<-as.factor(twin.df[,14])
twin.df[,15]<-as.factor(twin.df[,15])
twin.df[,16]<-as.factor(twin.df[,16])
twin.df$Secchi..cm.<-as.numeric(twin.df$Secchi..cm.)
twin.df$Weather<-as.factor(twin.df$Weather)
twin.df$Year<-as.factor(twin.df$Year)

twin.df<-twin.df[,-1]

data<-rbind(data,twin.df)

# dang that's some nice work. Up to 25k observations
```

__Center Lake!__
Last one here. Got to add Spirit Lake data to this format for a truly compiled data collection of all carp and buffalo tagging data up to 12/11/2018.

```{r}
center.df <- read.csv("2017 CenterLake Tag Data.csv", header = TRUE)
str(center.df)

center.df$Date<-center.df$ï..Date



# Date format
levels(center.df$Date )
center.df$Date<-as.Date(center.df$Date, format = "%m/%d/%Y")
str(center.df$Date)

# Lake is good
summary(center.df$Lake)


# Tag number: 
summary(center.df$TAG.NUMBER)
center.df$TAG.NUMBER[center.df$TAG.NUMBER == "."] <- NA # set non-tagged fish to NA
summary(center.df$TAG.NUMBER)


# Species
#
summary(center.df$Species)
center.df$Species<-as.character(center.df$Species)
center.df$Species[center.df$Species == "CAP"] <-"COC"
center.df$Species<-as.factor(center.df$Species)
summary(center.df$Species)
# no all-species samples in these lakes in 2017


# length
# Yeah awesome spirit lake crew for entering "." in empty cells
summary(center.df$Length..mm.)
length(center.df$Length..mm.[center.df$Length..mm. == "."])
center.df$Length..mm.[center.df$Length..mm. == "."]<- NA
center.df$Length..mm.<-as.numeric(as.character(center.df$Length..mm.))
summary(center.df$Length..mm.[center.df$Species == "COC"])
summary(center.df$Length..mm.[center.df$Species == "BIB"])

# Weight
summary(center.df$Weight..g.)
length(center.df$Weight..g.[center.df$Weight..g. == "."])
center.df$Weight..g.[center.df$Weight..g. == "."]<- NA
center.df$Weight..g.<-as.numeric(as.character(center.df$Weight..g.))
summary(center.df$Weight..g.[center.df$Species== "COC"])
summary(center.df$Weight..g.[center.df$Species== "BIB"])
# Looks OK (again most of these were checked last year)

# recaps
summary(center.df$Recap)
center.df$Recap<-as.character(center.df$Recap)
center.df$Recap[center.df$Recap == "."] <- NA
center.df$Recap[center.df$Recap == ""] <- NA
center.df$Recap<-as.factor(center.df$Recap)
center.df$Recap<-droplevels(center.df$Recap)
sum(!is.na(center.df$Recap)) 
summary(center.df$Recap)


# Tag Loss
summary(center.df$Tag.Loss)
sum(!is.na(center.df$Tag.Loss)) # 42 2017 tag losses

# gear
summary(center.df$Gear)
summary(data$Gear)
center.df$Gear<-as.character(center.df$Gear)
center.df$Gear[center.df$Gear == "EF"]<- "Electroshocking"
center.df$Gear[center.df$Gear == "CF Seine"]<- "Commercial Seine"
center.df$Gear<-as.factor(center.df$Gear)
summary(center.df$Gear)


# Site/Transect
summary(center.df$Site.Transect)
# no standard transect runs nor all-species runs in center lake 2017

# skipping to On.Time..s.
summary(center.df$On.Time..s.)
center.df$On.Time..s.<-as.character(center.df$On.Time..s.)
center.df$On.Time..s.[center.df$On.Time..s. =="."] <- NA
center.df$On.Time..s.<-as.numeric(center.df$On.Time..s.)
summary(center.df$On.Time..s.)
# seems OK after checking min and max

# water temp
summary(center.df$Water.Temp..C.)
# all good and numeric

# secchi
center.df$Secchi..cm.<-as.numeric(as.character(center.df$Secchi..cm.))

# weather is good

# year
center.df$Year<-as.factor(as.character(center.df$Year))

# see where we are at:
str(center.df)

# correct classes of columns:
center.df$Tag.Loss<-as.factor(as.character(center.df$Tag.Loss))
center.df[,14]<-as.factor(center.df[,14])
center.df[,16]<-as.factor(center.df[,16])

# remove junky year column
center.df<-center.df[,-1]

data<-rbind(data,center.df)



```

Holy cow! As of 12/11/2018 (1/4/2019) this data frame is now 26,615 observations long. Time to celebrate with more data checking. for now let's write a new csv in a new chunk to have the 2017 and 2018 data separate from the other 2018 only data

```{r}
write.csv(data,"2018 and 2017 Compiled Data.csv")
```





# Silver Lake Carp Summary
This was a quick estimate for MJW specific request. Still some good code here to plot length-frequency histograms and CPUE in June.
Carp only.

```{r}
# Length Frequency Histogram of COC in Silver Lake, IA

silver.coc<-subset(data, Species == "COC")
silver.coc<-subset(silver.coc, Lake == "Silver")
silver.coc<-subset(silver.coc, !is.na(silver.coc$Length..mm.))
head(silver.coc)
str(silver.coc)

library(ggplot2)

ggplot(silver.coc, aes(x=Length..mm., fill=Species))+
  geom_histogram(binwidth=10,colour="black",position="dodge")+
  scale_x_continuous(limits = c(0,1000))+
  scale_y_continuous(limits = c(0,60))

library(rcompanion)
groupwiseMean(Length..mm.~1,
              data = silver.coc,
              conf = 0.95,
              digits = 3)

sem<-sd(silver.coc$Length..mm.)/sqrt(length(silver.coc$Length..mm.))

# cpue for coc in silver lake in june
#
# CarpCPUE - n fish per hour electrofishing
#
summary(silver.coc)
# on-time in seconds to Effort in hours
       # Need to remove seine hauls and NAs
str(silver.coc$On.Time..s.)
silver.coc$On.Time..s.[silver.coc$On.Time..s. == ""] <- NA
silver.coc$On.Time..s.[silver.coc$On.Time..s. == "453/488"] <- "941"
silver.coc$On.Time..s.<-as.numeric(as.character(silver.coc$On.Time..s.))
silver.coc$Effort<-silver.coc$On.Time..s./3600



# now to paste date, then lake then run
silver.coc$DR<-factor(paste(silver.coc$Date,silver.coc$Site.Transect))

# Now to divide the count of captured fish by the mean effort per DLR
coc.cpue<-aggregate(Effort~DR, data = silver.coc, mean, na.action=NULL)
silver.coc$dummy<-1
coc.n<-aggregate(dummy~DR, data = silver.coc, sum)
coc.cpue$n<-coc.n$dummy
coc.cpue$cpue<-coc.cpue$n/coc.cpue$Effort
coc.cpue<-coc.cpue[!is.na(coc.cpue$cpue),]

june.coc.cpue<-coc.cpue[c(12:44),]
groupwiseMean(cpue~1,
              data=june.coc.cpue,
              conf = 0.95)

```

