---
title: "Initial Twin Lake Plots"
author: "Martin A. Simonson"
date: "September 14, 2017"
output: pdf_document
---
```{r,include=F}
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

packages(vegan)
packages(ggplot2)
packages(reshape2)
packages(nnet)
packages(multcomp)
packages(plyr)
packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(gridExtra)

```
# reading in data and splitting up lakes and species
```{r}
df <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/TwinLakesLength2017.csv")

summary(df)
# Keeping units of data sheets, although I would prefer cm to inches

NT.B<-subset(df,df$Lake=="N Twin" & df$Species == "BIB")
NT.C<-subset(df,df$Lake=="N Twin" & df$Species == "COC")
ST.B<-subset(df,df$Lake=="S Twin" & df$Species == "BIB")
ST.C<-subset(df,df$Lake=="S Twin" & df$Species == "COC")
```
# Length frequency histograms: two species of fish, two lakes.

```{r}


p1<-ggplot(NT.C,aes(x=Length..inches.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "goldenrod") + 
  labs(title="Length Frequency for Carp: North Twin Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

p2<-ggplot(NT.B,aes(x=Length..inches.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "cornflowerblue") + 
  labs(title="Length Frequency for Buffalo: North Twin Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

p3<-ggplot(ST.C,aes(x=Length..inches.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "goldenrod") + 
  labs(title="Length Frequency for Carp: South Twin Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

p4<-ggplot(ST.B,aes(x=Length..inches.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "cornflowerblue") + 
  labs(title="Length Frequency for Buffalo: South Twin Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

grid.arrange(p1,p2,p3,p4, nrow=2, ncol=2)
```
# reading in data and splitting up species
```{r}
df <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/CenterLake2017.csv",header=TRUE)

summary(df)
df$TL.in.[df$TL.in.=="."]<-NA
df$TL.in.[df$TL.in.=="0.0"]<-NA
df$TL.in.<-as.numeric(as.character(df$TL.in.))
# Keeping units of data sheets, although I would prefer cm to inches

CL.B<-subset(df, df$Spp == "BIB", na.rm=TRUE)
CL.C<-subset(df, df$Spp == "CAP", na.rm=TRUE)

summary(CL.B)
summary(CL.C)
```
# Length frequency histograms: two species of fish, one lake.

```{r}

p1<-ggplot(CL.B,aes(x=TL.in.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "cornflowerblue") + 
  labs(title="Length Frequency for Buffalo: Blue Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

CL.C$TL.in.<-as.numeric(CL.C$TL.in.)
p2<-ggplot(CL.C,aes(x=TL.in.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "goldenrod") + 
  labs(title="Length Frequency for Carp: Center Lake",
       x="Length (inches)",
       y="Frequency")+
  ylim(c(0,150))

grid.arrange(p1,p2,ncol=2,nrow=1)
```

# Length frequency histograms: One species of fish, one lake

```{r}
df <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/BL_Carp_Lengths.csv",header=TRUE)

summary(df)


#qplot(df$Length..in.,
#      geom="histogram",
#      main="Length Freqency for Carp: Blue Lake",
#      xlab="Length (inches)",
#      fill=I("goldenrod"),
#      col=I("red"),
#      alpha=0.5,
#      binwidth = 0.5,
#      ylim = c(0,125), xlim = c(8.5,27.5))

ggplot(df,aes(x=Length..in.)) + 
  geom_histogram(binwidth = 0.5, col = "black", fill = "goldenrod") + 
  labs(title="Length Frequency for Carp: Blue Lake",
       x="Length (inches)",
       y="Frequency") +
  ylim(c(0,150))

```


# As of Jan 12 2018

Tasked with creating 1 figure with 
- four length-freq plots (one for each lake)
- both species in each panel
- same y-axis limits

I want to have this in "long" form, where the spp is a column and the lengths are the other column. 

### Do this for 4 lakes and then do a grid.arrange?

```{r,}
#Example from stackoverflow and RPubs
# http://rpubs.com/sjackman/grid_arrange_shared_legend
library(ggplot2)
library(gridExtra)
grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    grid.arrange(
        do.call(arrangeGrob, lapply(plots, function(x)
            x + theme(legend.position="none"))),
        legend,
        ncol = 1,
        heights = unit.c(unit(1, "npc") - lheight, lheight))
}

# So this function didn't work for me:

# Error in g[[which(sapply(g, function(x) x$name) == "guide-box")]] : 
#  attempt to select less than one element in get1index
# In addition: Warning message:
# Removed 2 rows containing missing values (geom_bar). 


```

```{r,}
########################################################
#
#
# Twin Lakes
#
#
########################################################

df <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/TwinLakesLength2017.csv")

summary(df)
# Keeping units of data sheets, although I would prefer cm to inches

NT.B<-subset(df,df$Lake=="N Twin" & df$Species == "BIB")
NT.C<-subset(df,df$Lake=="N Twin" & df$Species == "COC")
ST.B<-subset(df,df$Lake=="S Twin" & df$Species == "BIB")
ST.C<-subset(df,df$Lake=="S Twin" & df$Species == "COC")

North.B<-ggplot(NT.B, aes(x=Length..inches., fill=Species)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(
                      breaks=1:2,
                      values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
  scale_x_continuous(limits = c(0,35)) +
  labs(x="Length (Inches)", y="Frequency",title="North Twin Lake: Buffalo")

North.C<-ggplot(NT.C, aes(x=Length..inches., fill=Species)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(
                      breaks=1:2,
                      values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
  scale_x_continuous(limits = c(0,35))+
  labs(x="Length (Inches)", y="Frequency",title="North Twin Lake: Carp")

South.B<-ggplot(ST.B, aes(x=Length..inches., fill=Species)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(breaks=1:2, 
                    values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
    scale_x_continuous(limits = c(0,35)) +
  labs(x="Length (Inches)", y="Frequency",title="South Twin Lake: Buffalo")

South.C<-ggplot(ST.B, aes(x=Length..inches., fill=Species)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(breaks=1:2, 
                    values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
    scale_x_continuous(limits = c(0,35)) +
  labs(x="Length (Inches)", y="Frequency",title="South Twin Lake: Carp")

#####################################################
#
#
# Center Lake
#
#
#####################################################
CL <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/CenterLake2017.csv",header=TRUE)

summary(CL)
CL$TL.in.[CL$TL.in.=="."]<-NA
CL$TL.in.[CL$TL.in.=="0.0"]<-NA
CL$TL.in.<-as.numeric(as.character(CL$TL.in.))
# Keeping units of data sheets, although I would prefer cm to inches
CL$Spp<-as.character(CL$Spp)
CL$Spp[CL$Spp == "CAP"]<-"COC"
CL$Spp<-as.factor(CL$Spp)
CL<-na.omit(CL)
summary(CL)

CL.B<-subset(CL,CL$Spp == "BIB")
CL.C<-subset(CL,CL$Spp == "COC")


Center.B<-ggplot(CL.B, aes(x=TL.in., fill=Spp)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(breaks=1:2, 
                    values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
  scale_x_continuous(limits = c(0,35))+
  labs(x="Length (Inches)", y="Frequency",title="Center Lake: Buffalo")

Center.C<-ggplot(CL.C, aes(x=TL.in., fill=Spp)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(breaks=1:2, 
                    values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
  scale_x_continuous(limits = c(0,35))+
  labs(x="Length (Inches)", y="Frequency",title="Center Lake: Carp")


#############################################
#
# Blue Lake
#
#############################################
BL <- read.csv("C:\\Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/BL_Carp_Lengths.csv",header=TRUE)
BL$Spp<-"COC"
BL$Spp<-as.factor(BL$Spp)
summary(BL)

Blue<-ggplot(BL, aes(x=Length..in., fill=Spp)) +
  geom_histogram(binwidth=0.5, colour="black", position="dodge") +
  scale_fill_manual(breaks=1:2, 
                    values=c("cornflowerblue","goldenrod")) +
  scale_y_continuous(limits = c(0, 130)) +
  scale_x_continuous(limits = c(0,35))+
  labs(x="Length (Inches)", y="Frequency",title="Blue Lake: Carp")

grid.arrange(Center.C,Center.B,
             North.C,North.B,
             South.C,South.B,
             Blue,
             ncol=2,nrow=4)




```
