---
title: "Iowa AFS 2020"
author: "Martin Simonson"
date: "2/28/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r, include=FALSE}
# front matter
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#packages(rcompanion)
packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
packages(GISTools)
#packages(maps)
#packages(mapproj)
packages(ggpmisc)
packages(lubridate)
```

# objective 
coallate the various code from Chapter 2 RMD working file into the "final products" needed for the Iowa AFS 2020 presentation. 

Potentially ditching CMR for this talk, focusing on 

- Survival (CJS models maybe, but MJW will be picky) [ same approach as Wr - differences ]

- Abundance/Biomass Density Changes

- Growth in Length

- Size Structure (how it changes after harvest) [ same approach as Wr - differences]

- Fecundity (recruitment of 2 yo and 3 yo fish)

## Another way to look at this is tax exploitation:

X axis is tax returns (exploitation) from 0 to 100%
Y is % change in population, from -100 to 100
One-one line from middle y axis (0) to 100% on x axis. Anything above that is compensatory response!

_problem_ is that the change in population is year-year, so we only have 2 data points for Center and N Twin and only have 1 data point for Silver, Storm. Plus, how the F do I handle the change in the variation around BMD from year to year? Top CI in 2018 minus top CI in 2017???

# length frequenccies


look at size structure of populations
```{r}
df<-read.csv("2017-2020 Compiled Data.csv", header = T)

df$Length.in<-df$Length..mm./25.4
df<-droplevels(subset(df, Gear != "Fyke"))
df$Gear<-as.character(df$Gear)
df$Gear[df$Gear == "Seine"] <- "Commercial Seine"
df$Gear<-as.factor(df$Gear)

df<-droplevels(subset(df, !is.na(Length..mm.) & Species == "COC" |
                          !is.na(Length..mm.) & Species == "BIB"))

df$Species<-as.character(df$Species)
df$Species[df$Species == "COC"] <- "Common Carp"
df$Species[df$Species == "BIB"] <- "Bigmouth Buffalo"
df$Species<-as.factor(df$Species)

df.b<-droplevels(subset(df, Species == "Bigmouth Buffalo" & Lake != "Blue"))
df.c<-droplevels(subset(df, Species == "Common Carp" & Lake != "Blue"))

buffalo.lf<-ggplot(df.b, aes(x=Length.in, fill = Gear))+
                         geom_histogram(binwidth = 1,
                                        colour = "black",
                                        position = "stack")+
                         coord_cartesian(xlim = c(0,45))+
                         facet_grid(Year ~ Lake)+
                         theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = c(0.91,0.8),
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))+
                         scale_fill_grey(start = 0.1, end = 0.9, aesthetics = "fill")+
                         labs(title = "Bigmouth Buffalo Length Frequencies 2017-2019",
                              x = "Length (inches)",
                              y = "Frequency")
buffalo.lf # ok that exists.

carp.lf<-ggplot(df.c, aes(x=Length.in, fill = Gear))+
                         geom_histogram(binwidth = 1,
                                        colour = "black",
                                        position = "stack")+
                         coord_cartesian(xlim = c(0,45))+
                         facet_grid(Year ~ Lake)+
                         theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = c(0.91,0.8),
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))+
                         scale_fill_grey(start = 0.1, end = 0.9, aesthetics = "fill")+
                         labs(title = "Common Carp Length Frequencies 2017-2019",
                              x = "Length (inches)",
                              y = "Frequency")
carp.lf # ok that exists.

```

# and now what?

The tag exploitation plot is a good idea actually.

Let's talk out the format of the data:

Lake    | Year Range    | Species    | % exploitation (% tags returned)     | % change in population (lbs/acre)

s Twin  | 2017 - 2018   | BIB        | 0 (no harvest)                       | ??
s Twin  | 2017 - 2018   | COC        | 0 (no harvest)                       | ??

- I should have code written up for the tag exploitation

- % change in population will be interesting. Go with biomass densities. 

```{r}
bmd17<-read.csv("2017_BiomassDensityEstimates.csv",header=T)
head(bmd17)
bmd18<-read.csv("2018_BiomassDensityEstimates.csv",header=T)
head(bmd18)
bmd19<-read.csv("2019_BiomassDensityEstimates.csv",header=T)
head(bmd19)

bmd<-rbind(bmd17,bmd18,bmd19)
levels(bmd$Lake)
bmd$Lake[bmd$Lake == "5 Island"] <- "Five Island"


# create new DF 
Lake <- c("Blue","Blue",
          "Center","Center","Center","Center",
          "Five Island","Five Island",
          "N Twin","N Twin","N Twin","N Twin",
          "S Twin","S Twin","S Twin","S Twin",
          "Silver","Silver",
          "Storm","Storm")
Species<-c("Common Carp","Common Carp", # two years blue
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species center
           "Common Carp","Bigmouth Buffalo", # one year two species five island
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species north twin
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species south twin
           "Common Carp","Bigmouth Buffalo", # one year two species silver
           "Common Carp","Bigmouth Buffalo") # one year two species storm
Year.Range<-c("2017-2018","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2018-2019","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2018-2019","2018-2019",
              "2018-2019","2018-2019")

bmd.change<-data.frame(cbind(Lake,Species,Year.Range))
head(bmd.change)

# now need to set up the change in biomass densities. The Response variable
## as a percent
### means the biomass density in year two divided by the biomass density in year one times 100
str(droplevels(bmd))
delta.bmd<-as.character(rep(NA,20))
delta.bmd[[1]]<-bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2018] / 
                bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2017] * 100
delta.bmd[[2]]<-bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2019] / 
                bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2018] * 100
delta.bmd[[1]]<-bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2018] / 
                bmd$BMD[bmd$Lake == "Blue" & bmd$Species == "Common Carp" & bmd$Year == 2017] * 100

```

```{r}
##### 
# Front Matter
setwd("~/BMB-COC")

#2019 tagging data
nineteen<-read.csv("2019 Carp and Buffalo Data Entry.csv")
str(nineteen)
levels(nineteen$Lake)
nineteen$Lake[nineteen$Lake == "N. Twin"] <- "N Twin"
nineteen<-droplevels(subset(nineteen, Species == "COC" |
                                      Species == "BIB"))
Silver19<-droplevels(subset(nineteen, Lake == "Silver"))
NTwin19<-droplevels(subset(nineteen, Lake == "N Twin"))

# 2019 harvest data
harvest<-read.csv("2019  Harvest Tag Recovery.csv")
str(harvest)
levels(harvest$Lake)
harvest$Lake[harvest$Lake == "N. Twin"] <- "N Twin"
SilverHarvest<-droplevels(subset(harvest, Lake == "Silver"))
NTwinHarvest<-droplevels(subset(harvest, Lake == "N Twin"))


# Silver first, most straightforward
## Find # of tags from harvest that came from each species in Silver Lake
## setdiff(x,y) provides values in X but not in Y, where X and Y can be vectors or data frames

length(setdiff(SilverHarvest$TAG.NUM, Silver19$TAG.NUM))
# 83 tags recovered were NOT from 2019 tagging events
tags19<-length(SilverHarvest$TAG.NUM)-83
# so 150 tags recovered ARE in 2019 tagging events


# First crack at exploitation:
#### Denominator is total number of 2019 tags deployed for that species
#### Numerator is total number of 2019 tags recovered by each species.
silver.buff.exploit<-(length(SilverHarvest$TAG.NUM)-length(setdiff(SilverHarvest$TAG.NUM,Silver19$TAG.NUM[Silver19$Species=="BIB"]))) /
                      (length(Silver19$TAG.NUM[Silver19$Species == "BIB"]))
silver.carp.exploit<-(length(SilverHarvest$TAG.NUM)-length(setdiff(SilverHarvest$TAG.NUM,Silver19$TAG.NUM[Silver19$Species=="COC"]))) /
                      (length(Silver19$TAG.NUM[Silver19$Species == "COC"]))
# Is overall exploitation the same as the sum of both exploitation rates?
silver.exploit<-(length(SilverHarvest$TAG.NUM)-length(setdiff(SilverHarvest$TAG.NUM,Silver19$TAG.NUM))) /
                  (length(Silver19$TAG.NUM)) # not currently. gotta check with someone else!

silver<-data.frame(silver.exploit,silver.buff.exploit,silver.carp.exploit)
silver


#### Try with N Twin now
NTwin.buff.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM[NTwin19$Species=="BIB"]))) /
  (length(NTwin19$TAG.NUM[NTwin19$Species == "BIB"]))
NTwin.carp.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM[NTwin19$Species=="COC"]))) /
  (length(NTwin19$TAG.NUM[NTwin19$Species == "COC"]))
# Is overall exploitation the same as the sum of both exploitation rates?
NTwin.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM))) /
  (length(NTwin19$TAG.NUM)) # not currently. gotta check with someone else!

NTWIN<-data.frame(NTwin.exploit,NTwin.buff.exploit,NTwin.carp.exploit)
NTWIN


# remember that final tagging day after population was already harvested then harvested again.(11/10/19)
### Could exclude that day up front and not use it in exploitation estimates?
## Shoot for my own consience I'll try to see how different N Twin is without that final tag day

levels(NTwin19$Date)
NTwin19.2<-droplevels(subset(NTwin19, Date != "11/10/19"))
# Same as before
NTwin.buff.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM[NTwin19.2$Species=="BIB"]))) /
  (length(NTwin19$TAG.NUM[NTwin19$Species == "BIB"]))
NTwin.carp.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM[NTwin19.2$Species=="COC"]))) /
  (length(NTwin19$TAG.NUM[NTwin19$Species == "COC"]))
# Is overall exploitation the same as the sum of both exploitation rates? Not so below.
NTwin.exploit<-(length(NTwinHarvest$TAG.NUM)-length(setdiff(NTwinHarvest$TAG.NUM,NTwin19$TAG.NUM))) /
  (length(NTwin19$TAG.NUM)) # not currently. gotta check with someone else!

NTWIN<-data.frame(NTwin.exploit,NTwin.buff.exploit,NTwin.carp.exploit)
NTWIN
# hardly changed. Good to see here.

# checking sum?
(NTWIN$NTwin.buff.exploit+NTWIN$NTwin.carp.exploit)
# checking average
(NTWIN$NTwin.buff.exploit+NTWIN$NTwin.carp.exploit)/2
# hell check the difference

(NTWIN$NTwin.buff.exploit-NTWIN$NTwin.carp.exploit)
# Closer... ask brandon upstairs
```

