---
title: "Iowa AFS 2020"
author: "Martin Simonson"
date: "2/28/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r, include=FALSE}
# front matter
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#packages(rcompanion)
packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
packages(GISTools)
#packages(maps)
#packages(mapproj)
packages(ggpmisc)
packages(lubridate)
```

# objective 
coallate the various code from Chapter 2 RMD working file into the "final products" needed for the Iowa AFS 2020 presentation. 

Potentially ditching CMR for this talk, focusing on 

- Survival (CJS models maybe, but MJW will be picky) [ same approach as Wr - differences ]

- Abundance/Biomass Density Changes

- Growth in Length

- Size Structure (how it changes after harvest) [ same approach as Wr - differences]

- Fecundity (recruitment of 2 yo and 3 yo fish)

## Another way to look at this is tax exploitation:

X axis is tax returns (exploitation) from 0 to 100%
Y is % change in population, from -100 to 100
One-one line from middle y axis (0) to 100% on x axis. Anything above that is compensatory response!

_problem_ is that the change in population is year-year, so we only have 2 data points for Center and N Twin and only have 1 data point for Silver, Storm. Plus, how the F do I handle the change in the variation around BMD from year to year? Top CI in 2018 minus top CI in 2017???

# length frequenccies


look at size structure of populations
```{r}
df<-read.csv("2017-2020 Compiled Data.csv", header = T)

df$Length.in<-df$Length..mm./25.4
df<-droplevels(subset(df, Gear != "Fyke"))
df$Gear<-as.character(df$Gear)
df$Gear[df$Gear == "Seine"] <- "Commercial Seine"
df$Gear<-as.factor(df$Gear)

df<-droplevels(subset(df, !is.na(Length..mm.) & Species == "COC" |
                          !is.na(Length..mm.) & Species == "BIB"))

df$Species<-as.character(df$Species)
df$Species[df$Species == "COC"] <- "Common Carp"
df$Species[df$Species == "BIB"] <- "Bigmouth Buffalo"
df$Species<-as.factor(df$Species)

df.b<-droplevels(subset(df, Species == "Bigmouth Buffalo" & Lake != "Blue"))
df.c<-droplevels(subset(df, Species == "Common Carp" & Lake != "Blue"))

buffalo.lf<-ggplot(df.b, aes(x=Length.in, fill = Gear))+
                         geom_histogram(binwidth = 1,
                                        colour = "black",
                                        position = "stack")+
                         coord_cartesian(xlim = c(0,45))+
                         facet_grid(Year ~ Lake)+
                         theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = c(0.91,0.8),
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))+
                         scale_fill_grey(start = 0.1, end = 0.9, aesthetics = "fill")+
                         labs(title = "Bigmouth Buffalo Length Frequencies 2017-2019",
                              x = "Length (inches)",
                              y = "Frequency")
buffalo.lf # ok that exists.

carp.lf<-ggplot(df.c, aes(x=Length.in, fill = Gear))+
                         geom_histogram(binwidth = 1,
                                        colour = "black",
                                        position = "stack")+
                         coord_cartesian(xlim = c(0,45))+
                         facet_grid(Year ~ Lake)+
                         theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = c(0.91,0.8),
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))+
                         scale_fill_grey(start = 0.1, end = 0.9, aesthetics = "fill")+
                         labs(title = "Common Carp Length Frequencies 2017-2019",
                              x = "Length (inches)",
                              y = "Frequency")
carp.lf # ok that exists.

```

# and now what?

The tag exploitation plot is a good idea actually.

Let's talk out the format of the data:

Lake    | Year Range    | Species    | % exploitation (% tags returned)     | % change in population (lbs/acre)

s Twin  | 2017 - 2018   | BIB        | 0 (no harvest)                       | ??
s Twin  | 2017 - 2018   | COC        | 0 (no harvest)                       | ??

- I should have code written up for the tag exploitation

- % change in population will be interesting. Go with biomass densities. 

```{r}
bmd17<-read.csv("2017_BiomassDensityEstimates.csv",header=T)
head(bmd17)
bmd18<-read.csv("2018_BiomassDensityEstimates.csv",header=T)
head(bmd18)
bmd19<-read.csv("2019_BiomassDensityEstimates.csv",header=T)
head(bmd19)


applejacks<-read.csv("BMD.17.18.19.csv",header=T)


bmd<-rbind(bmd17,bmd18,bmd19)
levels(bmd$Lake)
bmd$Lake[bmd$Lake == "5 Island"] <- "Five Island"


# create new DF 
Lake <- as.factor(as.character(c("Blue","Blue",
          "Center","Center","Center","Center",
          "Five Island","Five Island",
          "N Twin","N Twin","N Twin","N Twin",
          "S Twin","S Twin","S Twin","S Twin",
          "Silver","Silver",
          "Storm","Storm")))
Species<-as.factor(as.character(c("Common Carp","Common Carp", # two years blue
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species center
           "Common Carp","Bigmouth Buffalo", # one year two species five island
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species north twin
           "Common Carp","Bigmouth Buffalo","Common Carp","Bigmouth Buffalo", # two years two species south twin
           "Common Carp","Bigmouth Buffalo", # one year two species silver
           "Common Carp","Bigmouth Buffalo"))) # one year two species storm
Year.Range<-as.factor(as.character(c("2017-2018","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2018-2019","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2017-2018","2017-2018","2018-2019","2018-2019",
              "2018-2019","2018-2019",
              "2018-2019","2018-2019")))

bmd.change<-data.frame(Lake,Species,Year.Range)
head(bmd.change)
```


```{r,}
# now need to set up the change in biomass densities. The Response variable
## as a percent
### means the biomass density in year two divided by the biomass density in year one times 100
str(droplevels(bmd))
delta.bmd<-as.character(rep(".",20))
delta.bmd[[1]]<-((bmd$C.BMDensity[bmd$Lake == "Blue" & bmd$Year == 2018] / 
                bmd$C.BMDensity[bmd$Lake == "Blue" & bmd$Year == 2017])-1) * 100
delta.bmd[[2]]<-((bmd$C.BMDensity[bmd$Lake == "Blue" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "Blue" & bmd$Year == 2018])-1) * 100
#
delta.bmd[[3]]<-((bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018] / 
                bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2017])-1) * 100
delta.bmd[[4]]<-((bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018] / 
                bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2017])-1) * 100
delta.bmd[[5]]<-((bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100
delta.bmd[[6]]<-((bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100
#

delta.bmd[[7]]<-((bmd$C.BMDensity[bmd$Lake == "Five Island" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "Five Island" & bmd$Year == 2018])-1) * 100
delta.bmd[[8]]<-((bmd$B.BMDensity[bmd$Lake == "Five Island" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "Five Island" & bmd$Year == 2018])-1) * 100
#

delta.bmd[[9]]<-((bmd$C.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2018] / 
                bmd$C.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2017])-1) * 100
delta.bmd[[10]]<-((bmd$B.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2018] / 
                bmd$B.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2017])-1) * 100
delta.bmd[[11]]<-((bmd$C.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2018])-1) * 100
delta.bmd[[12]]<-((bmd$B.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "N Twin" & bmd$Year == 2018])-1) * 100
#

delta.bmd[[13]]<-((bmd$C.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2018] / 
                bmd$C.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2017])-1) * 100
delta.bmd[[14]]<-((bmd$B.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2018] / 
                bmd$B.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2017])-1) * 100
delta.bmd[[15]]<-((bmd$C.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2018])-1) * 100
delta.bmd[[16]]<-((bmd$B.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "S Twin" & bmd$Year == 2018])-1) * 100
#

delta.bmd[[17]]<-((bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100
delta.bmd[[18]]<-((bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100
#

delta.bmd[[19]]<-((bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$C.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100
delta.bmd[[20]]<-((bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2019] / 
                bmd$B.BMDensity[bmd$Lake == "Center" & bmd$Year == 2018])-1) * 100

bmd.change$delta.bmd<-as.numeric(as.character(delta.bmd))
bmd.change


```

```{r}
##### 
dframe<-read.csv("2017-2020 Compiled Data.csv",header = T)
dframe$Year<-as.factor(dframe$Year)
dframe$TAG.NUMBER<-as.factor(dframe$TAG.NUMBER)
dframe$Date<-as.Date(dframe$Date, "%m/%d/%Y")
dframe$On.Time..s.<-as.numeric(dframe$On.Time..s.)
str(dframe)
# remove all other spp
dframe$TAG.NUMBER[dframe$TAG.NUMBER == ""]<-NA
dframe<-droplevels(subset(dframe, Species == "BIB" |
                                  Species == "COC" ))
# remove all dissected fish and non-tagged fish
dframe<-droplevels(subset(dframe, !is.na(as.numeric(as.character(dframe$TAG.NUMBER)))))
# change five island
dframe$Lake<-as.character(dframe$Lake)
dframe$Lake[dframe$Lake == "5 Island"]<- "Five Island"
dframe$Lake<-as.factor(dframe$Lake)


# 
#Harvested tags recovered
harvest<-read.csv("Harvest Tag Recovery.csv",header=T)
str(harvest)
harvest$Year<-as.factor(harvest$Year)
harvest$TAG.NUMBER<-as.factor(harvest$TAG.NUMBER)
harvest$Date<-as.Date(harvest$Date, "%m/%d/%Y")
summary(harvest$Lake)
# can delete the single "five island" because it's an escaped fish caught in DSM
harvest<-droplevels(subset(harvest, Lake != "Five Island" &
                                    Lake != "Storm" &
                                    Lake != "S Twin" &
                                    Lake != "Blue"))
levels(harvest$Lake)
harvest$Lake<-as.character(harvest$Lake)
harvest$Lake[harvest$Lake == "N Twin "] <- "N Twin"
harvest$Lake[harvest$Lake == "5 Island"]<- "Five Island"
harvest$Lake<-as.factor(harvest$Lake)
levels(harvest$Notes)
harvest$detect<-1

levels(harvest$Gear)
harvest$Gear<-as.character(harvest$Gear)
harvest$Gear[harvest$Gear == "Seine"] <- "COMMERCIAL HARVEST"
harvest$Gear<-as.factor(harvest$Gear)


## CAN I SET SPECIES BY LOOKING UP PREV TAG NUMBER?
harvest$Species<-factor(harvest$Species, levels = c("BIB","COC"))
dframe.blend <- dframe[!duplicated(dframe$TAG.NUMBER),] # removing recaps
harvest.blend<-left_join(harvest,dframe.blend, by = "TAG.NUMBER")
sum(is.na(harvest.blend$Species.y)) # have about 21 tags recovered with no prior record at silver, Center, and N Twin
sum(is.na(harvest.blend$Species.y))/length(harvest.blend$TAG.NUMBER) # about 1.6% of all recovered tags do not have prior info
harvest$Species<-harvest.blend$Species.y
harvest<-droplevels(subset(harvest, !is.na(harvest$Species)))
#
# correct a couple things
names(dframe)
names(harvest)
#
# Try to merge the frames now!
df<-rbind(dframe,harvest)
str(df)
summary(df$detect)
levels(df$Species)
levels(dframe$Species)
levels(harvest$Species)
levels(harvest$Lake)
```

```{r}
ntwin<-droplevels(subset(df, Lake == "N Twin"))
str(ntwin)


unique(ntwin$Date[ntwin$Gear != "COMMERCIAL HARVEST"])
ntwin.tag<-droplevels(subset(ntwin, Gear != "COMMERCIAL HARVEST"))
ntwin.tag<-subset(ntwin.tag, Date != "2019-05-06") # fish not tagged May 6 2019, std runs for recaps only
ntwin.tag<-subset(ntwin.tag, Date != "2019-05-17") # fish not tagged May 17 2019, std runs for recaps only

unique(ntwin$Date[ntwin$Gear == "COMMERCIAL HARVEST"])
ntwin.har<-droplevels(subset(ntwin, Gear == "COMMERCIAL HARVEST"))

ntwin<-rbind(ntwin.tag,ntwin.har)
ntwin<-ntwin[order(as.Date(ntwin$Date, format = "%Y-%m-%d")),]


ntwin.c<-dcast(ntwin, Date+Gear ~ Species + detect)


# harvest date breaks at:
### 2019-04-04; 2019-04-05; 2019-04-07; 2019-04-08; 2019-04-09
### 2019-5-16; 2019-5-17
### 2019-11-23; 2019-11-30, 2019-11-06, 2019-11-25

# North Twin Cast frame (columns 3 and 4)
# targeting only the harvest in April and May 2019, and that exploitation rate

#  M_i is the number of fish marked at the start of the period (all tags - harvest)

M_1<-data.frame(t(colSums(ntwin.c[c(1:31),c(3,4)]))) # marked in period 1 = pre 2019 harvest
exp_2<- t(data.frame(t(colSums(ntwin.c[c(32:38),c(3,4)]))) / M_1)*100 # tags recovered april and may 2019






# Five Island Lake
ntwin<-droplevels(subset(df, Lake == "Five Island"))
str(ntwin)

unique(ntwin$Date[ntwin$Gear != "COMMERCIAL HARVEST"])
ntwin.tag<-droplevels(subset(ntwin, Gear != "COMMERCIAL HARVEST"))

unique(ntwin$Date[ntwin$Gear == "COMMERCIAL HARVEST"])
ntwin.har<-droplevels(subset(ntwin, Gear == "COMMERCIAL HARVEST"))

ntwin<-rbind(ntwin.tag,ntwin.har)
ntwin<-ntwin[order(as.Date(ntwin$Date, format = "%Y-%m-%d")),]

ntwin.c<-dcast(ntwin, Date+Gear ~ Species + detect)

M_1<-data.frame(t(colSums(ntwin.c[c(1:17),c(3,4)]))) # marked in period 1 = pre 2019 harvest
exp_3<- t(data.frame(t(colSums(ntwin.c[c(18:21),c(3,4)]))) / M_1)*100 # tags recovered april and may 2019



# Silver Lake
ntwin<-droplevels(subset(df, Lake == "Silver"))
str(ntwin)

unique(ntwin$Date[ntwin$Gear != "COMMERCIAL HARVEST"])
ntwin.tag<-droplevels(subset(ntwin, Gear != "COMMERCIAL HARVEST"))

unique(ntwin$Date[ntwin$Gear == "COMMERCIAL HARVEST"])
ntwin.har<-droplevels(subset(ntwin, Gear == "COMMERCIAL HARVEST" & 
                                    Date == "2018-10-24"))

ntwin<-rbind(ntwin.tag,ntwin.har)
ntwin<-ntwin[order(as.Date(ntwin$Date, format = "%Y-%m-%d")),]

ntwin.c<-dcast(ntwin, Date+Gear ~ Species + detect)

M_1<-data.frame(t(colSums(ntwin.c[c(1:20),c(3,4)]))) # marked in period 1 = pre 2019 harvest
exp_4<- t(data.frame(t(colSums(ntwin.c[21,c(3,4)]))) / M_1)*100 # tags recovered april and may 2019


# Center Lake
ntwin<-droplevels(subset(df, Lake == "Center"))
str(ntwin)

unique(ntwin$Date[ntwin$Gear != "COMMERCIAL HARVEST"])
ntwin.tag<-droplevels(subset(ntwin, Gear != "COMMERCIAL HARVEST"))

unique(ntwin$Date[ntwin$Gear == "COMMERCIAL HARVEST"])
ntwin.har<-droplevels(subset(ntwin, Gear == "COMMERCIAL HARVEST"))

ntwin<-rbind(ntwin.tag,ntwin.har)
ntwin<-ntwin[order(as.Date(ntwin$Date, format = "%Y-%m-%d")),]

ntwin.c<-dcast(ntwin, Date+Gear ~ Species + detect)

M_1<-data.frame(t(colSums(ntwin.c[c(1:32),c(3,4)]))) # marked in period 1 = pre 2019 harvest
exp_1<- t(data.frame(t(colSums(ntwin.c[33,c(3,4)]))) / M_1)*100 # tags recovered april and may 2019



Tag.exp<-c(0,0, # no harvest in blue
           NA,NA,exp_1[2,],exp_2[1,], # center
           exp_3[2,],exp_3[1,],
           0,0,exp_2[2,],exp_2[1,],
           0,0,0,0,
           exp_4[2,],exp_4[1,],
           0,0)


bmd.change$tag.exp<-Tag.exp
```

# get a plot in there:

```{r}
str(bmd.change)
max(bmd.change$delta.bmd)
bmd.change$tag.exp[bmd.change$Lake == "S Twin" & 
                   bmd.change$Species == "Bigmouth Buffalo" & 
                   bmd.change$Year.Range == "2017-2018"]<- NA
bmd.change$delta.bmd[bmd.change$Lake == "N Twin" & 
                     bmd.change$Species == "Common Carp" & 
                     bmd.change$Year.Range == "2017-2018"]<- ((34/21.2)-1)*100


fig5<-ggplot(bmd.change, aes(x=tag.exp, y = delta.bmd))+
          geom_point(aes(shape=Lake,color=Year.Range), size =5)+
          geom_abline(intercept=0,slope=-1,lty=2)+
          facet_grid(.~Species)+
          xlim(c(0,30))+
          ylim(c(-100,100))+
          scale_shape_manual(values = c(15,16,17,8,13,11,25)) +
                  theme(panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(),
                       panel.background = element_rect(fill="white",colour="gray50"), 
                       axis.line = element_line(colour = "black"),
                       legend.position= "bottom",
                       legend.direction = "horizontal",
                       legend.title = element_blank(),
                       legend.background = element_blank())+
                  guides(shape = guide_legend(nrow=1))+
                  labs( x = "Percent exploitation (tags recovered from harvest)",
                        y = "Percent change in population",
                        title = "Percent Change in Population vs. Percent tags recovered")
fig5
```

