---
title: "pre2017 Lake Monitoring Summary"
author: "Martin A. Simonson"
date: "December 4, 2017"
output: pdf_document
---
```{r,include=FALSE}
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

packages(vegan)
packages(ggplot2)
packages(reshape2)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
```

# Lake Monitoring data summary

This is an attempt to sort and analyze lake monitoring data from Iowa DNR to summarize lake quality measurements at the six(seven) lakes of my study:

- North Twin Lake (Calhoun)                                 [090]
- South Twin Lake (Calhoun) ** Not monitored **
- Center Lake (Dickinson)                                   [019]
- Silver Lake (Dickinson)                                   [105]
- Storm Lake (Buena Vista)                                  [113]
- Five Island Lake (Palo Alto)                              [036]
- Blue Lake (Monona)                                        [012]

# Variables of interest!

There's a lot going on here. ~70 variables in total. I need to pare this down a bit.



After looking at the summary, let's focus on some common things here:

-Zooplankton total biomass in micrograms/L                  [25]
-Phytoplankton Total biomass in milligrams/L                [24]
-Average Epilimnetic Turbidity in NTU                       [23]
-Average Epilimnetic Specific Conductivity in uS/cm         [21]
-Average Epilimnetic Dissolved Oxygen in mg/L               [20]
-Average Epilimnetic pH                                     [18]
-Average Epilimnetic Temp in Celsius                        [17]
-Average Secchi Depth in meters                             [16]
-Total Nitrogen in mg/L                                     [11]
-Total Phosphorus in micrograms/L                           [12]
-Soluble Reactive Phosphorus in micrograms/L                [13]
-Total Suspended Solids in mg/L                             [08]



```{r,include=F}
setwd("C:/Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/")
df<-read.csv("130Lakes_IowaLakesDatabase.csv",header=T) # load data 

head(df)

# Trim variables and set numeric/factors
df<-df[,c(2,4,8,11,12,13,16,17,18,20,21,23,24,25)]
lapply(df,class)
df$Lake.ID<-as.factor(df$Lake.ID)
df$Epilimnetic.Average.Turbidity..NTU.<-as.numeric(as.character(df$Epilimnetic.Average.Turbidity..NTU.))
head(df)

# Trim Lake Names
df$Lake.ID<-as.character(df$Lake.ID)
df$Lake.ID[df$Lake.ID == 12]<-"Blue"
df$Lake.ID[df$Lake.ID == 19]<-"Center"
df$Lake.ID[df$Lake.ID == 36]<-"5 Island"
df$Lake.ID[df$Lake.ID == 90]<-"N. Twin"
df$Lake.ID[df$Lake.ID == 105]<-"Silver"
df$Lake.ID[df$Lake.ID == 113]<-"Storm"
df$Lake.ID<-as.factor(df$Lake.ID)

# Sort by lake
Center<-df[df$Lake.ID == "Center",]
FiveIsland<-df[df$Lake.ID == "5 Island",]
Silver<-df[df$Lake.ID == "Silver",]
Blue<-df[df$Lake.ID == "Blue",]
Storm<-df[df$Lake.ID == "Storm",]
NorthTwin<-df[df$Lake.ID == "N. Twin",]

# colmeans
CenterMns<-colMeans(Center[,3:14],na.rm=T)
FiveIslandMns<-colMeans(FiveIsland[,3:14],na.rm=T)
SilverMns<-colMeans(Silver[,3:14],na.rm=T)
BlueMns<-colMeans(Blue[,3:14],na.rm=T)
StormMns<-colMeans(Storm[,3:14],na.rm=T)
TwinMns<-colMeans(NorthTwin[,3:14],na.rm=T)

# Standard Deviation
CenterSD<-sapply(Center[,3:14],sd,na.rm=T)
FiveIslandSD<-sapply(FiveIsland[,3:14],sd,na.rm=T)
SilverSD<-sapply(Silver[,3:14],sd,na.rm=T)
BlueSD<-sapply(Blue[,3:14],sd,na.rm=T)
StormSD<-sapply(Storm[,3:14],sd,na.rm=TRUE)
TwinSD<-sapply(NorthTwin[,3:14],sd,na.rm=T)

# Counts
CenterN<-45-colSums(sapply(Center[,3:14],is.na))
FiveIslandN<-45-colSums(sapply(FiveIsland[,3:14],is.na))
SilverN<-45-colSums(sapply(Silver[,3:14],is.na))
BlueN<-45-colSums(sapply(Blue[,3:14],is.na))
StormN<-45-colSums(sapply(Storm[,3:14],is.na))
TwinN<-45-colSums(sapply(NorthTwin[,3:14],is.na))

# Error (95%CI)
CenterE<-qnorm(0.975)*CenterSD/sqrt(CenterN)
FiveIslandE<-qnorm(0.975)*FiveIslandSD/sqrt(FiveIslandN)
SilverE<-qnorm(0.975)*SilverSD/sqrt(SilverN)
BlueE<-qnorm(0.975)*BlueSD/sqrt(BlueN)
StormE<-qnorm(0.975)*StormSD/sqrt(StormN)
TwinE<-qnorm(0.975)*TwinSD/sqrt(TwinN)
  
  
# Create new df with lakemean/sd
df2<-cbind(BlueMns,BlueE,
           CenterMns,CenterE,
           FiveIslandMns,FiveIslandE,
           TwinMns,TwinE,
           SilverMns,SilverE,
           StormMns,StormE)

print(df2,cutoff=3)


```

So that basically sums up the table I want to put in the proposal.

Maybe some plots would be better?

```{r,}
# Box and Whisker Plots for each variable?
df3<-rbind(Blue,Center,FiveIsland,NorthTwin,Silver,Storm) #df of just my lakes
colnames(df3)<-c("Lake.ID","Lake.Name","TSS (mg/L)","TN (mg/L)","TP (µg/L)","SRP (µg/L)",
                 "Secchi (m)","Temp (C)","pH","DO (mg/L)","Sp.Cond (uS/cm)",
                 "Turbidity (NTU)","Phytoplankton (mg/L)", "Zooplankton (µg/L)")




par(mfrow=c(3,4),mar=c(3,5,1,1))
mycol<-c(wes_palette("Zissou"),"darkgreen")
for(i in 3:14) boxplot(df3[,i]~factor(df3$Lake.ID),ylab=colnames(df3)[i],col=mycol,outline=F)




```


