---
title: "Biomass & Density"
author: "Marty Simonson"
date: "December 3, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=FALSE}
rm(list=ls())
# Front Matter

packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
```

This document is to complete the _[hopefully]_ more accurate method of estimating total biomass by length-specific average weights multiplied by the fraction of the population in that weight category times the abundance estimate, including 95% lower and upper bounds of the abundance estimate

Need to create a data frame to match the 2018 method with 14 columns:
- Lake (alphabetical)
- C.L.BM (carp lower biomass in pounds not as it is currently in grams)
- C. BM (carp biomass estimate pounds)
- C.U.BM (carp upper biomass estimate pounds)
- C.L.BMDensity (guess what? carp lower biomass density in pounds per acre)
- C.BMDensity (carp biomass density estimate lbs./acre)
- C.U.BMDensity (carp biomass density lbs./acre)
- B.L.BM (buffalo biomass density lower bound pounds)
- B.BM (buffalo biomass estimate pounds)
- B.U.BM (buffalo biomass estimate upper bound pounds)
- B.L.BMDensity (buffalo biomass density lower bound in pounds per acre)
- B.BMDensity (buffalo biomass density estimate lbs./acre)
- B.U.BMDensity (buffalo biomass density estimate upper bound pounds per acre)
- Year (factor 2017)


# length-specific biomass estimates
subset lake first. 


```{r,}
#Lakes
data<-read.csv("2018 and 2017 Compiled Data.csv", header = T)
levels(data$Lake)
Blue <- subset(data, Lake == "Blue")
Center<-subset(data, Lake == "Center")
FiveIsland<-subset(data, Lake == "5 Island")
NTwin<-subset(data, Lake == "N. Twin")
Silver<-subset(data, Lake == "Silver")
STwin<-subset(data, Lake == "S. Twin")
Storm<-subset(data,Lake == "Storm")


```

then subset with just COC and BIB, drop NAs, create length bins, and bin-specific averange weights.

# Blue
```{r,}
###########################
#
# Blue Lake 2017 only
#
# Carp:
Blue.coc<-subset(Blue, Species =="COC" &
                       Year == "2017")
Blue.coc<-subset(Blue.coc, !is.na(Blue.coc$Length..mm.))

#Buffalo not needed for Blue

#25mm bins
Blue.COC.25 <- Blue.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
Blue.COC.50 <- Blue.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
# so now we have percent frequency in 25/50mm (1-2") bins. Need to calculate average weight for those bins.
# I forsee my method taking too long but it's all I can think of right now:
#       to reduce repitition, let's go with the 50mm (2") bins first.
#       need 1 line of code to subset a new frame for each length bin, 
#       then take mean weight.

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 200 & Length..mm. < 250 & !is.na(Weight..g.))
w1<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 250 & Length..mm. < 300 & !is.na(Weight..g.))
w2<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 300 & Length..mm. < 350 & !is.na(Weight..g.))
w3<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w4<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w5<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w6<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w7<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w8<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w9<-mean(Blue.coc.temp$Weight..g.)

Blue.coc.temp<-subset(Blue.coc, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w10<-mean(Blue.coc.temp$Weight..g.)



Blue.COC.50$mweight<-c(w1,w2,w3,w4,w5,w6,w7,w8,w9,w10) 

#######
#
# Calculate the total weight through a sum of fractional weights
# 
# Blue Population Estimates: 31993- 60835 - 113049
# Blue Lake area: 269 acres
#
# Convert to lbs
# convert to density
#
# Should end up with a list of 12 blue lake values 2x(L.BM,BM,U.BM,L.BMDensity,BMDensity,U.BMDensity)

Blue.COC.50$percent<-Blue.COC.50$percent/100 # percent to decimal

Blue.COC.50$LowerBM<-Blue.COC.50$percent * Blue.COC.50$mweight * 31993 # multiply fraction, mean weight, lower pop.
C.L.BM<-sum(Blue.COC.50$LowerBM/453.592)


Blue.COC.50$BMest<-Blue.COC.50$percent * Blue.COC.50$mweight * 60835 # multiply fraction, mean weight, pop. est
C.BM<-sum(Blue.COC.50$BMest/453.592)

Blue.COC.50$UpperBM<-Blue.COC.50$percent * Blue.COC.50$mweight * 113049 # multiply fraction, mean weight, upper pop.
C.U.BM<-sum(Blue.COC.50$UpperBM/453.592)

C.L.BMDensity<-C.L.BM/269
C.BMDensity<-C.BM/269
C.U.BMDensity<-C.U.BM/269

B.L.BM<-NA
B.BM<-NA
B.U.BM<-NA
B.L.BMDensity<-NA
B.BMDensity<-NA
B.U.BMDensity<-NA


Blue.BM.Density<-c(C.L.BM,C.BM,C.U.BM,C.L.BMDensity,C.BMDensity,C.U.BMDensity,
                   B.L.BM,B.BM,B.U.BM,B.L.BMDensity,B.BMDensity,B.U.BMDensity)


```

#Center
```{r,}

###########################
#
# Center Lake 2017 only
#
# Carp:
Center.coc<-subset(Center, Species =="COC" &
                           Year == "2017") #                      check for mis-matches here!
Center.coc<-subset(Center.coc, !is.na(Center.coc$Length..mm.)) 
#Buffalo
Center.bib<-subset(Center, Species =="BIB" &
                           Year == "2017")
Center.bib<-subset(Center.bib, !is.na(Center.bib$Length..mm.))
#25mm bins
Center.COC.25 <- Center.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
Center.COC.50 <- Center.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
# so now we have percent frequency in 25/50mm (1-2") bins. Need to calculate average weight for those bins.
# I forsee my method taking too long but it's all I can think of right now:
#       to reduce repitition, let's go with the 50mm (2") bins first.
#       need 1 line of code to subset a new frame for each length bin, 
#       then take mean weight.

#Center.coc.temp<-subset(Center.coc, Length..mm. >= 50 & Length..mm. < 100 & !is.na(Weight..g.))
#w1<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 100 & Length..mm. < 150 & !is.na(Weight..g.))
w1<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 150 & Length..mm. < 200 & !is.na(Weight..g.))
w2<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 200 & Length..mm. < 250 & !is.na(Weight..g.))
w3<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w4<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w5<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w6<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w7<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w8<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w9<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w10<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w11<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 750 & Length..mm. < 800 & !is.na(Weight..g.))
w12<-mean(Center.coc.temp$Weight..g.)

Center.coc.temp<-subset(Center.coc, Length..mm. >= 800 & Length..mm. < 850 & !is.na(Weight..g.))
w13<-mean(Center.coc.temp$Weight..g.)


Center.COC.50$mweight<-c(w6,w6,w6,w6,w6,w6,w7,w8,w9,w10,w11,w12,w13) #changed last two to w11 and first to w3 

# Now Buffalo

#25mm bins
Center.bib.25 <- Center.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
Center.bib.50 <- Center.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()

Center.bib.temp<-subset(Center.bib, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w1<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w2<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w3<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w4<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w5<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w6<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w7<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w8<-mean(Center.bib.temp$Weight..g.)

Center.bib.temp<-subset(Center.bib, Length..mm. >= 750 & Length..mm. < 800 & !is.na(Weight..g.))
w9<-mean(Center.bib.temp$Weight..g.)

Center.bib.50$mweight<-c(w1,w2,w3,w4,w5,w6,w7,w8,w9)

#######
#
# Calculate the total weight through a sum of fractional weights
# 
# Center Carp Population Estimates: 6769 - 8675 - 11578
# Center Buff Population Estimates: 11243 - 12999 - 15463
# Center Lake area: 220 acres
#
# Convert to lbs
# convert to density
#
# Should end up with a list of 6 (six) blue lake values (L.BM,BM,U.BM,L.BMDensity,BMDensity,U.BMDensity)

Center.COC.50$percent<-Center.COC.50$percent/100 # percent to decimal

Center.COC.50$LowerBM<-Center.COC.50$percent * Center.COC.50$mweight * 6769 # multiply fraction, mean weight, lower pop.
C.L.BM<-sum(Center.COC.50$LowerBM/453.592)

Center.COC.50$BMest<-Center.COC.50$percent * Center.COC.50$mweight * 8675 # multiply fraction, mean weight, pop. est
C.BM<-sum(Center.COC.50$BMest/453.592)

Center.COC.50$UpperBM<-Center.COC.50$percent * Center.COC.50$mweight * 11578# multiply fraction, mean weight, upper pop.
C.U.BM<-sum(Center.COC.50$UpperBM/453.592)

C.L.BMDensity<-C.L.BM/220
C.BMDensity<-C.BM/220
C.U.BMDensity<-C.U.BM/220

# now buffalo
Center.bib.50$percent<-Center.bib.50$percent/100
Center.bib.50$LowerBM<-Center.bib.50$percent * Center.bib.50$mweight * 11243 # multiply fraction, mean weight, lower pop.
B.L.BM<-sum(Center.bib.50$LowerBM/453.592)

Center.bib.50$BMest<-Center.bib.50$percent * Center.bib.50$mweight * 12999 # multiply fraction, mean weight, pop. est
B.BM<-sum(Center.bib.50$BMest/453.592)

Center.bib.50$UpperBM<-Center.bib.50$percent * Center.bib.50$mweight * 15463 # multiply fraction, mean weight, upper pop.
B.U.BM<-sum(Center.bib.50$UpperBM/453.592)

B.L.BMDensity<-B.L.BM/220
B.BMDensity<-B.BM/220
B.U.BMDensity<-B.U.BM/220



Center.BM.Density<-c(C.L.BM,C.BM,C.U.BM,C.L.BMDensity,C.BMDensity,C.U.BMDensity,
                   B.L.BM,B.BM,B.U.BM,B.L.BMDensity,B.BMDensity,B.U.BMDensity)



```

# North Twin
```{r,}

###########################
#
# NTwin Lake
#
# Carp:
NTwin.coc<-subset(NTwin, Species =="COC" &
                         Year == "2017")
NTwin.coc<-subset(NTwin.coc, !is.na(NTwin.coc$Length..mm.))
#Buffalo
NTwin.bib<-subset(NTwin, Species =="BIB" &
                         Year == "2017")
NTwin.bib<-subset(NTwin.bib, !is.na(NTwin.bib$Length..mm.))
#25mm bins
NTwin.COC.25 <- NTwin.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
NTwin.COC.50 <- NTwin.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
# so now we have percent frequency in 25/50mm (1-2") bins. Need to calculate average weight for those bins.
# I forsee my method taking too long but it's all I can think of right now:
#       to reduce repitition, let's go with the 50mm (2") bins first.
#       need 1 line of code to subset a new frame for each length bin, 
#       then take mean weight.

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 200 & Length..mm. < 250 & !is.na(Weight..g.))
w1<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 250 & Length..mm. < 300 & !is.na(Weight..g.))
w2<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 300 & Length..mm. < 350 & !is.na(Weight..g.))
w3<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w4<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w6<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w7<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w8<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w9<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w10<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w11<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w12<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 750 & Length..mm. < 800 & !is.na(Weight..g.))
w13<-mean(NTwin.coc.temp$Weight..g.)

NTwin.coc.temp<-subset(NTwin.coc, Length..mm. >= 800 & Length..mm. < 850 & !is.na(Weight..g.))
w14<-mean(NTwin.coc.temp$Weight..g.)


NTwin.COC.50$mweight<-c(w1,w2,w3,w4,w6,w7,w8,w9,w10,w11,w12,w13,w14)

# Now Buffalo

#25mm bins
NTwin.bib.25 <- NTwin.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
NTwin.bib.50 <- NTwin.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()

#NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 50 & Length..mm. < 100 & !is.na(Weight..g.))
#w1<-mean(NTwin.bib.temp$Weight..g.)



NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 300 & Length..mm. < 350 & !is.na(Weight..g.))
w1<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w2<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w3<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w4<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w5<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w6<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w7<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w8<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w9<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 750 & Length..mm. < 800 & !is.na(Weight..g.))
w10<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 800 & Length..mm. < 850 & !is.na(Weight..g.))
w11<-mean(NTwin.bib.temp$Weight..g.)

NTwin.bib.temp<-subset(NTwin.bib, Length..mm. >= 850 & Length..mm. < 900 & !is.na(Weight..g.))
w12<-mean(NTwin.bib.temp$Weight..g.)


NTwin.bib.50$mweight<-c(w2,w2,w3,w4,w5,w6,w7,w8,w9,w10,w11,w11) #changed last two to w11 and first to w3 

#######
#
# Calculate the total weight through a sum of fractional weights
# 
# NTwin Carp Population Estimates: 1231 - 2057 - 3898
# NTwin Buff Population Estimates: 7276- 12963 - 23038
# NTwin Lake area: 453 acres
#
# Convert to lbs
# convert to density
#
# Should end up with a list of 6 (six) blue lake values (L.BM,BM,U.BM,L.BMDensity,BMDensity,U.BMDensity)

NTwin.COC.50$percent<-NTwin.COC.50$percent/100 # percent to decimal

NTwin.COC.50$LowerBM<-NTwin.COC.50$percent * NTwin.COC.50$mweight * 1231# multiply fraction, mean weight, lower pop.
C.L.BM<-sum(NTwin.COC.50$LowerBM/453.592)


NTwin.COC.50$BMest<-NTwin.COC.50$percent * NTwin.COC.50$mweight * 2057 # multiply fraction, mean weight, pop. est
C.BM<-sum(NTwin.COC.50$BMest/453.592)

NTwin.COC.50$UpperBM<-NTwin.COC.50$percent * NTwin.COC.50$mweight * 3898 # multiply fraction, mean weight, upper pop.
C.U.BM<-sum(NTwin.COC.50$UpperBM/453.592)

C.L.BMDensity<-C.L.BM/453
C.BMDensity<-C.BM/453
C.U.BMDensity<-C.U.BM/453

# now buffalo
NTwin.bib.50$percent<-NTwin.bib.50$percent/100
NTwin.bib.50$LowerBM<-NTwin.bib.50$percent * NTwin.bib.50$mweight * 7276 # multiply fraction, mean weight, lower pop.
B.L.BM<-sum(NTwin.bib.50$LowerBM/453.592)

NTwin.bib.50$BMest<-NTwin.bib.50$percent * NTwin.bib.50$mweight * 12963 # multiply fraction, mean weight, pop. est
B.BM<-sum(NTwin.bib.50$BMest/453.592)

NTwin.bib.50$UpperBM<-NTwin.bib.50$percent * NTwin.bib.50$mweight * 23038 # multiply fraction, mean weight, upper pop.
B.U.BM<-sum(NTwin.bib.50$UpperBM/453.592)

B.L.BMDensity<-B.L.BM/453
B.BMDensity<-B.BM/453
B.U.BMDensity<-B.U.BM/453



NTwin.BM.Density<-data.frame(cbind(C.L.BM,C.BM,C.U.BM,C.L.BMDensity,C.BMDensity,C.U.BMDensity,
                   B.L.BM,B.BM,B.U.BM,B.L.BMDensity,B.BMDensity,B.U.BMDensity))

```

#South Twin
```{r,}

###########################
#
# STwin Lake
#
# Carp:
STwin.coc<-subset(STwin, Species =="COC" &
                         Year == "2017" )
STwin.coc<-subset(STwin.coc, !is.na(STwin.coc$Length..mm.))
#Buffalo
STwin.bib<-subset(STwin, Species =="BIB" &
                         Year == "2017")
STwin.bib<-subset(STwin.bib, !is.na(STwin.bib$Length..mm.))
#25mm bins
STwin.COC.25 <- STwin.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
STwin.COC.50 <- STwin.coc %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
# so now we have percent frequency in 25/50mm (1-2") bins. Need to calculate average weight for those bins.
# I forsee my method taking too long but it's all I can think of right now:
#       to reduce repitition, let's go with the 50mm (2") bins first.
#       need 1 line of code to subset a new frame for each length bin, 
#       then take mean weight.

#STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 50 & Length..mm. < 100 & !is.na(Weight..g.))
#w1<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 250 & Length..mm. < 300 & !is.na(Weight..g.))
w1<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 300 & Length..mm. < 350 & !is.na(Weight..g.))
w2<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w3<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w4<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w5<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w6<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w7<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w8<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w9<-mean(STwin.coc.temp$Weight..g.)

STwin.coc.temp<-subset(STwin.coc, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w10<-mean(STwin.coc.temp$Weight..g.)


STwin.COC.50$mweight<-c(w2,w2,w3,w4,w5,w6,w7,w8,w9,w9)

# Now Buffalo

#25mm bins
STwin.bib.25 <- STwin.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=25))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()
#50mm bins
STwin.bib.50 <- STwin.bib %>%
  add_count(Species) %>%
  rename(Species_count = n) %>%
  group_by(Species, Species_count, gr=cut(Length..mm., breaks = seq(0,1000,by=50))) %>%
  summarise(n = n()) %>%
  mutate(freq=n, percent = ((n/Species_count)*100),
         cumfreq = cumsum(freq), cumpercent = cumsum(percent)) %>%
  ungroup()

#STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 250 & Length..mm. < 300 & !is.na(Weight..g.))
#w1<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 300 & Length..mm. < 350 & !is.na(Weight..g.))
w1<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 350 & Length..mm. < 400 & !is.na(Weight..g.))
w2<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 400 & Length..mm. < 450 & !is.na(Weight..g.))
w3<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 450 & Length..mm. < 500 & !is.na(Weight..g.))
w4<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 500 & Length..mm. < 550 & !is.na(Weight..g.))
w5<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 550 & Length..mm. < 600 & !is.na(Weight..g.))
w6<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 600 & Length..mm. < 650 & !is.na(Weight..g.))
w7<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 650 & Length..mm. < 700 & !is.na(Weight..g.))
w8<-mean(STwin.bib.temp$Weight..g.)

STwin.bib.temp<-subset(STwin.bib, Length..mm. >= 700 & Length..mm. < 750 & !is.na(Weight..g.))
w9<-mean(STwin.bib.temp$Weight..g.)


STwin.bib.50$mweight<-c(w1,w2,w3,w4,w5,w6,w7,w9,w9) 

#######
#
# Calculate the total weight through a sum of fractional weights
# 
# STwin Carp Population Estimates: 12227 - 23204 - 41883
# STwin Buff Population Estimates: 5535 - 6895 - 8748
# STwin Lake area: 600 acres
#
# Convert to lbs
# convert to density
#
# Should end up with a list of 6 (six) blue lake values (L.BM,BM,U.BM,L.BMDensity,BMDensity,U.BMDensity)

STwin.COC.50$percent<-STwin.COC.50$percent/100 # percent to decimal

STwin.COC.50$LowerBM<-STwin.COC.50$percent * STwin.COC.50$mweight * 12227 # multiply fraction, mean weight, lower pop.
C.L.BM<-sum(STwin.COC.50$LowerBM/453.592)


STwin.COC.50$BMest<-STwin.COC.50$percent * STwin.COC.50$mweight * 23204 # multiply fraction, mean weight, pop. est
C.BM<-sum(STwin.COC.50$BMest/453.592)

STwin.COC.50$UpperBM<-STwin.COC.50$percent * STwin.COC.50$mweight * 41883 # multiply fraction, mean weight, upper pop.
C.U.BM<-sum(STwin.COC.50$UpperBM/453.592)

C.L.BMDensity<-C.L.BM/600
C.BMDensity<-C.BM/600
C.U.BMDensity<-C.U.BM/600

# now buffalo
STwin.bib.50$percent<-STwin.bib.50$percent/100
STwin.bib.50$LowerBM<-STwin.bib.50$percent * STwin.bib.50$mweight * 5535 # multiply fraction, mean weight, lower pop.
B.L.BM<-sum(STwin.bib.50$LowerBM/453.592)

STwin.bib.50$BMest<-STwin.bib.50$percent * STwin.bib.50$mweight * 6895 # multiply fraction, mean weight, pop. est
B.BM<-sum(STwin.bib.50$BMest/453.592)

STwin.bib.50$UpperBM<-STwin.bib.50$percent * STwin.bib.50$mweight * 8748 # multiply fraction, mean weight, upper pop.
B.U.BM<-sum(STwin.bib.50$UpperBM/453.592)

B.L.BMDensity<-B.L.BM/600
B.BMDensity<-B.BM/600
B.U.BMDensity<-B.U.BM/600



STwin.BM.Density<-data.frame(cbind(C.L.BM,C.BM,C.U.BM,C.L.BMDensity,C.BMDensity,C.U.BMDensity,
                   B.L.BM,B.BM,B.U.BM,B.L.BMDensity,B.BMDensity,B.U.BMDensity))
```

# Bring it all together for total biomass and biomass density

```{r,}
# Rbind each of the 7 lists of twelve (data frames)

All.biomass<-rbind(Blue.BM.Density,
                   Center.BM.Density,
                   NTwin.BM.Density,
                   STwin.BM.Density)
Lake<-c("Blue","Center","N Twin","S Twin")
Year<-c("2017","2017","2017","2017")
All.biomass<-cbind(Lake,Year,All.biomass)
write.csv(All.biomass,"2017_BiomassDensityEstimates.csv")

dat<-read.csv("2017_BiomassDensityEstimates.csv")
dat
```