---
title: "Marcus_Project"
author: "Martin Simonson"
date: "March 1, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---



```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
#packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
#packages(lattice)
#packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
#packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
#packages(GISTools)
#packages(maps)
#packages(mapproj)
packages(splitstackshape)
```


# splitting and subsetting buffalo data for Marcus' analysis.

Here's the deal. Marcus can't handle all the buffalo structures we have, therefore, I need to figure out a way to subset our age samples that still provides an even distribution of the data.

- We can only use Center, Silver, and Five Island.
          Storm didn't have enough buffalo sampled and dissected. (I think, who knows anymore)
          We don't have the two otolith methods from 2017, but could go back and look for double otoliths in envelopes?
          
Start with 2018 man. Don't kill yourself on a friday. There's hockey tonight.

```{r}
df<-read.csv("2018 and 2017 Compiled Data.csv", header = T)
head(df)
# remove extraneous data
df<-df[,-c(1,8:10,12:21)]
df<-droplevels(subset(df, Species == "BIB"))
df.2018<-droplevels(subset(df, Year == "2018"))
df.2018<-droplevels(subset(df.2018, Lake == "Center"|
                                    Lake == "5 ISland"|
                                    Lake == "Silver"))

summary(df.2018$Notes)

# only one note is "aging fish" 
df.2018.ages<-droplevels(subset(df.2018, Notes == "Aging Fish"))
df.2018.alive<-droplevels(subset(df.2018, Notes != "Aging Fish"))

# I hate life today.
summary(df.2018.ages$Length..mm.)


df.c<-droplevels(subset(df.2018.ages, Lake == "Center"))

df.c.sub<-stratified(df.c, "Length..mm.", size = 3, seed = 1)

```






