---
title: "Harvest Data Summary"
author: "Martin A. Simonson"
date: "December 4, 2017"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r,include=FALSE}
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
```

# Harvest data summary

This is an attempt to collect harvest data from different Iowa DNR district offices to summarize harvest at the six(seven) lakes of my study:

- North Twin Lake (Calhoun) (090)
- South Twin Lake (Calhoun) (NA)
- Center Lake (Dickinson)   (019)
- Silver Lake (Dickinson)   (105)
- Storm Lake (Buena Vista)  (113)
- Five Island Lake (Palo Alto)  (036)
- Blue Lake (Monona)        (012)



As of 12/4/2017 I have 6 lakes of data (nothing for s. twin lake)

```{r,include=F}
df<-read.csv("SpiritHarvest.csv",header=T) # load data 


summary(df$Lake) 
summary(df$Carp)
df$Year<-as.factor(df$Year)
```

# Harvest per acre

To standardize these systems and make comparisons between lakes more meaningful, I need to scale the harvest (lbs.) by the size of each lake (acres). Then I can plot historical harvest in terms of lbs/acre and look at plots side by side.

```{r,}
df<-read.csv("SpiritHarvest.csv",header=T) # load data 
df$Year<-as.factor(df$Year)
head(df)
########################################################
df<-df[df$Lake == "Center Lake" |
       df$Lake == "Five Island Lake" |
       df$Lake == "Silver Lake (Dickinson)" |
       df$Lake == "Blue Lake" | 
       df$Lake == "Storm Lake" |
       df$Lake == "North Twin Lake",]
########################################################
summary(df)
df$Acres[df$Lake == "Center Lake"] <- 220
df$Acres[df$Lake == "Blue Lake"] <- 275
df$Acres[df$Lake == "Five Island Lake"] <- 973
df$Acres[df$Lake == "North Twin Lake"] <- 463
df$Acres[df$Lake == "Silver Lake (Dickinson)"] <- 1066
df$Acres[df$Lake == "Storm Lake"] <- 3140

df$Total.Harvest.byAcre<-(df$Carp+df$Buffalo+df$Drum)/df$Acres

df$Buffalo<-df$Buffalo/df$Acres
df$Drum<-df$Drum/df$Acres
df$Carp<-df$Carp/df$Acres
df$Total.Rough.Fish<-df$Total.Rough.Fish/df$Acres

summary(df)
df<-df[,c(1,4,5,7)]
```


```{r,}
# create individual data frames for each lake
Center<-df[df$Lake == "Center Lake",]
FiveIsland<-df[df$Lake == "Five Island Lake",]
Silver<-df[df$Lake == "Silver Lake (Dickinson)",]
Blue<-df[df$Lake == "Blue Lake",]
Storm<-df[df$Lake == "Storm Lake",]
NorthTwin<-df[df$Lake == "North Twin Lake",]


```


# Plotting

I think the best option is to go lake by lake, and try to have the two species' harvest summed for each Year
  
Then perhaps a final plot with all three lakes on them together, with all 2 variables (two spp. and total).

```{r,}

#####################
# 
#
#    Center Lake
#
#
#####################
# Melt!
Center.m<-melt(Center)
# Harvest plot by year
CenterPlot<-ggplot(Center.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Center Lake Commercial Harvest (Different Y-axis scale)",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,300)+
                theme(axis.text.x=element_text(angle=-90)) +
                theme(axis.text.y=element_text(colour="red")) +
                theme(axis.title.y=element_text(colour="red"))

#####################
# 
#
#    Silver Lake
#
#
#####################
# Melt!
Silver.m<-melt(Silver)
# Harvest plot by year
SilverPlot<-ggplot(Silver.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Silver Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    5 Island Lake
#
#
#####################
# Melt!
FiveIsland.m<-melt(FiveIsland)
# Harvest plot by year
FivePlot<-ggplot(FiveIsland.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Five Island Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))


#####################
# 
#
#    Storm Lake
#
#
#####################
# Melt!
Storm.m<-melt(Storm)
# Harvest plot by year
StormPlot<-ggplot(Storm.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Storm Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    Blue Lake
#
#
#####################
# Melt!
Blue.m<-melt(Blue)
# Harvest plot by year
BluePlot<-ggplot(Blue.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Blue Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)") +
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    North Twin Lake
#
#
#####################
# Melt!
NorthTwin.m<-melt(NorthTwin)
# Harvest plot by year
TwinPlot<-ggplot(NorthTwin.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "North Twin Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

grid.arrange(BluePlot,CenterPlot,FivePlot,SilverPlot,StormPlot,TwinPlot,
             ncol=2,nrow=3)

```






```{r,}

#####################
# 
#
#    All 6 Lakes
#
#
#####################
#All<-rbind(Silver,FiveIsland,Center,Storm,Blue,NorthTwin)
# Melt!
All.m<-melt(df)
# Harvest plot by year
ggplot(All.m, aes(x=Lake, y=value, fill=variable), ylab = "Harvest (lbs. per acre)", main = "All Lake Harvest") + 
                geom_bar(stat='identity', position = "dodge", colour = "black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Cumulative Commercial Harvest by Lake",
                     y = "Harvest (lbs/acre)")

```


# Overall Harvest regardless of lake size


```{r,}
df<-read.csv("SpiritHarvest.csv",header=T) # load data 
df$Year<-as.factor(df$Year)
head(df)
########################################################
df<-df[df$Lake == "Center Lake" |
       df$Lake == "Five Island Lake" |
       df$Lake == "Silver Lake (Dickinson)" |
       df$Lake == "Blue Lake" | 
       df$Lake == "Storm Lake" |
       df$Lake == "North Twin Lake",]
########################################################
head(df)

summary(df)
df<-df[,c(1,4,5,7)]
```


```{r,}
# create individual data frames for each lake
Center<-df[df$Lake == "Center Lake",]
FiveIsland<-df[df$Lake == "Five Island Lake",]
Silver<-df[df$Lake == "Silver Lake (Dickinson)",]
Blue<-df[df$Lake == "Blue Lake",]
Storm<-df[df$Lake == "Storm Lake",]
NorthTwin<-df[df$Lake == "North Twin Lake",]


```


# Plotting

I think the best option is to go lake by lake, and try to have the two speces for each Year

This will be total harvest, not by lake size. This is to see the catch in recent years in regards to 2019 harvest contracts.

```{r,}

#####################
# 
#
#    Center Lake
#
#
#####################
# Melt!
Center.m<-melt(Center)
# Harvest plot by year
CenterPlot2<-ggplot(Center.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Center Lake Commercial Harvest",
                     y = "Harvest (lbs.)")+
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    Silver Lake
#
#
#####################
# Melt!
Silver.m<-melt(Silver)
# Harvest plot by year
SilverPlot2<-ggplot(Silver.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Silver Lake Commercial Harvest",
                     y = "Harvest (lbs.)")+
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    5 Island Lake
#
#
#####################
# Melt!
FiveIsland.m<-melt(FiveIsland)
# Harvest plot by year
FivePlot2<-ggplot(FiveIsland.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Five Island Lake Commercial Harvest",
                     y = "Harvest (lbs.)")+
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))


#####################
# 
#
#    Storm Lake
#
#
#####################
# Melt!
Storm.m<-melt(Storm)
# Harvest plot by year
StormPlot2<-ggplot(Storm.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Storm Lake Commercial Harvest",
                     y = "Harvest (lbs.)")+
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    Blue Lake
#
#
#####################
# Melt!
Blue.m<-melt(Blue)
# Harvest plot by year
BluePlot2<-ggplot(Blue.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Blue Lake Commercial Harvest",
                     y = "Harvest (lbs.)") +
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    North Twin Lake
#
#
#####################
# Melt!
NorthTwin.m<-melt(NorthTwin)
# Harvest plot by year
TwinPlot2<-ggplot(NorthTwin.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Royal1")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "North Twin Lake Commercial Harvest",
                     y = "Harvest (lbs.)")+
                ylim(0,80000)+
                theme(axis.text.x=element_text(angle=-90))

grid.arrange(BluePlot2,CenterPlot2,FivePlot2,SilverPlot2,StormPlot2,TwinPlot2,
             ncol=2,nrow=3)

```


# Then put both together: harvest per acre and overall harvest.
```{r}
grid.arrange(SilverPlot,SilverPlot2,StormPlot,StormPlot2,TwinPlot,TwinPlot2,
             ncol=2,nrow=3)
```

