---
title: "Harvest Data Summary"
author: "Martin A. Simonson"
date: "December 4, 2017"
output: pdf_document
---
```{r,include=FALSE}
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
packages(lattice)
packages(tikzDevice)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
```

# Harvest data summary

This is an attempt to collect harvest data from different Iowa DNR district offices to summarize harvest at the six(seven) lakes of my study:

- North Twin Lake (Calhoun) (090)
- South Twin Lake (Calhoun) (NA)
- Center Lake (Dickinson)   (019)
- Silver Lake (Dickinson)   (105)
- Storm Lake (Buena Vista)  (113)
- Five Island Lake (Palo Alto)  (036)
- Blue Lake (Monona)        (012)



As of 12/4/2017 I have 6 lakes of data (nothing for s. twin lake)

```{r,include=F}
setwd("C:/Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/")
df<-read.csv("SpiritHarvest.csv",header=T) # load data 


summary(df$Lake) 
summary(df$Carp)
df$Year<-as.factor(df$Year)
```

# Harvest per acre

To standardize these systems and make comparisons between lakes more meaningful, I need to scale the harvest (lbs.) by the size of each lake (acres). Then I can plot historical harvest in terms of lbs/acre and look at plots side by side.

```{r,}
head(df)
########################################################
df<-df[df$Lake == "Center Lake" |
       df$Lake == "Five Island Lake" |
       df$Lake == "Silver Lake (Dickinson)" |
       df$Lake == "Blue Lake" | 
       df$Lake == "Storm Lake" |
       df$Lake == "North Twin Lake",]
########################################################
summary(df)
df$Acres[df$Lake == "Center Lake"] <- 220
df$Acres[df$Lake == "Blue Lake"] <- 275
df$Acres[df$Lake == "Five Island Lake"] <- 973
df$Acres[df$Lake == "North Twin Lake"] <- 463
df$Acres[df$Lake == "Silver Lake (Dickinson)"] <- 1066
df$Acres[df$Lake == "Storm Lake"] <- 3140

df$Total.Harvest<-(df$Carp+df$Buffalo+df$Drum)/df$Acres

df$Buffalo<-df$Buffalo/df$Acres
df$Drum<-df$Drum/df$Acres
df$Carp<-df$Carp/df$Acres
df$Total.Rough.Fish<-df$Total.Rough.Fish/df$Acres

summary(df)
df<-df[,c(1,4,5:8)]
```


```{r,}
# create individual data frames for each lake
Center<-df[df$Lake == "Center Lake",]
FiveIsland<-df[df$Lake == "Five Island Lake",]
Silver<-df[df$Lake == "Silver Lake (Dickinson)",]
Blue<-df[df$Lake == "Blue Lake",]
Storm<-df[df$Lake == "Storm Lake",]
NorthTwin<-df[df$Lake == "North Twin Lake",]


```


# Time to import some water quality data
```{r,}
WQ<-read.csv("C:/Users/M_Sim/OneDrive/IOWA/Iowa State University/CarpR/WaterQuality/130Lakes_IowaLakesDatabase.csv", header = TRUE)
str(WQ)
summary(WQ$Lake.ID)

BlueWQ<-WQ[WQ$Lake.ID == 12,]
CenterWQ<-WQ[WQ$Lake.ID == 19,]
FiveIslandWQ<-WQ[WQ$Lake.ID == 36,]
NTwinWQ<-WQ[WQ$Lake.ID == 90,]
SilverWQ<-WQ[WQ$Lake.ID == 105,]
StormWQ<-WQ[WQ$Lake.ID == 113,]

# Need to average by year maybe...
BlueWQ<-aggregate(BlueWQ, by=list(BlueWQ$Year), FUN=mean, na.rm = T)
CenterWQ<-aggregate(CenterWQ, by=list(CenterWQ$Year), FUN=mean, na.rm = T)
FiveIslandWQ<-aggregate(FiveIslandWQ, by=list(FiveIslandWQ$Year), FUN=mean, na.rm = T)
NTwinWQ<-aggregate(NTwinWQ, by=list(NTwinWQ$Year), FUN=mean, na.rm = T)
SilverWQ<-aggregate(SilverWQ, by=list(SilverWQ$Year), FUN=mean, na.rm = T)
StormWQ<-aggregate(StormWQ, by=list(StormWQ$Year), FUN=mean, na.rm = T)


```

#Merge harvest and water quality data frames ---- maybe
```{r,}
Center<-merge(Center,CenterWQ,by="Year",all=T)
# The problem with this is it drops values that aren't common between both datasets....







```


# Plotting

I think the best option is to go lake by lake, and try to have the two species (and then total) for each Year
  
Then perhaps a final plot with all three lakes on them together, with all 3 variables (two spp. and total).

```{r,}

#####################
# 
#
#    Center Lake
#
#
#####################
# Melt!
Center.m<-melt(Center)
# Harvest plot by year
CenterPlot<-ggplot(Center.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Center Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,300)+
                theme(axis.text.x=element_text(angle=-90)) +
                geom_line(data=CenterWQ, aes(x=Year, y=CenterWQ[,26]))

#####################
# 
#
#    Silver Lake
#
#
#####################
# Melt!
Silver.m<-melt(Silver)
# Harvest plot by year
SilverPlot<-ggplot(Silver.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Silver Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    5 Island Lake
#
#
#####################
# Melt!
FiveIsland.m<-melt(FiveIsland)
# Harvest plot by year
FivePlot<-ggplot(FiveIsland.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Five Island Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))


#####################
# 
#
#    Storm Lake
#
#
#####################
# Melt!
Storm.m<-melt(Storm)
# Harvest plot by year
StormPlot<-ggplot(Storm.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Storm Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    Blue Lake
#
#
#####################
# Melt!
Blue.m<-melt(Blue)
# Harvest plot by year
BluePlot<-ggplot(Blue.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Blue Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)") +
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

#####################
# 
#
#    North Twin Lake
#
#
#####################
# Melt!
NorthTwin.m<-melt(NorthTwin)
# Harvest plot by year
TwinPlot<-ggplot(NorthTwin.m, aes(x=Year, y=value, fill=variable)) + 
                geom_bar(stat='identity', position = "dodge", colour="black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "North Twin Lake Commercial Harvest",
                     y = "Harvest (lbs. per acre)")+
                ylim(0,100)+
                theme(axis.text.x=element_text(angle=-90))

grid.arrange(BluePlot,CenterPlot,FivePlot,SilverPlot,StormPlot,TwinPlot,
             ncol=2,nrow=3)

```






```{r,}

#####################
# 
#
#    All 6 Lakes
#
#
#####################
#All<-rbind(Silver,FiveIsland,Center,Storm,Blue,NorthTwin)
# Melt!
All.m<-melt(df)
# Harvest plot by year
ggplot(All.m, aes(x=Lake, y=value, fill=variable), ylab = "Harvest (lbs. per acre)", main = "All Lake Harvest") + 
                geom_bar(stat='identity', position = "dodge", colour = "black") +
                scale_fill_manual(values = wes_palette("Darjeeling2")) +
                theme(legend.position = "bottom",legend.title = element_blank()) +
                labs(title = "Cumulative Commercial Harvest by Lake",
                     y = "Harvest (lbs/acre)")

```



