---
title: "Ecostats Final Project JAGS and Graphics"
author: "Marty Simonson, Brandon Maahs, Bobby Cope"
date: "December 18, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r}
# Front Matter
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}
# Working Directory and Packages
setwd("~/BMB-COC/STAT534/Final Project/STAT534_FishCPUE_Part2/") # set to wherever you downloaded files
packages(rjags)
packages(mcmcplots)
packages(ggplot2)
packages(gridExtra)
```

# Intro/ReadMe

## Part 1:

This is the initial project proposal, using data from Blue Lake in 2018. However, the beta densities weren't very strong and we wanted to include water temp, so we changed the model in Part 2, circa line 217.

Last updated by MAS 12/13/2019

```{r,}
cpue <- read.csv('EcostatFinalCPUE.csv') # read in Blue Lake 2018 carp CPUE and env. covariates

# define model
cpue.m <- jags.model(
  'Marti.bug',
  data=list(logPop=log(72140), sdlogPop=0.079, # log pouplation size and 
                                               # log sd of population size from data (sdlogpop becomes Tau) 
    n=dim(cpue)[1], # samples
    logCPUE=log(cpue$CPUE), # log-transformed CPUE from data 
    AirTemp=cpue$AirTemp, # air temp in celcius (Subract 20 in .bug file)
    Wind=cpue$Wind), # wind speed in mph, (subtract 10 in .bug file)
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method
  
coef(cpue.m)

# sample the MCMC chains 1000 times as burn-in, ignore the samples. this gets thrown away
update(cpue.m, 1000)

# we want to plot:
## draw curve of estimated catchabaility as function of air temp and wind using betas
## uncertainty 
cpue.i <- coda.samples(cpue.m, # model
                       c('logcatch','b0','b1','b2','sd'), # parameters we want back
                       1000) # number of samples

par(mar=c(3,3,3,3)+0.2, mgp=c(2,0.8,0))
plot(cpue.i)

summary(cpue.i)

# test of package 'mcmcplots' for b1 trace, density
denplot(cpue.i, parms = c('b1', collapse = FALSE, auto.layout = TRUE))
traplot(cpue.i, parms = c('b1'))
```

# Part 1 plots

```{r,}
# extracting a parameter, plotting it's confidence region

# beta 0: intercept
varnames(cpue.i) # reference variables in model
str(cpue.i) # structure of lists
beta0.out <- cbind(cpue.i[[1]][,1], cpue.i[[2]][,1], cpue.i[[3]][,1]) # subsetting for beta1 mcmc output
attributes(beta0.out) <- NULL # dropping attributes
dens.beta0 <- density(beta0.out) # density
q25.beta0 <- quantile(beta0.out, .025) # lower quantile
q975.beta0 <- quantile(beta0.out, .975) # upper quantile
dd.beta0 <- with(dens.beta0, data.frame(x,y)) # density dataframe

beta0<-qplot(x, y, data = dd.beta0, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta0, x>q25.beta0 & x<q975.beta0),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 0 - Intercept")+
  ylab("Density of Beta 0")+
  ggtitle("Blue Lake Carp Model: Intercept")
beta0










# beta 1: air temp
varnames(cpue.i) # reference variables in model
str(cpue.i) # structure of lists
beta1.out <- cbind(cpue.i[[1]][,2], cpue.i[[2]][,2], cpue.i[[3]][,2]) # subsetting for beta1 mcmc output
attributes(beta1.out) <- NULL # dropping attributes
dens.beta1 <- density(beta1.out) # density
q25.beta1 <- quantile(beta1.out, .025) # lower quantile
q975.beta1 <- quantile(beta1.out, .975) # upper quantile
dd.beta1 <- with(dens.beta1, data.frame(x,y)) # density dataframe

beta1<-qplot(x, y, data = dd.beta1, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta1, x>q25.beta1 & x<q975.beta1),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 1: Slope of log(AirTemp)")+
  ylab("Density")+
  ggtitle("Density of Beta 1")
beta1

# beta 2, slope of wind speed
varnames(cpue.i)
str(cpue.i)
beta2.out <- cbind(cpue.i[[1]][,3], cpue.i[[2]][,3], cpue.i[[3]][,3]) # subsetting beta2 mcmc output
attributes(beta2.out) <- NULL # dropping attributes
dens.beta2 <- density(beta2.out)
q25.beta2 <- quantile(beta2.out, .025)
q975.beta2 <- quantile(beta2.out, .975)
dd.beta2 <- with(dens.beta2, data.frame(x,y))

beta2<-qplot(x, y, data = dd.beta2, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta2, x>q25.beta2 & x<q975.beta2),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 2: Slope of log(Wind Speed)")+
  ylab("Density")+
  ggtitle("Density of Beta 2")
beta2
# BETA 2 DOES NOT APPEAR TO HAVE STRONG EFFECT

## Combine two plots
grid.arrange(beta0,beta1,beta2,ncol=3)
```

# new model *without* wind, and subsequent regression plot and expected vs. observed cpue plot

- note the new .bug file "Marti2.bug"

```{r,}
## New model without Wind
cpue.b.2 <- jags.model(
  'Marti2.bug',
  data=list(logPop=log(72140), sdlogPop=0.079, # log pouplation size and log sd of population size from data (sdlogpop becomes Tau) 
            n=dim(cpue)[1], # samples
            logCPUE=log(cpue$CPUE), # log-transformed CPUE from data 
            AirTemp=cpue$AirTemp), # air temp in celcius (Subract 20 in .bug file)
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method

coef(cpue.b.2)

# sample the MCMC chains 1000 times as burn-in, ignore the samples. this gets thrown away
update(cpue.b.2, 1000)

cpue.b <- coda.samples(cpue.b.2, # model
                       c('logN','b0','b1','sd'), # parameters we want back
                       1000) # number of samples

cpue.b.out<-summary(cpue.b)
varnames(cpue.b.out)
str(cpue.b.out)
linear <- data.frame(cpue.b.out[[1]][c(1:3),], cpue.b.out[[2]][c(1:3),c(1,5)])
str(linear)

catch.int<-cpue.b.out[[1]][1,1]
catch.slope<-cpue.b.out[[1]][2,1]


m <- cpue.b.out$statistics[,"Mean"]
l <- cpue.b.out$quantiles[,"2.5%"]
u <- cpue.b.out$quantiles[,"97.5%"]

dim(cpue)
length(m)

#####################################################################
#
# Plot of linear regression, x = ln(air temp) and y = ln(CPUE)
#
######################################################################

ggplot(cpue, aes(x = log(AirTemp), y = log(CPUE/72140))) +
  geom_point(size = 3) +
  geom_abline(slope = catch.slope, intercept = catch.int, size = 1.5, lty = 2)+
  xlim(1,4)+
  ylim(-8,-4)+
  geom_ribbon(aes(ymin = catch.int-l[1], ymax = catch.int+u[1]), alpha = 0.2)+ # can't get this to work
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("ln (Air Temperature)")+
  ylab("ln (CPUE)")+
  ggtitle("Linear Model of CPUE as a function of log air temperature")

#####################################################################
#
# Plot of expected vs. observed cpue
#
######################################################################

# plot of distribution of ElogCPUE in boxplots, overlaid with observed CPUE 
cpue.k <- coda.samples(cpue.b.2, 
                       c('ElogCPUE'), # parameters we want back
                       1000) # number of samples
summary(cpue.k)
# storing model outputas object
cpue.k.out<-summary(cpue.k)

cpue.plot.data<-cbind(cpue,cpue.k.out[[1]][c(1:26),],cpue.k.out[[2]][c(1:26),c(1:5)])
str(cpue.plot.data)

ggplot(cpue.plot.data, aes(x = Date, y = log(CPUE)))+
         geom_boxplot(aes(ymin = cpue.plot.data$`2.5%`, lower = cpue.plot.data$`25%`, 
                           middle = cpue.plot.data$Mean, 
                           upper = cpue.plot.data$`75%`, ymax = cpue.plot.data$`97.5%`),
  stat = "identity", fill = "cornflowerblue")+
  geom_point(size=5)+
  theme_classic()+
  theme(axis.text=element_text(size=8),
        axis.text.x = element_text(angle=90),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Date")+
  ylab("Expected and observed ln (CPUE)")+
  ggtitle("Blue Lake Carp Comparison of Expected Observed ln (CPUE)")

####################################################################################################
```

## Part 2:

We've taken the provided rjags code (Mari.r and chunks above) and model fit in the .bug file (Marti.bug - full and Marti2.bug - reduced) and adapted it for a new scenario. New files are this R Markdown (Ecostats Final Project JAGS and Graphics.Rmd) and bugs model (NTWIN.bug)

We switched study system (lake) from Blue Lake to North Twin lake, for a few reasons:

- wind speed didn't have strong effect in linear model of catchability. We include it here but may reduce our model if needed later

- air temp was OK in the blue lake model, but we are fairly confident that water temp would be a better covariate than air temp. Still, air temp is kept in this model and it may be dropped from the reduced model later.

- We now have water temperatures for all but 2 of the 17 sample days (will be removed)

- We now have two data files (NTwin18CPUEbuff.csv and NTwin18CPUEcarp.csv), one for common carp and one for bigmouth buffalo, so our extended research question will address if we observe major differences in estimated parameters between species using the same hierarchical models of catchability and predicted CPUE.

# Front Matter

```{r}
#Load in electrofishing CPUE data with 3 covariates (water temp, noon air temp, and noon wind speed)
#remove NA obs of water temp for 2 occasions
carp<-read.csv("NTwin18CPUEcarp.csv")
carp<-carp[!is.na(carp$Water.Temp..C.), ]
buff<-read.csv("NTwin18CPUEbuff.csv")
buff<-buff[!is.na(buff$Water.Temp..C.), ]
```

# define the models

```{r}
###############################################################
#
#     Carp Model
#
###############################################################
cpue.c <- jags.model( # cpue.c to denote carp model
  'NTWIN_Full.bug', # new model file
  data=list(logPop=log(3816), sdlogPop=.0081, # log pouplation size (from data: schnabel est) and 
                                               # log sd of population size from data (sdlogpop becomes Tau) 
    n=dim(carp)[1], # number samples (data)
    logCPUE=log(carp$CPUE), # log-transformed CPUE from data 
    AirTemp=carp$AirTemp, # air temp in celcius (Subract 20 in .bug file)
    Wind=carp$Wind, # wind speed in mph, (subtract 10 in .bug file)
    Water=carp$Water.Temp..C.), # water temp in Celcius
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method
  
coef(cpue.c)

###############################################################
#
#     Buffalo Model
#
###############################################################
cpue.b <- jags.model( # cpue.b to denote buffalo model
  'NTWIN_Full.bug', # same model file between species
  data=list(logPop=log(18776), sdlogPop=.0038, # log pouplation size (from data: schnabel est) and 
                                               # log sd of population size from data (sdlogpop becomes Tau) 
    n=dim(buff)[1], # number samples (data)
    logCPUE=log(buff$CPUE), # log-transformed CPUE from data 
    AirTemp=buff$AirTemp, # air temp in celcius (Subract 20 in .bug file)
    Wind=buff$Wind, # noon wind speed in mph (subracted 10 in .bug file)
    Water=buff$Water.Temp..C.), # water temp in Celcius
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method
  
coef(cpue.b)
```

# *CARP* coda.samples and MCMC simulations, diagnostic plots

```{r}
##########################################################################
#
#                 Carp
#
##########################################################################
update(cpue.c, 1000) # burn-in of 1000 samples

cpue.0 <- coda.samples(cpue.c, # model
  c('logN','b0','b1','b2','b3','sd','ElogCPUE'), # parameters we want back 
  1000) # number of samples

par(mar=c(3,3,2,2)+0.2, mgp=c(2,0.8,0))
plot(cpue.0) 
# A lot wackier trace plots than part 1. Increasing burn-in and samples



# adjustment to burn-in
update(cpue.c, 5000) # burn-in of 5000 samples

cpue.1 <- coda.samples(cpue.c, # model
  c('logN','b0','b1','b2','b3','sd', 'ElogCPUE'), # parameters we want back (posterior distributions of priors)
  5000) # number of samples

plot(cpue.1) 
# Much much better trace plots.  

summary(cpue.1)



# another model outputting logcatch

update(cpue.c, 5000) # burn-in of 5000 samples

cpue.extra <- coda.samples(cpue.c, # model
  c('logcatch'), # parameters we want back (posterior distributions of priors)
  5000) # number of samples


summary(cpue.extra)

exp(summary(cpue.extra[[1]][c(1:15)]))
```

## Part 2 *CARP* diagnostic plots

```{r,}
# extracting a parameter, plotting it's confidence region

# beta 0: intercept
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta0.out <- cbind(cpue.1[[1]][,16], cpue.1[[2]][,16], cpue.1[[3]][,16]) # subsetting for beta0 mcmc output
attributes(beta0.out) <- NULL # dropping attributes
dens.beta0 <- density(beta0.out) # density
q25.beta0 <- quantile(beta0.out, .025) # lower quantile
q975.beta0 <- quantile(beta0.out, .975) # upper quantile
dd.beta0 <- with(dens.beta0, data.frame(x,y)) # density dataframe

beta0<-qplot(x, y, data = dd.beta0, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta0, x>q25.beta0 & x<q975.beta0),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 0: intercept")+
  ylab("Density")+
  ggtitle("Density of Beta 0: intercept")
beta0
# does not overlap 0, -4 to -6.5



# beta 1: air temp
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta1.out <- cbind(cpue.1[[1]][,17], cpue.1[[2]][,17], cpue.1[[3]][,17]) # subsetting for beta1 mcmc output
attributes(beta1.out) <- NULL # dropping attributes
dens.beta1 <- density(beta1.out) # density
q25.beta1 <- quantile(beta1.out, .025) # lower quantile
q975.beta1 <- quantile(beta1.out, .975) # upper quantile
dd.beta1 <- with(dens.beta1, data.frame(x,y)) # density dataframe

beta1<-qplot(x, y, data = dd.beta1, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta1, x>q25.beta1 & x<q975.beta1),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 1: Slope of air temp")+
  ylab("Density")+
  ggtitle("Density of Beta 1: Air Temp")
beta1
# does not overlap 0, and actually very similar distribution to blue lake model (yay!)



# beta 2:  wind speed
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta2.out <- cbind(cpue.1[[1]][,18], cpue.1[[2]][,18], cpue.1[[3]][,18]) # subsetting for beta2 mcmc output
attributes(beta2.out) <- NULL # dropping attributes
dens.beta2 <- density(beta2.out) # density
q25.beta2 <- quantile(beta2.out, .025) # lower quantile
q975.beta2 <- quantile(beta2.out, .975) # upper quantile
dd.beta2 <- with(dens.beta2, data.frame(x,y)) # density dataframe

beta2<-qplot(x, y, data = dd.beta2, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta2, x>q25.beta2 & x<q975.beta2),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 2: Slope of wind speed")+
  ylab("Density")+
  ggtitle("Density of Beta 2: wind speed")
beta2
# overlaps 0, just like blue lake carp model. Good sign i suppose


#
# beta 3: WATER TEMP
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta3.out <- cbind(cpue.1[[1]][,19], cpue.1[[2]][,19], cpue.1[[3]][,19]) # subsetting for beta1 mcmc output
attributes(beta3.out) <- NULL # dropping attributes
dens.beta3 <- density(beta3.out) # density
q25.beta3 <- quantile(beta3.out, .025) # lower quantile
q975.beta3 <- quantile(beta3.out, .975) # upper quantile
dd.beta3 <- with(dens.beta3, data.frame(x,y)) # density dataframe

beta3<-qplot(x, y, data = dd.beta3, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta3, x>q25.beta3 & x<q975.beta3),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 3: water temp")+
  ylab("Density")+
  ggtitle("Density of Beta 3: water temp")
beta3
# does overlap 0, but less than 1. Interesting sauce. Wanted it to be more significant!

## Combine all linear beta plots
grid.arrange(beta0,beta1,
             beta2,beta3,
             ncol=2) # If I'm gonna do this plot in the presentation I need to standardize both axes
```


## Part 2 Buffalo Models


```{r,}

##########################################################################
#
#                 Buffalo
#
##########################################################################
update(cpue.b, 1000) # burn-in of 1000 samples

cpue.3 <- coda.samples(cpue.b, # model
  c('logN','b0','b1','b2','b3','sd',"ElogCPUE"), # parameters we want back 
  1000) # number of samples

plot(cpue.3)
# b1 and b3 look pretty whack, which is expected after the carp results.
# will try to increase burn-in and # resamples


# increase burn in 5 times
update(cpue.b, 5000) # burn-in of 5000 samples

cpue.4 <- coda.samples(cpue.b, # model
  c('logN','b0','b1','b2','b3','sd',"ElogCPUE"), # parameters we want back 
  5000) # number of samples

plot(cpue.4)
```


## Part 2 *BUFFALO* diagnostic plots

```{r,}
# extracting a parameter, plotting it's confidence region

# beta 0: intercept
varnames(cpue.4) # reference variables in model
str(cpue.4) # structure of lists
beta0.out <- cbind(cpue.4[[1]][,16], cpue.4[[2]][,16], cpue.4[[3]][,16]) # subsetting for beta0 mcmc output
attributes(beta0.out) <- NULL # dropping attributes
dens.beta0 <- density(beta0.out) # density
q25.beta0 <- quantile(beta0.out, .025) # lower quantile
q975.beta0 <- quantile(beta0.out, .975) # upper quantile
dd.beta0 <- with(dens.beta0, data.frame(x,y)) # density dataframe

beta0<-qplot(x, y, data = dd.beta0, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta0, x>q25.beta0 & x<q975.beta0),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 0: intercept")+
  ylab("Density")+
  ggtitle("Density of Carp Model Beta 0: intercept")+
  xlim(-10,0)+
  ylim(0,1)

beta0
# does not overlap 0, -5 to -9



# beta 1: air temp
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta1.out <- cbind(cpue.4[[1]][,17], cpue.4[[2]][,17], cpue.4[[3]][,17]) # subsetting for beta1 mcmc output
attributes(beta1.out) <- NULL # dropping attributes
dens.beta1 <- density(beta1.out) # density
q25.beta1 <- quantile(beta1.out, .025) # lower quantile
q975.beta1 <- quantile(beta1.out, .975) # upper quantile
dd.beta1 <- with(dens.beta1, data.frame(x,y)) # density dataframe

beta1<-qplot(x, y, data = dd.beta1, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta1, x>q25.beta1 & x<q975.beta1),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 1: Slope of air temp")+
  ylab("Density")+
  ggtitle("Density of Beta 1: Air Temp")
beta1
# overlaps 0! Different than carp model



# beta 2:  wind speed
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta2.out <- cbind(cpue.4[[1]][,18], cpue.4[[2]][,18], cpue.4[[3]][,18]) # subsetting for beta2 mcmc output
attributes(beta2.out) <- NULL # dropping attributes
dens.beta2 <- density(beta2.out) # density
q25.beta2 <- quantile(beta2.out, .025) # lower quantile
q975.beta2 <- quantile(beta2.out, .975) # upper quantile
dd.beta2 <- with(dens.beta2, data.frame(x,y)) # density dataframe

beta2<-qplot(x, y, data = dd.beta2, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta2, x>q25.beta2 & x<q975.beta2),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 2: Slope of wind speed")+
  ylab("Density")+
  ggtitle("Density of Beta 2: wind speed")
beta2
# overlaps 0, just like blue lake carp model and north twin carp model. Good sign i suppose


#
# beta 3: WATER TEMP
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta3.out <- cbind(cpue.4[[1]][,19], cpue.4[[2]][,19], cpue.4[[3]][,19]) # subsetting for beta1 mcmc output
attributes(beta3.out) <- NULL # dropping attributes
dens.beta3 <- density(beta3.out) # density
q25.beta3 <- quantile(beta3.out, .025) # lower quantile
q975.beta3 <- quantile(beta3.out, .975) # upper quantile
dd.beta3 <- with(dens.beta3, data.frame(x,y)) # density dataframe

beta3<-qplot(x, y, data = dd.beta3, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta3, x>q25.beta3 & x<q975.beta3),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 3: water temp")+
  ylab("Density")+
  ggtitle("Density of Beta 3: water temp")
beta3
# does overlap 0, but less than 1. Interesting sauce. Wanted it to be more significant!

## Combine all linear beta plots
grid.arrange(beta0,beta1,
             beta2,beta3,
             ncol=2) # If I'm gonna do this plot in the presentation I need to standardize both axes
```











# conclusions from the models and diagnostics:
- carp models behaved much the same as blue lake. Water temp wasn't significant
- seems that adding water temp covariate required much more simulations

- buffalo models are unique from carp, no slopes are significant.

- discouraging sauce.












*NEED TO HAVE GROUP DISCUSSION AND CONVERSATION WITH PHIL BEFORE PROCEEDING*

########################
#
#  Part 2.b the reduced model
#
#
########################







## define new reduced models

### changing "cpue.c" and "cpue.b" model structure, but keeping object name the same for following code (plots). *Note this change when going back to run code from any chunks above!*

```{r}
###############################################################
#
#     Carp Model
#
###############################################################
cpue.c <- jags.model( # cpue.c to denote carp model
  'NTWIN_Reduced.bug', # new model file
  data=list(logPop=log(3816), sdlogPop=.0081, # log pouplation size (from data: schnabel est) and 
                                               # log sd of population size from data (sdlogpop becomes Tau) 
    n=dim(carp)[1], # number samples (data)
    logCPUE=log(carp$CPUE), # log-transformed CPUE from data 
#    AirTemp=carp$AirTemp, # air temp in celcius (Subract 20 in .bug file)
#    Wind=carp$Wind, # wind speed in mph, (subtract 10 in .bug file)
    Water=carp$Water.Temp..C.), # water temp in Celcius
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method
  
coef(cpue.c)

###############################################################
#
#     Buffalo Model
#
###############################################################
cpue.b <- jags.model( # cpue.b to denote buffalo model
  'NTWIN_Reduced.bug', # same model file between species
  data=list(logPop=log(18776), sdlogPop=.0038, # log pouplation size (from data: schnabel est) and 
                                               # log sd of population size from data (sdlogpop becomes Tau) 
    n=dim(buff)[1], # number samples (data)
    logCPUE=log(buff$CPUE), # log-transformed CPUE from data 
#    AirTemp=buff$AirTemp, # air temp in celcius (Subract 20 in .bug file)
#    Wind=buff$Wind, # noon wind speed in mph (subracted 10 in .bug file)
    Water=buff$Water.Temp..C.), # water temp in Celcius
  n.chains=3, # number of simulation sequences
  n.adapt=100) # tells jags how much effort to spend in finding right resampling method
  
coef(cpue.b)
```

## coda.samples and diagnostic plots for reduced models



# *CARP* coda.samples and MCMC simulations, diagnostic plots

```{r}
##########################################################################
#
#                 Carp
#
##########################################################################
update(cpue.c, 1000) # burn-in of 1000 samples

cpue.0 <- coda.samples(cpue.c, # model
  c('logN','b0','b1','sd','ElogCPUE'), # parameters we want back 
  1000) # number of samples

par(mar=c(3,3,2,2)+0.2, mgp=c(2,0.8,0))
plot(cpue.0) 
# A lot wackier trace plots than part 1. Increasing burn-in and samples



# adjustment to burn-in
update(cpue.c, 5000) # burn-in of 5000 samples

cpue.1 <- coda.samples(cpue.c, # model
  c('logN','b0','b1','sd', 'ElogCPUE'), # parameters we want back (posterior distributions of priors)
  5000) # number of samples

plot(cpue.1) 
# Much much better trace plots.  

summary(cpue.1)
```

## Part 2 *CARP* diagnostic plots

```{r,}
# extracting a parameter, plotting it's confidence region

# beta 0: intercept
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta0.out <- cbind(cpue.1[[1]][,16], cpue.1[[2]][,16], cpue.1[[3]][,16]) # subsetting for beta0 mcmc output
attributes(beta0.out) <- NULL # dropping attributes
dens.beta0 <- density(beta0.out) # density
q25.beta0 <- quantile(beta0.out, .025) # lower quantile
q975.beta0 <- quantile(beta0.out, .975) # upper quantile
dd.beta0 <- with(dens.beta0, data.frame(x,y)) # density dataframe

beta0.c<-qplot(x, y, data = dd.beta0, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta0, x>q25.beta0 & x<q975.beta0),
              aes(ymax = y), ymin = 0, fill = "darkgoldenrod", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 0: intercept")+
  ylab("Density of Beta 0")+
  ggtitle("Density of N Twin Carp Model: Intercept")+  
  xlim(-10,0)+
  ylim(0,1)

beta0.c
# does not overlap 0, -4 to -6.5



# beta 1: water temp
varnames(cpue.1) # reference variables in model
str(cpue.1) # structure of lists
beta1.out <- cbind(cpue.1[[1]][,17], cpue.1[[2]][,17], cpue.1[[3]][,17]) # subsetting for beta1 mcmc output
attributes(beta1.out) <- NULL # dropping attributes
dens.beta1 <- density(beta1.out) # density
q25.beta1 <- quantile(beta1.out, .025) # lower quantile
q975.beta1 <- quantile(beta1.out, .975) # upper quantile
dd.beta1 <- with(dens.beta1, data.frame(x,y)) # density dataframe

beta1.c<-qplot(x, y, data = dd.beta1, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta1, x>q25.beta1 & x<q975.beta1),
              aes(ymax = y), ymin = 0, fill = "darkgoldenrod", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 1: Slope of water temp")+
  ylab("Density of Beta 1")+
  ggtitle("N Twin Carp Model: Slope Water Temp")+
  ylim(0,13)+
  xlim(-0.35,0.35)
beta1.c
# overlaps 0!



```


# *Buffalo* coda.samples and MCMC simulations, diagnostic plots

```{r}
##########################################################################
#
#                 Buffalo
#
##########################################################################
update(cpue.b, 1000) # burn-in of 1000 samples

cpue.2 <- coda.samples(cpue.b, # model
  c('logN','b0','b1','sd','ElogCPUE'), # parameters we want back 
  1000) # number of samples

par(mar=c(3,3,2,2)+0.2, mgp=c(2,0.8,0))
plot(cpue.2) 
# A lot wackier trace plots than part 1. Increasing burn-in and samples



# adjustment to burn-in
update(cpue.b, 5000) # burn-in of 5000 samples

cpue.3 <- coda.samples(cpue.b, # model
  c('logN','b0','b1','sd', 'ElogCPUE'), # parameters we want back (posterior distributions of priors)
  5000) # number of samples

plot(cpue.3) 
# Much much better trace plots.  

summary(cpue.3)
```

## Part 2 *Buffalo* diagnostic plots

```{r,}
# extracting a parameter, plotting it's confidence region

# beta 0: intercept
varnames(cpue.3) # reference variables in model
str(cpue.3) # structure of lists
beta0.out <- cbind(cpue.3[[1]][,16], cpue.3[[2]][,16], cpue.3[[3]][,16]) # subsetting for beta0 mcmc output
attributes(beta0.out) <- NULL # dropping attributes
dens.beta0 <- density(beta0.out) # density
q25.beta0 <- quantile(beta0.out, .025) # lower quantile
q975.beta0 <- quantile(beta0.out, .975) # upper quantile
dd.beta0 <- with(dens.beta0, data.frame(x,y)) # density dataframe

beta0.b<-qplot(x, y, data = dd.beta0, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta0, x>q25.beta0 & x<q975.beta0),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 0: intercept")+
  ylab("Density of Beta 0")+
  ggtitle("N Twin Buffalo Model: intercept")+
  xlim(-10,0)+
  ylim(0,1)

beta0.b
# does not overlap 0, -5.5 to -7.5



# beta 1: water temp
varnames(cpue.3) # reference variables in model
str(cpue.3) # structure of lists
beta1.out <- cbind(cpue.3[[1]][,17], cpue.3[[2]][,17], cpue.3[[3]][,17]) # subsetting for beta1 mcmc output
attributes(beta1.out) <- NULL # dropping attributes
dens.beta1 <- density(beta1.out) # density
q25.beta1 <- quantile(beta1.out, .025) # lower quantile
q975.beta1 <- quantile(beta1.out, .975) # upper quantile
dd.beta1 <- with(dens.beta1, data.frame(x,y)) # density dataframe

beta1.b<-qplot(x, y, data = dd.beta1, geom = "line", ylab = "", xlab = "") +
  geom_ribbon(data = subset(dd.beta1, x>q25.beta1 & x<q975.beta1),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA, alpha = 0.5)+
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Beta 1: Slope of water temp")+
  ylab("Density of Beta 1")+
  ggtitle("N Twin Buffalo Model: Slope Water Temp")+
  ylim(0,13)+
  xlim(-0.35,0.35)
beta1.b
# does not overlap 0, and actually very similar distribution to blue lake model (yay!)



```




```{r}
grid.arrange(beta0.b,beta1.b,
             beta0.c,beta1.c, ncol = 2)
```










# Results plots

## linear regression of log(cpue) as a function of log(water temp)

```{r}
##############################################################################
#
#               Carp linear regression
#
##############################################################################
cpue.1.out<-summary(cpue.1)
varnames(cpue.1.out)
str(cpue.1.out)
linear1 <- data.frame(cpue.1.out[[1]][c(16,17),], cpue.1.out[[2]][c(16,17),c(1,5)])
str(linear1)

catch.int<-cpue.1.out[[1]][16,1]
catch.slope<-cpue.1.out[[1]][17,1]


m <- cpue.1.out$statistics[,"Mean"]
l <- cpue.1.out$quantiles[,"2.5%"]
u <- cpue.1.out$quantiles[,"97.5%"]

dim(cpue)
length(m)

#
# Plot of linear regression, x = ln(water temp) and y = ln(CPUE)
#

carp.lin<-ggplot(carp, aes(x = log(Water.Temp..C.), y = log(CPUE/3816))) +
  geom_point(size = 3, color = "darkgoldenrod") +
  geom_abline(slope = catch.slope, intercept = catch.int, size = 1.5, lty = 2, color = "darkgoldenrod")+
  xlim(1,4)+
  ylim(-10,-4)+
  geom_ribbon(aes(ymin = catch.int-l[1], ymax = catch.int+u[1]), alpha = 0.2)+ # can't get this to work
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("ln (Water Temperature)")+
  ylab("ln (Catchability)")+
  ggtitle("N Twin Model of Ln Carp CPUE as a function of Ln water temperature")
carp.lin

##############################################################################
#
#               Buffalo linear regression
#
##############################################################################
cpue.3.out<-summary(cpue.3)
varnames(cpue.3.out)
str(cpue.3.out)
linear2 <- data.frame(cpue.3.out[[1]][c(16,17),], cpue.3.out[[2]][c(16,17),c(1,5)])
str(linear2)

catch.int<-cpue.3.out[[1]][16,1]
catch.slope<-cpue.3.out[[1]][17,1]


m <- cpue.3.out$statistics[,"Mean"]
l <- cpue.3.out$quantiles[,"2.5%"]
u <- cpue.3.out$quantiles[,"97.5%"]

dim(cpue)
length(m)

#
# Plot of linear regression, x = ln(water temp) and y = ln(CPUE)
#

buff.lin<-ggplot(buff, aes(x = log(Water.Temp..C.), y = log(CPUE/18776))) +
  geom_point(size = 3, color = "red") +
  geom_abline(slope = catch.slope, intercept = catch.int, size = 1.5, lty = 2, color = "red")+
  xlim(1,4)+
  ylim(-10,-4)+
  geom_ribbon(aes(ymin = catch.int-l[1], ymax = catch.int+u[1]), alpha = 0.2)+ # can't get this to work
  theme_classic()+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))+
  xlab("ln (Water Temperature)")+
  ylab("ln (Catchability)")+
  ggtitle("N Twin Model of Ln Buffalo CPUE as a function of Ln Water Temperature")
buff.lin

grid.arrange(carp.lin,buff.lin,ncol=2)
```

## Comparison of expected log CPUE vs. observed log CPUE, by date

```{r}
#####################################################################
#
#                    CARP Plot of expected vs. observed cpue
#
######################################################################

# plot of distribution of ElogCPUE in boxplots, overlaid with observed CPUE 
cpue.4 <- coda.samples(cpue.c, 
                       c('ElogCPUE'), # parameters we want back
                       5000) # number of samples
summary(cpue.4)
# storing model output as object
cpue.4.out<-summary(cpue.4)

cpue.plot.data.carp<-cbind(carp,cpue.4.out[[1]][c(1:15),],cpue.4.out[[2]][c(1:15),c(1:5)])
str(cpue.plot.data.carp)

Carp.EvO.CPUE<-ggplot(cpue.plot.data.carp, aes(x = Date, y = log(CPUE)))+
         geom_boxplot(aes(ymin = cpue.plot.data.carp$`2.5%`, lower = cpue.plot.data.carp$`25%`, 
                           middle = cpue.plot.data.carp$Mean, 
                           upper = cpue.plot.data.carp$`75%`, ymax = cpue.plot.data.carp$`97.5%`),
  stat = "identity", fill = "darkgoldenrod")+
  geom_point(size=5)+
  theme_classic()+
  theme(axis.text=element_text(size=8),
        axis.text.x = element_text(angle=90),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Date")+
  ylab("Expected and observed ln (CPUE)")+
  ggtitle("Comparison of Expected Ln Carp CPUE Distribution vs. Observed Ln Carp CPUE")+
  ylim(0,5)
Carp.EvO.CPUE
####################################################################################################

#####################################################################
#
#                    Buffalo Plot of expected vs. observed cpue
#
######################################################################

# plot of distribution of ElogCPUE in boxplots, overlaid with observed CPUE 
cpue.5 <- coda.samples(cpue.b, 
                       c('ElogCPUE'), # parameters we want back
                       5000) # number of samples
summary(cpue.5)
# storing model outputas object
cpue.5.out<-summary(cpue.5)

cpue.plot.data.buff<-cbind(buff,cpue.5.out[[1]][c(1:15),],cpue.5.out[[2]][c(1:15),c(1:5)])
str(cpue.plot.data.buff)

Buff.EvO.CPUE<-ggplot(cpue.plot.data.buff, aes(x = Date, y = log(CPUE)))+
         geom_boxplot(aes(ymin = cpue.plot.data.buff$`2.5%`, lower = cpue.plot.data.buff$`25%`, 
                           middle = cpue.plot.data.buff$Mean, 
                           upper = cpue.plot.data.buff$`75%`, ymax = cpue.plot.data.buff$`97.5%`),
  stat = "identity", fill = "red", alpha = 0.5)+
  geom_point(size=5)+
  theme_classic()+
  theme(axis.text=element_text(size=8),
        axis.text.x = element_text(angle=90),
        axis.title=element_text(size=14,face="bold"))+
  xlab("Date")+
  ylab("Expected and observed ln (CPUE)")+
  ggtitle("Comparison of Expected Ln Buffalo CPUE Distribution vs. Observed Ln Buffalo CPUE")+
  ylim(0,5)
Buff.EvO.CPUE
####################################################################################################


grid.arrange(Carp.EvO.CPUE,Buff.EvO.CPUE,ncol=2)


```

# END
