---
title: "2020 rmark input file"
author: "Martin Simonson"
date: "2/5/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r}
rm(list=ls())
library(RMark)
library(dplyr)
library(tidyr)
setwd("~/BMB-COC/MARK/2020")
dframe<-read.csv("2017-2020 Compiled Data.csv",header=T)
dframe$Date<-as.Date(dframe$Date, "%m/%d/%Y")
dframe$Year.Month<-factor(format(dframe$Date, "%Y.%m"))
dframe$Year<-factor(format(dframe$Date, "%Y"))
dframe


str(dframe)
```

# getting capture history together

```{r}
# for the moment, let's just do center lake 

# and for the moment buffalo only:

df<-droplevels(subset(dframe, Lake == "Center" &
                          Species == "BIB" &
                          !is.na(dframe$TAG.NUMBER)))
# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))
levels(df$Notes)



# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.b.ch.year<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Year - capture histories of buff condensed to years

# making it work for months over years (robust design)
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.b.ch.month<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Month - capture histories of buff condensed to Months
c.b.ch.month$freq<-1

############
#
# Carp
#
###########
df<-droplevels(subset(dframe, Lake == "Center" &
                          Species == "COC" ))
# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))




# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.c.ch.year<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Year - capture histories of carp condensed to years

# condense to tag number, month, and detect
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.c.ch.month<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Month - capture histories of carp condensed to months
c.c.ch.month$freq<-1
# attempt at saving to try in MARK itself
write.csv(c.b.ch.month, "CenterBuffRD.inp",row.names = F)
write.csv(c.c.ch.month, "CenterCarpRD.inp",row.names=F)
```


# robust design models with monthly capture histories - Center Lake Buffalo - probably worth deleting now

```{r}


run.robust=function(x)
{
#
# data from Robust.dbf with MARK
# 5 primary sessions with secondary sessions of length 2,6,6
#
time.intervals=c(0, 1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
#
#

rd <- process.data(data = x,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)
rd$nocc
rd$nocc.secondary

S.time=list(formula=~time)
p.time.session=list(formula=~-1+session:time,share=TRUE)
model.1=mark(data = rd, model = "Robust", 
            time.intervals=time.intervals,
            model.parameters=list(S=S.time,
                                  p=p.time.session,
                                  Gamma` = 0,
                                  Gamma`` = 0,
                                  ),threads=2)

S.time.session=list(formula=~-1+session:time,share=TRUE)
p.time.session=list(formula=~-1+session:time,share=TRUE)
model.2=mark(data = rd, model = "Robust",
             time.intervals=time.intervals,
             model.parameters=list(S=S.time,
                                   p=p.time.session,
                                   ),threads=2)

S.time.session=list(formula=~1,share=TRUE)
p.time.session=list(formula=~1,share=TRUE)
model.3=mark(data=rd,model="Robust",
             time.intervals=time.intervals,
             model.parameters = list(S = S.time.session,
                                     p = p.time.session),threads=2)


return(collect.models())
}
c.b.robust.results=run.robust(c.b.ch.month)

c.b.robust.results

cleanup(ask=F)
```


# University of Montana edition, on Center Lake Buffalo

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6
#
time.intervals=c(0, 1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
#

rd <- process.data(data = c.b.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}
# survival varies by time, recapture prob varies by month and year
Center.Buff.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
Center.Buff.Params
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
Center.Buff.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)
Center.Buff.Nhat
# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)

```


# Center Lake Carp

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6
#
time.intervals=c(0, 1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
#

rd <- process.data(data = c.c.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}
# survival varies by time, recapture prob varies by month and year
Center.Carp.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
Center.Carp.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)

```

# move on to North Twin Lake input files

## getting capture history together

```{r}
# for the moment, let's just do center lake 

# and for the moment buffalo only:

df<-droplevels(subset(dframe, Lake == "N Twin" &
                          Species == "BIB" &
                          !is.na(dframe$TAG.NUMBER)))

# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))
levels(df$Notes)



# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.b.ch.year<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Year - capture histories of buff condensed to years

# making it work for months over years (robust design)
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
nt.b.ch.month<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Month - capture histories of buff condensed to Months
nt.b.ch.month$freq<-1

############
#
# Carp
#
###########
df<-droplevels(subset(dframe, Lake == "N Twin" &
                          Species == "COC" ))
# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))

levels(df$Year.Month)


# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.c.ch.year<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Year - capture histories of carp condensed to years

# condense to tag number, month, and detect
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
nt.c.ch.month<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Month - capture histories of carp condensed to months
nt.c.ch.month$freq<-1
# attempt at saving to try in MARK itself
write.csv(nt.b.ch.month, "NTwinBuffRD.inp",row.names = F)
write.csv(nt.c.ch.month, "NTwinCarpRD.inp",row.names=F)
```

# University of Montana edition, on North Twin Lake Buffalo

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6 - same as center lake
#
levels(df$Year.Month)
time.intervals=c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
#

rd <- process.data(data = nt.b.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies by eyar, recapture varies in season and in year
   # mod.Stime.pTime = mark(data = rd, 
    #                               model.parameters = list(
     #                               S = S.time,
      #                              p = p.time,
       #                             GammaDoublePrime = GammaDoublePrime.zero,
        #                            GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)

# survival varies by time, recapture prob varies by month and year
NTwin.Buff.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
NTwin.Buff.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

```


## North Twin Lake Carp

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6
#

time.intervals=c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
#

rd <- process.data(data = nt.c.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}
# survival varies by time, recapture prob varies by month and year
NTwin.Carp.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
NTwin.Carp.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)
```

# South Twin Lake

## getting capture history together

```{r}
# for the moment, let's just do center lake 

# and for the moment buffalo only:

df<-droplevels(subset(dframe, Lake == "S Twin" &
                          Species == "BIB" &
                          !is.na(dframe$TAG.NUMBER)))

# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))
levels(df$Notes)

df3<-levels(df$Year.Month)


# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.b.ch.year<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Year - capture histories of buff condensed to years

# making it work for months over years (robust design)
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
st.b.ch.month<-capt.hist[,2] # code: Center.Buffalo.CaptureHistory.Month - capture histories of buff condensed to Months
st.b.ch.month$freq<-1

############
#
# Carp
#
###########
df<-droplevels(subset(dframe, Lake == "S Twin" &
                          Species == "COC" ))
# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))

levels(df$Year.Month)


# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.c.ch.year<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Year - capture histories of carp condensed to years

# condense to tag number, month, and detect
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
st.c.ch.month<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Month - capture histories of carp condensed to months
st.c.ch.month$freq<-1
# attempt at saving to try in MARK itself
write.csv(st.b.ch.month, "StBuffRD.inp",row.names = F)
write.csv(st.c.ch.month, "StCarpRD.inp",row.names=F)
```

# University of Montana edition, on South Twin Lake Buffalo

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6 - same as center lake
#
df3
time.intervals=c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)

#

rd <- process.data(data = st.b.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
   # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                  model.parameters = list(
                                   S = S.time,
                                   p = p.time,
                                   GammaDoublePrime = GammaDoublePrime.zero,
                                   GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)

# survival varies by time, recapture prob varies by month and year
STwin.Buff.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
STwin.Buff.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

```


## South Twin Lake Carp

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6
#

time.intervals=c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
#

rd <- process.data(data = st.c.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
  mod.dot = mark(data = rd,
                    model.parameters = list(
                      S = S.dot,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}
# survival varies by time, recapture prob varies by month and year
STwin.Carp.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
Stwin.Carp.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)
```

# Blue Lake

## getting capture history together

```{r}

############
#
# Carp
#
###########
df<-droplevels(subset(dframe, Lake == "Blue" &
                          Species == "COC" ))
# remove dissections and others
df$TAG.NUMBER[df$TAG.NUMBER == ""]<-NA
df<-droplevels(subset(df, !is.na(as.numeric(as.character(df$TAG.NUMBER)))))

levels(df$Year.Month)


# condense to year and tag number and detect

df2<-df[,c(3,21,22)]


capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
c.c.ch.year<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Year - capture histories of carp condensed to years

# condense to tag number, month, and detect
df2<-df[,c(3,23,22)]
capt.hist<-df2 %>%
  # remove duplicates when fish are caught more than once in an event
  distinct() %>%
  # spread out data. Fill = 0 adds rows for combinations of ID and event w/o observations
  spread(Year.Month, detect, fill = 0) %>%
  # for every individual
  group_by(TAG.NUMBER) %>%
  # paste together 0s and 1s
  # unite is similar to paste
  unite("ch",2:tail(names(.),1),sep = "")

# well that worked, now to make it a .inp file
b.c.ch.month<-capt.hist[,2] # code: Center.Carp.CaptureHistory.Month - capture histories of carp condensed to months
b.c.ch.month$freq<-1
# attempt at saving to try in MARK itself
write.csv(b.c.ch.month, "StCarpRD.inp",row.names=F)
```

# University of Montana edition, on blue lake carp

```{r,}
# set up time intervals for 5 primary occasions each with 3 secondary occasions
# 5 primary sessions with secondary sessions of length 2,6,6 - same as center lake
#
levels(df$Year.Month)
time.intervals=c(0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)

#

rd <- process.data(data = b.c.ch.month,
                  model = "Robust",
                  time.intervals = time.intervals)
                  
names(rd)

# take a look at the data
head(rd$data)

# check the number of primary occasions 
rd$nocc

# check the number of secondary occasions 
rd$nocc.secondary
```

### Set up models to run

we want only specific combinations of S, p, gamma' and gamma'' and so need to dictate which ones we want one model at a time.

```{r}
run.robust <- function() {

  # In all models, gamma' and gamma'' are fixed at 0 (no temporary emigration)
  
  # 2 variations of survival
  
  # 3 variations of recapture prob
  
  # Apparent survival varies by year
  S.time = list(formula =  ~ time)
  # Apparent survival is constant
  S.dot = list(formula = ~ 1)
  
  # p varies by primary session but not among secondaries within primary
  # session = varies by primary, time=varies by secondary
  p.session = list(formula =  ~ session, share = TRUE)
  # session:time = varies by both
  p.session.time = list(formula =  ~ session:time, share = TRUE)
  # p=c due to use of "share=TRUE"
  p.dot = list(formula = ~ 1, share = TRUE)
  p.time = list(formula = ~ time, share = T)
  
  # we are excluding temporary emigration, fixing these at 0 for all models
  GammaDoublePrime.zero = list(formula =  ~ 1,
                                 fixed = 0)
  GammaPrime.zero = list(formula = ~ 1,
                         fixed = 0)
  

  # Model 1: Dot Model for survival and recapture prob
 # mod.dot = mark(data = rd,
  #                  model.parameters = list(
   #                   S = S.dot,
    #                  p = p.dot,
     #                 GammaDoublePrime = GammaDoublePrime.zero,
      #                GammaPrime = GammaPrime.zero))
  
  # Model 2: Survival varies by year, recapture is constant
  
  mod.Stime.pDot = mark(data = rd,
                    model.parameters = list(
                      S = S.time,
                      p = p.dot,
                      GammaDoublePrime = GammaDoublePrime.zero,
                      GammaPrime = GammaPrime.zero))
  
  # model 3: survival varies by all periods, recapture is null (dot)
  
    #mod.Ssessiontime.p.dot = mark(data = rd, 
     #                         model.parameters = list(
      #                         S = S.session.time,
       #                        p = p.dot,
        #                       GammaDoublePrime = GammaDoublePrime.zero,
         #                      GammaPrime = GammaPrime.zero))
           
  # Model 4 survival is constant, recapture varies within years but is same between years?
       # I don't think this makes sense, unequal sample periods
  
  mod.Sdot.pSession = mark(data = rd, 
                      model.parameters = list(
                         S = S.dot,
                         p = p.session,
                         GammaDoublePrime = GammaDoublePrime.zero,
                         GammaPrime = GammaPrime.zero))
  
  # model 5: survival is constant, recapture is constant within on year but varies between years
  mod.Sdot.pTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  
  
  # model 6: survival is constant, recapture varies within years and between years
  mod.Sdot.pSessionTime = mark(data = rd, 
                              model.parameters = list(
                               S = S.dot,
                               p = p.session.time,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
  
  # model 7: survival varies by years, recapture varies by session but not year
    mod.Stime.pSession = mark(data = rd, 
                              model.parameters = list(
                               S = S.time,
                               p = p.session,
                               GammaDoublePrime = GammaDoublePrime.zero,
                               GammaPrime = GammaPrime.zero))
    
   # model 8: survival varies by eyar, recapture varies in season and in year
    mod.Stime.pTime = mark(data = rd, 
                                  model.parameters = list(
                                   S = S.time,
                                   p = p.time,
                                   GammaDoublePrime = GammaDoublePrime.zero,
                                   GammaPrime = GammaPrime.zero))
    
  # model 8: survival varies year, recapture varies in season and in year
    mod.Stime.pSessionTime = mark(data = rd, 
                                   model.parameters = list(
                                    S = S.time,
                                    p = p.session.time,
                                    GammaDoublePrime = GammaDoublePrime.zero,
                                    GammaPrime = GammaPrime.zero))
    

# Return model table and list of models
return(collect.models())
}
```

### Run the models and examine the output

It's very simple to then use the function to run each of the models. As implemented here, the output from each of the models appears in the console where it can be reviewed. One can also examine model-specific output in other ways as shown below. 

```{r}
robust.results=run.robust()
```

### Examine model-selection table

```{r}
robust.results
names(robust.results)
```

### Examine output from top two models

```{r}

# survial is dot, recapture is varied by month and year
round(robust.results$mod.Sdot.pSessionTime$results$real[, 1:4], 3)
round(robust.results$mod.Sdot.pSessionTime$results$derived$`N Population Size`, 3)

# survival varies by time, recapture prob varies by month and year
Blue.Carp.Params<-round(robust.results$mod.Stime.pSessionTime$results$real[, 1:4], 3)
# Survival for two intervals, capture prob for 14 intervals, number unmarked in three years
Blue.Carp.Nhat<-round(robust.results$mod.Stime.pSessionTime$results$derived$`N Population Size`, 3)

```


##################################################################################################################################
#
## now to get some plots together maybe

The model that we are interested in, and was one of top 2 models for each lake and species, is:

mod.Stime.pSessionTime

- Need to compile objects of params above

```{r}

BuffSurvival<-rbind(Center.Buff.Params[c(1,2),],
                    NTwin.Buff.Params[c(1,2),],
                    STwin.Buff.Params[c(1,2),])
BuffSurvival$Lake<-as.factor(c("Center Lake","Center Lake","North Twin Lake","North Twin Lake","South Twin Lake","South Twin Lake"))
BuffSurvival$Year<-as.factor(as.character(c("2017","2018","2017","2018","2017","2018")))

CarpSurvival<-rbind(Center.Carp.Params[c(1,2),],
                    NTwin.Carp.Params[c(1,2),],
                    STwin.Carp.Params[c(1,2),],
                    Blue.Carp.Params[c(1,2),])
CarpSurvival$Lake<-as.factor(c("Center Lake","Center Lake",
                               "North Twin Lake","North Twin Lake",
                               "South Twin Lake","South Twin Lake",
                               "Blue Lake","Blue Lake"))
CarpSurvival$Year<-as.factor(as.character(c("2017","2018","2017","2018","2017","2018","2017","2018")))

```

### plotting survival:

side-by-side boxplot, or using lattice to split Open vs. Winter
- potential bargraph but those are whack sauce

survival (phi) on y axis
Year on x axis




Now to create 2 carp plots, 
- Carp actual top model (Phi(Center(.),N=S(.)) p (g*t))
```{r,}
library(ggplot2)
# Carp FIrst
carp1<-ggplot(CarpSurvival, aes(x=Year, y=estimate)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = ucl, ymin = lcl),width=0.2) +
  labs(title = "Common Carp Survival",
       x = "Year",
       y = "Survival") +
   facet_grid(Lake ~ .)+
        theme(panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.background = element_rect(fill="white",colour="grey50"), 
                      axis.line = element_line(colour = "black"),
                      legend.position = "none")
carp1


```

```{r}

# Buffalo
Buff1<-ggplot(BuffSurvival, aes(x=Year, y=estimate)) +
  geom_point(stat="identity", size = 4) +
  scale_y_continuous(limits=c(0,1))+
  geom_errorbar(aes(ymax = ucl, ymin = lcl), width = 0.2) +
  labs(title = "Buffalo Survival",
       x = "Year",
       y = "Survival") +
  facet_grid(Lake ~ .)+
        theme(panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.background = element_rect(fill="white",colour="grey50"), 
                      axis.line = element_line(colour = "black"),
                      legend.position = "none")
Buff1

```


### considerations:

- no accounting for tag loss
- low low buffalo recaptures
- add lakes where 2018 survival is available?






























################ new new

# 2020 input files, split by species, all lakes, censoring tags that get harvested. 

```{r}
# carp first:
c.c.ch.month$Lake <- "Center"
nt.c.ch.month$Lake <- "North Twin"
st.c.ch.month$Lake<- "South Twin"
b.c.ch.month$Lake<- "Blue"

carp.ch.month<-rbind(c.c.ch.month,
                     nt.c.ch.month,
                     st.c.ch.month,
                     b.c.ch.month)

```

