---
title: 'Preliminary Exams: Dr. Kaiser'
subtitle: 'Revised Submission'
author: "Marty Simonson"
date: "4/15/2020"
output: word_document
editor_options: 
  chunk_output_type: console
---


```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}


#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
#packages(FSA)
#packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
#packages(arm)
#packages(lattice)
#packages(tikzDevice)
#packages(FSAdata)
#packages(nlstools)
#packages(GGally)
#packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(tidyr)
#packages(rcompanion)
#packages(GISTools)
#packages(maps)
#packages(mapproj)
packages(fBasics)
packages(scales)
```

# The generalized Gamma Distribution

Suppose that $X$ has a standard gamma distribution with the probability density function, for $\alpha > 0$ and $\beta > 0$,

\[
f(x|\alpha,\beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1}e^{-\beta x}; x>0
\]

And a generalized gamma distribution results from the transformation

\[
Y = X^c
\]

where $c$ is a known constant such that $c>0$. Larger values of $c$ result in more extreme right tails in the resulting density function.

## 1.) Derive the P.D.F. of $Y$ using the Jacobian method

The general formula for the Jacobian transformation is 

\[
pdf(Y) = | \frac{\partial x}{\partial y} | * pdf(X)
\]

where the pdf of $X$ is given above. And if $Y=X^c$, then $X=Y^\frac{1}{c}$. The Jacobian determinant of this term is $| \frac{y^{(\frac{1}{c-1})}}{c} |$, and $y^\frac{1}{c}$ can be substituted in the distribution function for $x$, thus our new equation is:

\[
pdf(Y) = |\frac{y^{(\frac{1}{c})-1}}{c}| * \frac{\beta^\alpha}{\Gamma(\alpha)} * (y^\frac{1}{c})^{\alpha-1} * e^{(-\beta *y^\frac{1}{c})}
\]

Which, for positive nonzero $\alpha$, $\beta$, and $c$, simplifies to:

\[
pdf(Y) = \frac{ \beta^\alpha * 
          y^{\frac{\alpha-c}{c}} * 
          e^{(\beta*-y^{(1/c)})}}
          {c * \Gamma(\alpha)}
\]

for all values of $y \geq 0$, 0 elsewhere.

## 2.) Sample Statistics

Simulate a set of data of size $n=100$ from a generalized gamma distribution with $\alpha=9$, $\beta = 3$, and $c=3.25$. Note that this is most easily accomplished by first simulating the set of data from a gamma distribution with $\alpha = 10$ and $\beta=3$. 

```{r,}
# Simulating Data

## Original Gamma Distribution
n<-10000 # sample size of data
alpha<-10 # shape parameter
beta<-3 # scale parameter
gamma.sim<-data.frame(rgamma(n,shape=alpha,scale=beta))
colnames(gamma.sim)<-c("Value")# set column name for our 100 simulated values.

hist(gamma.sim$Value, prob=T, main = "Gamma, Alpha = 10 and Beta = 3",xlab="Variate value (n=10,000)", ylab = "density")

```


### For this example, first use the _original sample of the gamma distribution_ to compute:

- The sample mean
- Sample variance
- Skewness

```{r}
# Sample Mean
mean(gamma.sim$Value)
# Variance
var(gamma.sim$Value)
# Skewness
s<-skewness(gamma.sim$Value)
s[[1]]
```


### Second, use the _generalized gamma function_ to compute:

- The sample mean
- Sample variance
- Skewness

```{r}
# Mew simulated data - Generalized gamma distribution
f<-function(n,c.gen=3.25,beta.gen=3,alpha.gen=10){
 (rgamma(n,shape=alpha.gen,rate=beta.gen))^c.gen
}
n=10000
gen.gamma.sim<-data.frame(f(n))

# Sample Mean
mean(gen.gamma.sim$f.n.)
# Variance
var(gen.gamma.sim$f.n.)
# Skewness
s2<-skewness(gen.gamma.sim$f.n.)
s2[[1]]

hist(gen.gamma.sim$f.n., prob=T, main = "Generalized Gamma: Alpha = 10, Beta = 3, c = 3.25", xlab = "Variate value (n=10,000)",
     ylab="Density")
```



## 3.) Plot Empirical vs. Theoretical densities

Graph the empirical density of the generalized gamma sample (histogram on a probability scale) and overlay a curve giving the theoretical probability density function that was derived in part 1. 

__The area shaded green is the density curve for the simulated (empirical) data, the red line (without very light shading) is the density curve for the theoretical density from the derived formula.__

```{r,}
# paramaters
c.gen<-3.25
alpha.gen<-9
beta.gen<-3
# max value
n<-700
# sequence of values
values<-seq(0.1,n,.01) 
#values<-runif(100,0,700) ---- ended up providing the same curve

# break down PDF(y) into parts
prt1<-(beta.gen^alpha.gen)*values^((alpha.gen-c.gen)/c.gen)
prt2<-exp(beta.gen*(-values^(1/c.gen)))
prt3<-c.gen*gamma(alpha.gen)
pdfy<-data.frame(values,((prt1*prt2)/prt3))

# plot both transformed distributions
gen.gamma.sim %>%
  ggplot( aes(x=f.n.)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.9)+
      theme(legend.position = "bottom",
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            panel.background = element_rect(fill="white",
                                            colour="grey50"), 
            axis.line = element_line(colour = "black"))+
  labs(title="Densities of Transformed Gamma R.V.s",
       x="Value",
       y="Density")+
  geom_ribbon(data=pdfy,
            aes(x=values,ymin=0,ymax=X..prt1...prt2..prt3.),
            color="red",fill="red",alpha=0.1)
```


