---
title: "Scientific Collection Permit Reporting"
author: "Martin Simonson"
date: "1/6/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r,include=FALSE}
rm(list=ls())


packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#remove.packages(c("ggplot2", "data.table"))
#install.packages('Rcpp', dependencies = TRUE)
#install.packages('ggplot2', dependencies = TRUE)
#install.packages('data.table', dependencies = TRUE)

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
#packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
#packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
#packages(GISTools)
#packages(maps)
#packages(mapproj)
```

# code from "2019 annual report.rmd"

# DNR Permit data

```{r, include = F}
datframe<-read.csv("2019 Carp and Buffalo Data Entry - Jan 2020.csv", header = T)
str(datframe)



# Lakes
levels(datframe$Lake)
datframe$Lake[datframe$Lake == "5 Island "] <- "5 Island"
datframe$Lake[datframe$Lake == "N. Twin"] <- "N Twin"
datframe$detect<-1
junk<-melt(datframe, id.var = c("Species","Lake"),
           measure.vars = "detect")
LakeTotals<-dcast(junk,Species~Lake)
# this is good for overall total catch, HOWEVER:
#   For DNR permit (and probably subsequent IACUC) 
#   I need to split by released alive and euthanized
str(LakeTotals)




levels(datframe$Notes)
length(datframe$Species[datframe$Notes == "euthanized"])
datframe$Notes[datframe$Notes == "euthanized -  cut tail"] <- "euthanized"
datframe$Notes<-droplevels(datframe$Notes)
levels(datframe$Notes)
length(datframe$Species[datframe$Notes == "euthanized"])

# standardize/fix species names
levels(datframe$Species)
datframe$Species<-as.character(datframe$Species)
datframe$Species[datframe$Species == "Bullhead"] <- "BBH" # Brown Bullhead
datframe$Species[datframe$Species == "BLK CRP"] <- "BLC" # Black Crappie
datframe$Species[datframe$Species == "CCF "] <- "CCF" # Channel Catfish
datframe$Species[datframe$Species == "CFF"] <- "CCF" # channel catfish
datframe$Species[datframe$Species == "Crappie"] <- "BLC" # Black crappie in center lake
datframe$Species[datframe$Species == "dummy"] <- "" # used as placeholder for 0 fish transects
datframe$Species[datframe$Species == "GNSF"] <- "GSF" # green sunfish
datframe$Species[datframe$Species == "Green Sunfish"] <- "GSF" # green sunfish
datframe$Species[datframe$Species == "NOP not silver"] <- "NOP" # northern pike
datframe$Species[datframe$Species == "NOP silver"] <- "NOP" # northern pike, formerly stocked as 'silver' phenotype
datframe$Species[datframe$Species == "P-seed"] <- "PSD" # pumpkinseed
datframe$Species[datframe$Species == "Pumpkin seed"] <- "PSD" # pumpkinseed
datframe$Species[datframe$Species == "Spot tail shiner"] <- "Spottail Shiner"
datframe$Species[datframe$Species == "Spot Tail Shiner"] <- "Spottail Shiner"
datframe$Species[datframe$Species == "Spot tal shiner"] <- "Spottail Shiner"
datframe$Species[datframe$Species == "Spotted Shiner"] <- "Spottail Shiner"
datframe$Species[datframe$Species == "Sucker"] <- "WHS" # white sucker
datframe$Species[datframe$Species == "WCRP"] <- "WHC" # white crappie
datframe$Species[datframe$Species == "White Bass"] <- "WHB" # white bass
datframe$Species[datframe$Species == "WHT bass"] <- "WHB" # white bass
datframe$Species[datframe$Species == "WHT CRP"] <- "WHC" # white crappie
datframe$Species[datframe$Species == "Yellow Bull Head"] <- "YBH" # yellow bullhead
datframe$Species[datframe$Species == "Yellow Bullhead"] <- "YBH" # yellow bullhead
datframe$Species<-as.factor(datframe$Species)
datframe$Species<-droplevels(datframe$Species)
levels(datframe$Species)


# Gear
levels(datframe$Gear)
datframe$Gear[datframe$Gear == "Electrofishing "] <- "Electroshocking"
datframe$Gear[datframe$Gear == "Electrofishing"] <- "Electroshocking"
datframe$Gear[datframe$Gear == "Electroschocking"] <- "Electroshocking"
datframe$Gear[datframe$Gear == "Fyke Net"] <- "Fyke"
datframe$Gear[datframe$Gear == "Seine "] <- "Seine"
datframe$Gear<-droplevels(datframe$Gear)
levels(datframe$Gear)

```


# remove data collected by IA DNR
#### look at google sheets where ISU actually sampled blue or twin(s)... then subset
Blue Lake exclude all except: 5/30/2019 & 6-27-2019
North Twin exclude all except: 5-17-2019 & 5-21-2019 
-- Note that  11-3-2019 seine was estimated number of fish removed by seine, not handled by me
-- 11-10-2019 tagging not included on google sheets due to previous week's violation of closure.



```{r,}
levels(datframe$Date)

# Blue Lake
datframe$Lake[datframe$Date == "5/30/19"]
datframe2<-droplevels(subset(datframe, Lake == "Blue" &
                                       Date == "5/30/19" |
                                       Lake == "Blue" &
                                       Date == "6/27/19"))
datframe<-subset(datframe, Lake != "Blue")
datframe<-rbind(datframe,datframe2)


# N Twin Lake
datframe$Lake[datframe$Date == "5/17/19"]
datframe$Lake[datframe$Date == "5/21/19"]
datframe$Lake[datframe$Date == "11/10/19"] # legitly 
datframe3<-droplevels(subset(datframe, Lake == "N Twin" &
                                       Date == "11/10/19" |
                                       Lake == "N Twin" &
                                       Date == "5/21/19" |
                                       Lake == "N Twin" &
                                       Date == "5/17/19"))
datframe<-subset(datframe, Lake != "N Twin")
datframe<-rbind(datframe,datframe3)

# S Twin Lake
datframe<-subset(datframe, Lake != "S Twin")


```

# final counts

```{r,}
# Released Alive 2019
Alive<-droplevels(subset(datframe, Notes != "euthanized"))
A.melt<-melt(Alive, id.var = c("Species","Gear","Lake"),
              measure.vars = "detect")
AliveTotal<-dcast(A.melt,Species+Lake~Gear)
AliveTotal

# Euthanized 2018
Euth<-droplevels(subset(datframe, Notes == "euthanized"))
E.melt<-melt(Euth, id.var = c("Species","Gear","Lake"),
             measure.vars = "detect")
EuthTotal<-dcast(E.melt, Species+Lake~Gear)
EuthTotal

```
