---
title: "Stat 534 Homework 1"
author: "Marty Simonson"
date: "October 3, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# 1.) Alseis blackiana

Given a data file on abundance of _Alseis blackiana_ in two habitats (OldHigh and OldSlope). Assume that the counts follow a Negative Binomial distribution, and paramaterize in terms of the mean $\mu$ and overdispersion parameter $\theta$.

## a) 

Ignore habitat and find the MLE of mean and $\theta$, the overdispersion parameter. 

```{r,include=F}

packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}
setwd("C:/Users/martysim/Documents/BMB-COC/")
ab<-read.csv("STAT534/Ab.csv",header=T)
str(ab)
```
```{r,echo=F}
NBlnl<- function(beta,data) {
  mu <- beta[1]
  size <- beta[2]
  sum(dnbinom(data,size,mu=mu,log=t))
}

# Numerical MLE for mu and theta

one.a<-optim(c(20,5), NBlnl, method = 'BFGS', data = ab$Count,
      control = list(fnscale=-1))
mu<-one.a$par[1]
size<-one.a$par[2]

theta<- ((mu+(mu^2/size))-mu)/mu^2

mu
theta
```

The overall numerical optimization of the maximum likelihood parameters $\mu$ and $\theta$ are 20.400 and 0.307, respectively.

## b)

Fit a model that allows habitat-specific means and shared overdispersion. Find the mle's of the difference in means (as OldHigh - OldSlope) and the overdispersion.

```{r,echo=F}

NBlnl2<- function(beta, data) {
  mu<-beta[1] + data$X*beta[2]
  size <- beta[3]
  sum(dnbinom(data$Count,size,mu=mu,log=T))
}

#levels(ab$Habitat)
ab$X[ab$Habitat == "OldSlope"]<-0
ab$X[ab$Habitat == "OldHigh"]<-1

fit.b<-optim(c(20,
               20,
               5),
             NBlnl2,method='BFGS',data=ab,
             control=list(fnscale=-1),
             hessian=T)
mu.difference<-fit.b$par[2]
theta<-1/fit.b$par[3]

mu.difference

theta

```

The MLE for the difference in means between OldHigh - OldSlope is 17.875 and the MLE of the overdispersion parameter theta is 0.14


## c) 

Use the results from (b) to construct a normal approximation 95% confidence interval for the difference in means.

```{r,echo=F}
betahat<-fit.b$par
fit.b.vc<-solve(-fit.b$hessian)
fit.b.se<-sqrt(diag(fit.b.vc))

mu.difference.ci<-cbind(
  lower = betahat - qnorm(0.975)*fit.b.se,
  upper = betahat + qnorm(0.975)*fit.b.se)
mu.difference.ci[2,]
```

The 95% normal approximation for the difference in means between habitat is between 8.255 to 27.494.

## d)

Use results from this model (and perhaps other models) to construct a likelihood ratio test __LRT__ of the $H_0: \mu_{OldHigh} = \mu_{OldSlope}$. Report the value of the test statistic and and p-value.

```{r}
# reduced - full model
LRT1<- -2*(one.a$value - fit.b$value)
LRT1

pval.lrt1<-pchisq(LRT1, df=1, lower.tail=FALSE)
pval.lrt1
```

The test statistic is 13.507 with an associated p-value of 0.000238, leading us to the conclusion that there is evidence to reject hte null hypothesis that the mean counts between habitats are different.


## e)

Fit a model (or models) that allows habitat-specific means and habitat specific overdispersion. Find the MLE's of the two means and the two overdispersion parameters.

```{r, echo = F}
NBlnl3<- function(beta, data) {
  mu<-beta[1] + data$X*beta[2]
  size <- beta[3] + data$X*beta[4]
  sum(dnbinom(data$Count,size,mu=mu,log=T))
}

# We have 4 params, order is important (dummy variable X = 1 for Old High habitat)
# beta1 is mu for Old Slope
# beta2 is difference in mu between old slope and old high
# beta3 is theta (size) for old Slope
# beta4 is the difference in theta between old slope and old high

fit.e<-optim(c(10, 10, 5, 1),
             NBlnl3,
             method='BFGS',
             data=ab,
             control=list(fnscale=-1),
             hessian=T)
#fit.e


mu.diff<-fit.e$par[2]
theta.diff<-fit.e$par[4]


theta.oldSlope<-1/fit.e$par[3]
theta.oldHigh<-1/(fit.e$par[3] + theta.diff)

theta.oldSlope
theta.oldHigh

mu.oldSlope<-fit.e$par[1]
mu.oldHigh<-mu.oldSlope+mu.diff

mu.oldSlope
mu.oldHigh

```

The MLE's for the mean in habitats Old Slope and Old High are 13.249 and 31.127, respectively. The MLE's for overdispersion in Old Slope and Old High habitats are 0.0997 and 0.181, respectively.


## f)

Use results from this model (and perhaps other models) to construct a likelihood ratio test of $H_0: \theta_{OldHigh} = \theta_{OldSlope}$. Report the value of your test statistic and a p-value.

```{r}
# -2 * (reduced - full model)
LRT2<- -2*(fit.b$value - fit.e$value)
LRT2

pval.lrt2<-pchisq(LRT2, df=1, lower.tail=FALSE)
pval.lrt2
```

The test statistic is 0.461 with an associated p-value of 0.497, suggesting that there is no evidence to support the alternative hypothesis that there is a difference in $\theta$ between the two habitats.

# 3 Missouri Fish Communities

- (a) Is there an interaction between site and habitat, i.e., is the difference between the two habitats similar (or not) at the three sites?

- (b) If there is an interaction, which species are significant contributors to that interaction?

```{r, include = F}
packages("mvabund")
packages("vegan")
setwd("C:/Users/martysim/Documents/BMB-COC/")
fish.env<-read.csv("STAT534/newfishenv.csv")
fish<-read.csv("STAT534/newfish.csv")
str(fish) # columns are unique species, rows are sample events, cells are counts
str(fish.env)
# set up matrix in mvabund form:
fish.mva<-mvabund(fish)


#first model: habitat
fish.glm<-manyglm(fish.mva ~ habitat + site + gear + site*gear + gear*habitat + site*habitat, data = fish.env)
fish.glm0<-manyglm(fish.mva ~ habitat + site + gear + habitat*gear + gear*site, data = fish.env)

# species-specific lnL and AIC
logLik(fish.glm)
logLik(fish.glm0)
AIC(fish.glm)
AIC(fish.glm0)

# estimated overdispersion, as phi: Var = mu + phi*mu^2
fish.glm$phi

# residual plot
#plot(fish.glm)
```

## Answer:

### Methods

Missouri fish sampling was conducted where two habitats (main channel and chute) were sampled at each of three sites using five different gears. Species of fish found in fewer than 10% of samples have been omitted from the data. The objective is to test whether or not species abundance is affected by an interaction between site and habitat. If an interaction exists, we will determine which species are significant contributers to the observed interaction of site and habitat.

I assume that since we are given count data that our models are overdispersed and will create a model under the negative binomial distribution. I further assume that each species is independent.

The full model includes additive effects of habitat, site, and gear as well as interactions of site:gear, gear:habitat, and our parameter of interest site:habitat. The reduced (nested) model included the same individual effects of habitat, site, and gear but also had interactions of habitat:gear and gear:site. 

Both models were implemented using the R packages "mvabund" and "vegan", particularly the function _manyglm()_ that is used to calculate parameter estimates of generalized linear models fitted to each of many variable simultaneously. A lon-link function was used with the default negative binomial distribution. Significance for the interaction between site and habitat was tested using in the full model using a bootstrap analysis of variance function set to 999 iterations (known as PIT-trap resampling). 

Fish species that were significant contributors to any observed interaction between site and habitat were determined by plotting the deviance, or likelihood ratio test, between our full and reduced models for each species. The threshold for significance was determined first by a Bonferroni-corrected chi-square test and then by a traditional chi-square test, each with $\alpha = 0.05$ and two degrees of freedom.

### Results

The ANOVA for the full model showed that there was some evidence for an interaction between site and habitat (p-value ~ 0.05). The model for 'gear' was highly significant (p-value ~ 0.01) and models with 'site' and 'site:gear' were observed to be significant given the data.
```{r, echo = F}
# resampling residuals maintaining rows 999 times
anova(fish.glm)
```

Plotting the deviance between models yielded zero species that significantly contributed to the observed interaction between site and habitat under the Bonferroni-adjusted chi-square test (dotted line). However, the traditional chi-square threshold (dahsed line) determined that there were four species that were likely significant contributors to the observed interaction between sample site and sample habitat.
```{r, echo = F}
# Plot the deviance, or -2lnL, or likelihood ratio test, between our full and reduced (nested) model.
plot.new()
plot(1:30, 2*(logLik(fish.glm)-logLik(fish.glm0)),
     xlab = 'Fish Code',ylab='Deviance', main = 'Deviance between site:habitat model and reduced model',
     ylim = c(0,15))
abline(h=qchisq(1-(1-0.95)/30,2),lty=3) # bonferroni adjustment
abline(h=qchisq(0.95,2),lty=5) # traditional
```

The four species that were common carp  _Cyprinus carpio_, channel catfish _Ictalurus punctatus_, red shiner _Cyprinella lutrensis_, and  and silver chub _Machrybopsis sotreriana_. The first two species are large-bodied bottom feeders, and may use habitat differently based on water current (chute vs. main channel). The other two species are small-bodied individuals and may not have the strength to withstand current in the main channel.