---
title: "Ecostats Homework 4"
date: "November 18, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=F}
# Front matter
rm(list=ls())  # removes objects from environment

# Function to load and install packages
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

# Setting Working Directory
setwd("C:/Users/martysim/Documents/BMB-COC/STAT534/")

# setting up packages
packages(ggplot2)
packages(RMark)
packages(gtools)
packages(popbio)
```

# 1) Shrub transition martices

## (a) Asymptotic growth rate

```{r, include = F}
shrub<-read.csv("shrub0.csv",header=T)
str(shrub)

ev<-eigen(shrub)
lambda1<-Re(ev$values[1])
lambda1
```

- The asymptoic growth rate, $\lambda$, is 0.94.

## (b) Stable age distribution

```{r, include = F}
u<-Re(ev$vectors[,1]) # associated real eigenvector
u<-u/sum(u)
u
```

- The stable age distribution for stages 1-8 are 0.28, 0.03, 0.53, 0.01, 0.06, 0.06, 0.02, and 0.01, respectively.

## (c) Stable age distribution ignoring seeds

```{r, include = F}
uc<-Re(ev$vectors[,1]) # associated real eigenvector
uc
uc<-uc[-1]
uc<-uc/sum(uc)
sum(uc)
uc
```

- The stable age distribution for stages 2-8 (ignoring seeds) is 0.04, 0.73, 0.02, 0.08, 0.09, 0.02, and 0.01, respectively

## (d) Increasing matrix elements

```{r, include = F}
ev<-eigen(shrub)
u<-Re(ev$vectors[,1]) # associated real eigenvector
u<-u/sum(u)
u
# then repeat for transpose of a to get left eigen vectors
ev <- eigen(t(shrub));

v <- Re(ev$vectors[,1])
v <- v/sum(v);

# sensitivity matrix
s <- outer(v,u) / sum(v*u)
s # element row 5 col 3 is largest
# but that does not correspond to a non-zero element in the shrub matrix
# that's OK according to one resource
# it's like making sub-adults reproduce or something
s[5,3]
```

- The largest element in the sensitivity matrix is row 5, column 3: 2.90. Since this is a rate(slope), changing that element by 110% (1.1) would result in a change in $\lambda$ of 3.19.

## (e) Effect of canopy cover

```{r, include = F}
# function LTRE from package 'popbio'
shrub25<-read.csv("shrub25.csv",header=T)
str(shrub25)
g<-LTRE(shrub25,shrub)

## the sensitivity matrix approach
ev<-eigen(shrub25)
u<-Re(ev$vectors[,1]) # associated real eigenvector

lambda2<-Re(ev$values[1])
lambda2

u<-u/sum(u)
u
# then repeat for transpose of a to get left eigen vectors
ev <- eigen(t(shrub25));

v <- Re(ev$vectors[,1])
v <- v/sum(v);

# sensitivity matrix
s <- outer(v,u) / sum(v*u)
s

```

The elements [8,1] = 237; [7,1] = 160; [8,7] = 89 are the three highest values in the sensitivity matrix.

## (f) part one: LTRE and lambda

```{r, include = F}
difference<-lambda2-lambda1
difference

sum(s)
```

not even a little close to each other

## (f) part two: transition matrix logic

```{r}
shrub
```

well, the matrix as given shows a plant that flowered in 2010 was a seed in 2011 that had to transition to seedling with 0.1 in 2012. So two years. CONTRARY to reasoning as seeds produced in 2010 would become seedlings in 2011.

## (g) re-written transition matrix

```{r}
# my idea is to multiply fecundity by 0.1 for sm-vlg
shrub2<-shrub[1,]*0.1
shrub2<-rbind(shrub2,shrub[2:8,])
# trim first column and second row
shrub2<-shrub2[-2,-1]
shrub2
```

## (h) lambda with new transition matrix

```{r}
ev<-eigen(shrub2)

lambda3<-Re(ev$values[1])
lambda3

# stable age distribution
u2<-Re(ev$vectors[,1]) # associated real eigenvector
u2<-u2/sum(u2)
round(u2,2)
```

## (i)

```{r}
round(uc-u2, 4)
```

The stable age distributions between 1(c) and 1(h) are very similar, in fact the ratios are only off by a few hudredths at most. The largest difference is the method from part (h) has more 3% more 'lg' in the stable age distribution.

# 2) Shrubs and hurricanes

```{r}
shrubEnv<-read.csv("allshrubRev.csv")
str(shrubEnv)

allAdata <- shrubEnv

allA <- array(NA, dim=c(8,8,4))
# empty array

allA[,,1] <- as.matrix(allAdata[allAdata$Env==1, 3:10])
allA[,,2] <- as.matrix(allAdata[allAdata$Env==2, 3:10])
allA[,,3] <- as.matrix(allAdata[allAdata$Env==3, 3:10])
allA[,,4] <- as.matrix(allAdata[allAdata$Env==4, 3:10])
# extract each enviroment, convert to matrix, and store in allA

allA




hurricane<-read.csv("hurricane.csv")
hurricane


rand.env <- function(B,allA,nburn=100,nsim=1000) {
# function to simulate a random sequence of
#    environments, then use that to estimate stochastic lambda

#  input:
#  B: matrix that specifies the environmental transition matrix
#   each col gives P[move to env i | currently in that env]
#    i.e. B21 give P[move to env 2 | currently in env 1]

#  allA: array giving all transition matrices, A in each env
#    dimensions: k x k x e 
#    where e is number of environments, and k x k is dimension of A

# optionally, can specify # iterates to ignore (nburn)
#   and # iterates to use to estimate stochastic lambda (nsim)

  lambda <- env <- rep(0,nburn+nsim)
  env[1] <- 1		# arbitrarily start in env 1

  N <- matrix(0,nrow=dim(allA)[1],ncol=nburn+nsim)
  N[1,1] <- 1			# Initialize N to (1,0,...)

  for (i in 2:(nburn+nsim)) {
    curr.env <- 1+sum(cumsum(B[,env[i-1]]) < runif(1)) 
    env[i] <- curr.env

    curr.A <- allA[,,curr.env]

    N[,i] <- curr.A %*% N[,i-1]
    lambda[i] <- sum(N[,i])

    N[,i] <- N[,i]/sum(N[,i])
    }

 logl <- log(lambda[-(1:nburn)])
 meanlogl <- mean(logl)
 varlogl <- var(logl)

 list(env=env, N = N[,-(1:nburn)], lambda=lambda, 
   meanlogl = meanlogl, varlogl = varlogl)
}

one.a<-rand.env(hurricane,allA,nburn=100,nsim=1000)
exp(one.a$meanlogl)
one.a$varlogl
```

The population seems to be growing by about 4.35% annually.

## (b) variance of lambda

The variance of stochastic log lambda reports the precision of year to year variability, since it is a function of the log-transformed lamda values. This is the squared standard deviation of the distribution of mean lambdas.

## (c) Re-written environmental variability matrix

```{r, include = F}
# get new hurricane matrix that allows hurricanes to affect any site
hurricane

hurricane2<-hurricane[]

#


```

