---
title: "Ecostats Homework 4"
date: "November 22, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=F}
# Front matter
rm(list=ls())  # removes objects from environment

# Function to load and install packages
packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

# Setting Working Directory
setwd("C:/Users/martysim/Documents/BMB-COC/STAT534/")

# setting up packages
packages(ggplot2)
packages(RMark)
packages(gtools)
packages(popbio)
```

# 1) Shrub transition martices

## (a) Asymptotic growth rate

```{r, include = F}
shrub<-read.csv("shrub0.csv",header=T)
str(shrub)

ev<-eigen(shrub)
lambda1<-Re(ev$values[1])
lambda1
```

- The asymptoic growth rate, $\lambda$, is 0.94.

## (b) Stable age distribution

```{r, include = F}
u<-Re(ev$vectors[,1]) # associated real eigenvector
u<-u/sum(u)
round(u,2)
```

- The stable age distribution for stages 1-8 are 0.28, 0.03, 0.53, 0.01, 0.06, 0.06, 0.02, and 0.01, respectively.

## (c) Stable stage distribution ignoring seeds

```{r, include = F}
shrub
shrub.noseed<-shrub[-2,-1]
#shrub.noseed[1,1]<-.1
ev<-eigen(shrub.noseed)
uc<-Re(ev$vectors[,1]) # associated real eigenvector
uc
uc<-uc/sum(uc)
sum(uc)
round(uc,2)
l.c <- Re(ev$values[1])	# lambda, convert to real
l.c
```

- The stable stage distribution for stages 2-8 (ignoring seeds) is 0.09, 0.82, 0.02, 0.04, 0.02, 0.00, and 0.00, respectively.

## (d) Increasing matrix elements

```{r, include = F}
ev<-eigen(shrub)
u<-Re(ev$vectors[,1]) # associated real eigenvector
u<-u/sum(u)
u
# then repeat for transpose of a to get left eigen vectors
ev <- eigen(t(shrub));

v <- Re(ev$vectors[,1])
v <- v/sum(v);

# sensitivity matrix
s <- outer(v,u) / sum(v*u)
s # element row 5 col 3 is largest
l <- Re(ev$values[1])	# lambda, convert to real

# elasticity matrix
e<-shrub*s/l
e
# the largest value in the elasticity matrix is [5,5]
# 0.3178

# 10% increase of element [5,5] according to the elatsicity matrix, is
1.1*e[5,5]

# now change that value in the original shrub
shrub.d<-shrub
new.param<-shrub.d[5,5]*1.1
shrub.d[5,5]<-new.param
# new lambda
ev<-eigen(shrub.d)
lambda.d<-Re(ev$values[1])
lambda.d
lambda.d-lambda1
```

- The largest element in the sensitivity matrix is row 5, column 5: 0.35. The new lambda is 1.01, which is an increase of 0.06 from the first lambda in part (a).

## (e) Effect of canopy cover

```{r, include = F}
# function LTRE from package 'popbio'
shrub25<-read.csv("shrub25.csv",header=T)
str(shrub25)
ev<-eigen(shrub25)
lambda.e<-Re(ev$values[1])
lambda.e

a.star<-(shrub+shrub25)/2
a.star

ev<-eigen(a.star)
u<-Re(ev$vectors[,1]) # associated real eigenvector



u<-u/sum(u)
u
# then repeat for transpose of a to get left eigen vectors
ev <- eigen(t(a.star));

v <- Re(ev$vectors[,1])
v <- v/sum(v);

# sensitivity matrix
s <- outer(v,u) / sum(v*u)
s

rcm<-(shrub25-shrub)*s
rcm
sum(rcm)
```

- The elements [4,3] = 0.241; [1,5] = 0.087; [1,4] = 0.084 are the three highest values in the relative contribution to lambda matrix.

## (f) part one: LTRE and lambda

```{r, include = F}
difference<-lambda.e-lambda1
difference
sum(rcm)
difference-sum(rcm)
```

- These values are fairly close: the calculated difference between lambda for both transition matrices equals 0.618, while the estimated difference in lambda from the LTRE is 0.560. These values differ by about 0.058.

## (f) part two: transition matrix logic

- Well, the matrix as given shows a plant that flowered in 2010 was a seed in 2011 that had to transition to seedling with 0.1 in 2012. So it takes two years.

## (g) re-written transition matrix

```{r, include = F}
# my idea is to multiply fecundity by 0.1 for sm-vlg
shrub.f<-shrub[1,]*0.1
shrub2<-shrub[2:8,]
shrub2[1,]<-shrub.f
# trim first column and second row
shrub2<-shrub2[-1]
```

```{r}
shrub2
```

## (h) lambda with new transition matrix

```{r, include = F}
ev<-eigen(shrub2)

lambda3<-Re(ev$values[1])
lambda3

# stable age distribution
u2<-Re(ev$vectors[,1]) # associated real eigenvector
u2<-u2/sum(u2)
round(u2,2)
```

- The asymptotic growth rate is 0.94 and the stable stage distribution is 0.04, 0.73, 0.02, 0.08, 0.09, 0.02, 0.01. for stages seedling-large adult, respectively.

## (i) comparing the two no-seed models

```{r}
round(uc-u2, 4)
round(l.c-lambda3, 4)
```

- For the method from part (c), the first two stages (seedling and juvenile) have higher relative abundances than the distribution from part (h) by 5% and 9%, respectively. The pre-adult stage is nearly the same in both approaches. All the adult stages have lower relative abundances in part (c) compared to (h), by about 1% to 7%. The population growth rate from part (c) is about 4% larger than in part (h). Since both the stable stage distribution and the population growth rates are roughly similar, this shows that there are many ways to organize a transition matrix and obtain similar estimates. Being thoughtful of the biology and the research question will inform how a transition matrix should be constructed.

# 2) Shrubs and hurricanes

```{r, include = F}
shrubEnv<-read.csv("allshrubRev.csv")
str(shrubEnv)

allAdata <- shrubEnv

allA <- array(NA, dim=c(8,8,4))
# empty array

allA[,,1] <- as.matrix(allAdata[allAdata$Env==1, 3:10])
allA[,,2] <- as.matrix(allAdata[allAdata$Env==2, 3:10])
allA[,,3] <- as.matrix(allAdata[allAdata$Env==3, 3:10])
allA[,,4] <- as.matrix(allAdata[allAdata$Env==4, 3:10])
# extract each enviroment, convert to matrix, and store in allA

allA




hurricane<-read.csv("hurricane.csv")
hurricane


rand.env <- function(B,allA,nburn=100,nsim=1000) {
# function to simulate a random sequence of
#    environments, then use that to estimate stochastic lambda

#  input:
#  B: matrix that specifies the environmental transition matrix
#   each col gives P[move to env i | currently in that env]
#    i.e. B21 give P[move to env 2 | currently in env 1]

#  allA: array giving all transition matrices, A in each env
#    dimensions: k x k x e 
#    where e is number of environments, and k x k is dimension of A

# optionally, can specify # iterates to ignore (nburn)
#   and # iterates to use to estimate stochastic lambda (nsim)

  lambda <- env <- rep(0,nburn+nsim)
  env[1] <- 1		# arbitrarily start in env 1

  N <- matrix(0,nrow=dim(allA)[1],ncol=nburn+nsim)
  N[1,1] <- 1			# Initialize N to (1,0,...)

  for (i in 2:(nburn+nsim)) {
    curr.env <- 1+sum(cumsum(B[,env[i-1]]) < runif(1)) 
    env[i] <- curr.env

    curr.A <- allA[,,curr.env]

    N[,i] <- curr.A %*% N[,i-1]
    lambda[i] <- sum(N[,i])

    N[,i] <- N[,i]/sum(N[,i])
    }

 logl <- log(lambda[-(1:nburn)])
 meanlogl <- mean(logl)
 varlogl <- var(logl)

 list(env=env, N = N[,-(1:nburn)], lambda=lambda, 
   meanlogl = meanlogl, varlogl = varlogl)
}

one.a<-rand.env(hurricane,allA,nburn=10000,nsim=100000)
```
```{r}
round(one.a$meanlogl, 2)
round(one.a$varlogl, 2)
```

The population seems to be growing by about 6% annually.

## (b) variance of lambda

The variance of the stochastic log lambda is estimated from the collection of 90,000 lambda values of all simulated years, which captures year-to-year variability. It does not factor in sample size, which would be needed for standard error.  

## (c) Re-written environmental variability matrix

```{r, include = F}
# get new hurricane matrix that allows hurricanes to affect any site
hurricane

hurricane2<-hurricane
hurricane2[1,2]<-0.95
hurricane2[4,2]<-0.05
hurricane2[2,3]<-0.95
hurricane2[4,3]<-0.05
hurricane2[3,4]<-0.95
hurricane2[4,4]<-0.05
#
one.c<-rand.env(hurricane2,allA,nburn=10000,nsim=100000)
```
```{r,}
hurricane2
round(one.c$meanlogl, 2)
round(one.c$varlogl,2)
```

## (d)

```{r, include = F}
hurricane

hurricane3<-hurricane
hurricane3[1,1]<-0.99
hurricane3[4,1]<-0.01
hurricane3[1,2]<-0.99
hurricane3[4,2]<-0.01
hurricane3[2,3]<-0.99
hurricane3[4,3]<-0.01
hurricane3[3,4]<-0.99
hurricane3[4,4]<-0.01

one.d<-rand.env(hurricane3,allA,nburn=10000,nsim=100000)
```

```{r,}
hurricane3
round(one.d$meanlogl,2)
round(one.d$varlogl,2)
```

## (e)

```{r, include = F}
hurricane4<-hurricane2
hurricane4

hurricane4[1,4]<-0.5
hurricane4[2,4]<-0.45
hurricane4[3,4]<-0
hurricane4[4,4]<-0.05

one.e<-rand.env(hurricane4,allA,nburn=10000,nsim=100000)
```

```{r,}
hurricane4
round(one.e$meanlogl,2)
round(one.e$varlogl,2)
```

