---
title: "Marcus Age Comparison Plots"
author: "Martin Simonson"
date: "April 19, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r,include=F}
rm(list=ls())

packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

packages(vegan)
packages(ggplot2)
packages(reshape2)
packages(nnet)
packages(multcomp)
packages(plyr)
packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(gridExtra)
packages(dplyr)
packages(scales)
```


```{r,}
## Age comparison plots for North Twin Lake
#
# Buffalo and Carp
#
#
age<-read.csv("Marcus/BuffaloAges_Compiled.csv",header=T)
summary(age)
head(age)

##############################################################

# Marcus
        # Pectoral v Dorsal
p1<-ggplot(age, aes(carp,x=Marcus_Pectoral, y=Marcus_Dorsal))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "Reader 1",
       x = "",
       title ="Dorsal Spines")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)+
  theme(plot.title = element_text(hjust = 0.5))

p1

model<-lm(Marcus_Dorsal~-1 + Marcus_Pectoral, data = age)
summary(model)
confint(model)





        # Pectoral v sectioned otolith
p2<-ggplot(age, aes(x=Marcus_Pectoral, y=Marcus_SectOto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "",
       x="",
       title = "Sectioned Otoliths")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)+
  theme(plot.title = element_text(hjust = 0.5))

p2

model<-lm(Marcus_SectOto~Marcus_Pectoral, data = age)
summary(model)
confint(model)

        # Pectoral v cracked otlith
p3<-ggplot(age, aes(x=Marcus_Pectoral, y=Marcus_CrackOto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "",
       x = "",
       title = "Cracked and Sanded Otoliths")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)+
  theme(plot.title = element_text(hjust = 0.5))

p3

model<-lm(Marcus_CrackOto~Marcus_Pectoral, data = age)
summary(model)
confint(model)


##############################################################
        

# Drake


        # dorsal
p4<-ggplot(age, aes(carp,x=Drake_Pectoral, y=Drake_Dorsal))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "Reader 2",
       x = "")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p4

model<-lm(Drake_Dorsal~Drake_Pectoral, data = age)
summary(model)
confint(model)



        # sectioned otlith
p5<-ggplot(age, aes(x=Drake_Pectoral, y=Drake_SectOto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "",
       x = "")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p5

model<-lm(Drake_SectOto~Drake_Pectoral, data = age)
summary(model)
confint(model)

        # cracked otolith
p6<-ggplot(age, aes(x=Drake_Pectoral, y=Drake_Crackoto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(y = "",
       x = "")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p6

model<-lm(Drake_Crackoto~Drake_Pectoral, data = age)
summary(model)
confint(model)

##############################################################

# Marty

        # Dorsal
p7<-ggplot(age, aes(carp,x=Marty_Pectoral, y=Marty_Dorsal))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  labs(x = "Pectoral Age",
       y = "Reader 3")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p7

model<-lm(Marty_Dorsal~Marty_Pectoral, data = age)
summary(model)
confint(model)

        # sectioned otoliths
p8<-ggplot(age, aes(x=Marty_Pectoral, y=Marty_SectOto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  geom_smooth(method = "lm")+
  labs(x = "Pectoral Age",
       y = "")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p8

model<-lm(Marty_SectOto~Marty_Pectoral, data = age)
summary(model)
confint(model)

        # cracked otlith
p9<-ggplot(age, aes(x=Marty_Pectoral, y=Marty_Crackoto))+
  geom_point(size=5, alpha = 0.2)+
  geom_abline(slope = 1, intercept = 0)+
  geom_smooth(method="lm")+
  labs(x = "Pectoral Age",
       y = "")+
  theme_classic()+
  xlim(0,20)+
  ylim(0,20)
p9

model<-lm(Marty_Crackoto~Marty_Pectoral, data = age)
summary(model)
confint(model)

##############################################################

grid.arrange(p1,p2,p3,
             p4,p5,p6,
             p7,p8,p9,
             ncol=3)
```

```{r,}

##############################################################

#

#

#       New plots: based on fact that pectoral spines have least CV

#       We will place Pectoral Age as the common X axis (but different for each reader)

#       Each of the other three structures will be columns

#       Each of the readers will be rows

#       Therefore a 3x3 grid

#       Within each plot we want to have the 1 to 1 line,

#       then the average age estimate from the reader for the column structure with error bars

##############################################################################################
ages<-read.csv("Marcus/BuffaloAges_Compiled.csv",header=T)
head(ages)
ages$Drake_Pectoral<-as.factor(ages$Drake_Pectoral)
ages$Marcus_Pectoral<-as.factor(ages$Marcus_Pectoral)
ages$Marty_Pectoral<-as.factor(ages$Marty_Pectoral)

# trim the fat
ages<-ages[,-c(1:7)]
str(ages)
d.ages<-ages[,c(5,2,8,11)]
m.ages<-ages[,c(4,1,7,10)]
my.ages<-ages[,c(6,3,9,12)]

d.ages.m<-melt(d.ages)
d.ages.c<-dcast(d.ages.m, 
                Drake_Pectoral ~ variable, 
                fun.aggregate = mean)
d.ages.c2<-dcast(d.ages.m, 
                Drake_Pectoral ~ variable, 
                fun.aggregate = sd)
drake<-merge.data.frame(d.ages.c,d.ages.c2, by = "Drake_Pectoral")




m.ages.m<-melt(m.ages)
m.ages.c<-dcast(m.ages.m, 
                Marcus_Pectoral ~ variable, 
                fun.aggregate = mean)
m.ages.c2<-dcast(m.ages.m, 
                Marcus_Pectoral ~ variable, 
                fun.aggregate = sd)

my.ages.m<-melt(my.ages)
my.ages.c<-dcast(my.ages.m, 
                Marty_Pectoral ~ variable, 
                fun.aggregate = mean)
my.ages.c<-dcast(my.ages.m, 
                Marty_Pectoral ~ variable, 
                fun.aggregate = sd)


ggplot(ages, aes(x=))




##############################################################
ages<-ages[,-c(1:7)]
ages2<-melt(ages)

ggplot(ages, aes(x=))


```


# Coefficient of Variation computation

FOR EACH STRUCTURE:

\[
CV_j = 100\% * \frac{\sqrt{\Sigma_{i=1}^{R}\frac{(X_{ij}-X_j)^2}{R-1}}}{X_j}
\]

Where $CV_j$ is the age precision estimate for the $j$th fish. $X_{ij}$ is the $i$th age determination of the $j$th fish, $X_j$ is the mean age estimate of the $j$th fish, and $R$ is the number of times each fish is aged. 

Therefore, we need:

- $X_{ij}$: each of the 3 age estimates for each fish 

- $X_j$: the MEAN age estimate for each fish

- $R$: number of times each fish is aged. = 3

let's get to work!

```{r,}

age$j<-paste(age$Lake,age$TAG.NUMBER,sep = ".")
age$R<-3 # easy one. number of times fish is aged.

age<-age[,-c(1:7)]

head(age)

# MeanAge = Xj

###########################
#
#    Dorsal
#
###########################

dorsal<-age[,c(10,1,2,3,11)]
head(dorsal)

library(dplyr)
dorsal <- dorsal %>% mutate(MeanAge=(Marcus_Dorsal+Drake_Dorsal+Marty_Dorsal)/3)
head(dorsal)

# items to sum before the square root sign of denominator of CV equation.
dorsal$MarcusDenom<-((dorsal$Marcus_Dorsal-dorsal$MeanAge)^2)/(dorsal$R-1)
dorsal$DrakeDenom<-((dorsal$Drake_Dorsal-dorsal$MeanAge)^2)/(dorsal$R-1)
dorsal$MartyDenom<-((dorsal$Marty_Dorsal-dorsal$MeanAge)^2)/(dorsal$R-1)
head(dorsal)

# CV
dorsal$CV<-100*((sqrt(dorsal$MarcusDenom+dorsal$MartyDenom+dorsal$DrakeDenom))/dorsal$MeanAge)
summary(dorsal$CV)
packages(psych)
describe(dorsal$CV)
###########################
#
#    Pectoral
#
###########################
head(age)
pectoral<-age[,c(10,4,5,6,11)]
head(pectoral)

library(dplyr)
pectoral <- pectoral %>% mutate(MeanAge=(Marcus_Pectoral+Drake_Pectoral+Marty_Pectoral)/3)
head(pectoral)

# items to sum before the square root sign of denominator of CV equation.
pectoral$MarcusDenom<-((pectoral$Marcus_Pectoral-pectoral$MeanAge)^2)/(pectoral$R-1)
pectoral$DrakeDenom<-((pectoral$Drake_Pectoral-pectoral$MeanAge)^2)/(pectoral$R-1)
pectoral$MartyDenom<-((pectoral$Marty_Pectoral-pectoral$MeanAge)^2)/(pectoral$R-1)
head(pectoral)

# CV
pectoral$CV<-100*((sqrt(pectoral$MarcusDenom+pectoral$MartyDenom+pectoral$DrakeDenom))/pectoral$MeanAge)
summary(pectoral$CV)
describe(pectoral$CV)

###########################
#
#    sectioned otoliths
#
###########################
head(age)
SectOto<-age[,c(10,7,8,9,11)]
head(SectOto)

library(dplyr)
SectOto <- SectOto %>% mutate(MeanAge=(Marcus_SectOto+Drake_SectOto+Marty_SectOto)/3)
head(SectOto)

# items to sum before the square root sign of denominator of CV equation.
SectOto$MarcusDenom<-((SectOto$Marcus_SectOto-SectOto$MeanAge)^2)/(SectOto$R-1)
SectOto$DrakeDenom<-((SectOto$Drake_SectOto-SectOto$MeanAge)^2)/(SectOto$R-1)
SectOto$MartyDenom<-((SectOto$Marty_SectOto-SectOto$MeanAge)^2)/(SectOto$R-1)
head(SectOto)

# CV
SectOto$CV<-100*((sqrt(SectOto$MarcusDenom+SectOto$MartyDenom+SectOto$DrakeDenom))/SectOto$MeanAge)
summary(SectOto$CV)
describe(SectOto$CV)
###########################
#
#    Cracked and Sanded otoliths
#
###########################
age<-read.csv("Marcus/CrackSandOnly.csv",header=T)
head(age)
age$j<-paste(age$Lake,age$Slide..,sep = ".")
age$R<-3 # easy one. number of times fish is aged.

CrackOto<-age[,-c(1,2,6)]
head(CrackOto)
str(CrackOto)

library(dplyr)
CrackOto <- CrackOto %>% mutate(MeanAge=(Marcus+Drake+Marty)/3)
head(CrackOto)

# items to sum before the square root sign of denominator of CV equation.
CrackOto$MarcusDenom<-((CrackOto$Marcus-CrackOto$MeanAge)^2)/(CrackOto$R-1)
CrackOto$DrakeDenom<-((CrackOto$Drake-CrackOto$MeanAge)^2)/(CrackOto$R-1)
CrackOto$MartyDenom<-((CrackOto$Marty-CrackOto$MeanAge)^2)/(CrackOto$R-1)
head(CrackOto)

# CV
CrackOto$CV<-100*((sqrt(CrackOto$MarcusDenom+CrackOto$MartyDenom+CrackOto$DrakeDenom))/CrackOto$MeanAge)
summary(CrackOto$CV)
library(psych)
describe(CrackOto$CV)

```
