---
title: "Live-Dead Input Files"
author: "Martin Simonson"
date: "3/10/2020"
output: word_document
editor_options: 
  chunk_output_type: console
---


# front matter
```{r,include=FALSE}
rm(list=ls())
```

```{r,}

packages<-function(x, repos="http://cran.r-project.org", ...){
x<-as.character(match.call()[[2]])
if (!require(x,character.only=TRUE)){
install.packages(pkgs=x, repos=repos, ...)
require(x,character.only=TRUE)
}
}

#packages(rcompanion)
#packages(vegan)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(FSA)
packages(nnet)
#packages(multcomp)
#packages(plyr)
#packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(FSAdata)
packages(nlstools)
#packages(GGally)
packages(wesanderson)
#packages(DescTools)
#packages(mvnormtest)
#packages(HH)
#packages(biotools)
#packages(car)
#packages(psych)
#packages(GPArotation)
packages(doBy)
packages(dplyr)
packages(tidyr)
packages(rcompanion)
packages(GISTools)
packages(maps)
packages(mapproj)
packages(ggpmisc)
```

#loading in data

```{r}
long.LD<-read.csv("2020/National AFS Meeting/Buffalo/LiveDead_LongFormat.csv",header=T)
str(long.LD)
long.LD$Year<-as.factor(long.LD$Year)
```

# converting to shorter capture history in LD format

- need to keep lake variable
- could drop carp at first to make figuring it out easier
- length may not be important
- tag loss most likely irrelevant
- Gear is there to help ID when fish were recovered dead
- Year is factor at the moment, can also set year.month column if collapsing capture history to months
- good luck!

```{r}

# create dummy capture data; id is animal and date is the date it was captured or recaptured
levels(long.LD$Gear)
df=droplevels(subset(long.LD, Gear !="COMMERCIAL HARVEST")) # live encounters stored in "df"
ddf=droplevels(subset(long.LD, Gear == "COMMERCIAL HARVEST")) # dead recoveries stored in "ddf"


#create date intervals that accompany 
levels(ddf$Date)
intervals=as.Date(c("2017-04-01"))
intervals





# cut the dates into intervals
occasions=cut(df$date,intervals)



#create the count table with id for rows and years for columns
## dch is for dead recoveries
ch=with(df,table(id,occasions))



# can be caught multiple times in an occasion; change all >0 to 1
ch[ch>0]=1
ch


# create capture history as a string
chstr=apply(ch,1,paste,collapse="")




# to match the id and occasions between ch and dch
recapture$id=factor(as.character(recapture$id),levels=levels(capture$id))




fullmat=matrix(NA,nrow=nrow(ch),ncol=2*ncol(ch))
fullmat[,seq(1,2*ncol(ch),2)]=ch
fullmat[,seq(2,2*ncol(ch),2)]=dch
fullch=apply(fullmat,1,paste,collapse="")


# create dataframe for Rmark....
markdata=data.frame(ch=chstr,stringsAsFactors=FALSE)





```




